---
title: "Battle Creek"
author: "[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com), [Erin Cain](mailto: ecain@flowwest.com)"
date: "August 16, 2019"
output: html_document
---
```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, out.width = '100%')
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
```
## Future Data Improvements
None Planned

## Instream Spawning and Rearing Habitat
**Data Source:** [Thomas R. Payne & Associates 1995](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/Payne1995_BattleCreekIFIM.pdf){target="_blank"} (pg C-1)

Instream spawning and rearing habitat for Fall Run Chinook Salmon and Steelhead in Battle Creek are based on data from Thomas R. Payne & Associates instream flow evaluations conducted for the California Department of Fish and Wildlife. These evaluations developed relationships between flow and suitable spawning and rearing habitat for Battle Creek from seven study reaches. We used the consolidated WUA for all mainstem reaches for Fall Run Chinook and Steelhead data from Thomas Payne and Associates 1995 located in this spreadsheet (see'~cvpiaHabitat/data-raw/weighted_composite_battle.csv).   

### Combine Spawning and Rearing Data
The instream spawning and rearing habitat data described above for Fall Run Chinook Salmon (FR) and Steelhead (ST) is combined for use in the DSM in the following format. Units are in square feet per 1000 feet. 
```{r}
battle_ST <- read.csv('data-raw/watershed/battle/data/Weighted_composite_Battle.csv') 

battle_creek_instream <- battle_ST %>% 
  select(flow_cfs = Flow, FR_spawn_wua = Spawning,
         FR_fry_wua = 'Fry.Rearing', FR_juv_wua = 'Juv.Rearing',
         ST_adult_wua = 'adult.trout', ST_spawn_wua = 'spawning.trout.ST',
         ST_fry_wua = 'fry.rearing.ST', ST_juv_wua = 'juv.rearing.ST') %>%
  mutate(watershed = 'Battle Creek')

knitr:: kable(head(battle_creek_instream, 5), align = 'c', 
              caption = "Header Descriptions: flow_cfs = flow in cubic feet per second,
              FR_spawn_wua = Fall Run Chinook Spawning wua,  FR_fry_wua = 
              Fall Run Chinook Fry wua, FR_juv_wua = Fall Run Chinook Juvenile wua, 
              ST_adult_wua = Steelhead Adult wua, ST_spawn_wua = Steelhead Spawning wua,
              ST_fry_wua = Steelhead Fry wua, ST_juv_wua = Steelhead Juvenile wua") 

# usethis::use_data(battle_creek_instream, overwrite = TRUE)
```

*...with 30 more rows.* 

### Spawning WUA    
The following plot shows the weighted usable spawning area in square feet per thousand feet of river for Fall Run Chinook Salmon and Steelhead. These area per length rates are multiplied by the total spawning reach length mapped by the SIT.    
```{r, echo=FALSE}
battle_creek_instream %>% 
  select(flow_cfs, `Fall Run Chinook` = FR_spawn_wua, `Steelhead` = ST_spawn_wua) %>% 
  gather(Species, WUA, -flow_cfs) %>% 
  filter(!is.na(WUA)) %>% 
  ggplot(aes(flow_cfs, WUA, color = Species)) +
  geom_line() +
  labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') + 
  theme_minimal() + 
  scale_color_manual(values = c('#d95f02','#7570b3')) +
  theme(legend.justification = c(1,0), legend.position = c(.9,.7))

```

### Rearing WUA
The following plot shows the weighted usable rearing area in square feet per thousand feet of river for Fall Run Chinook Salmon and Steelhead fry and juvenile. These rates are multiplied by the total rearing reach length mapped by the SIT. 
``` {r}
battle_creek_instream %>% 
  select(FR_juv_wua, FR_fry_wua, ST_juv_wua, ST_fry_wua, flow_cfs, watershed) %>%
  gather(species_lifestage, acres, -flow_cfs, -watershed) %>% 
  mutate(
    species = case_when( 
      str_detect(species_lifestage, 'ST') ~ 'Steelhead', 
      str_detect(species_lifestage, 'FR') ~ 'Fall Run'),
      lifestage = ifelse(str_detect(species_lifestage, 'juv'), 'Juvenile', 'Fry')) %>% 
  ggplot(aes(flow_cfs, acres, color = species)) +
  geom_line() +
  theme_minimal() +
  facet_wrap(~lifestage, ncol = 1, scales = 'free_y') +
  scale_color_brewer(palette = 'Dark2') +
  labs(x = 'Flow (cfs)', y = 'Suitable Acres') + 
  theme(legend.justification = c(1,0), legend.position = c(.95,.3))
```
```{r, include = FALSE}
source('data-raw/watershed/floodplain_utils.R')
```

## Floodplain Rearing Habitat
**Data Source**: [Scaled from a Deer Creek flow to floodplain area relationship generated with a 1D HEC-RAS hydraulic model](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/DRAFT+Deer+Creek+Hydraulic+Models+Tech+Memo+6-08-07.pdf){target="_blank"}

### Fall Run
`r print_model_details('Battle Creek', 'fr')`

### Spring Run and Steelhead 
`r print_model_details('Battle Creek', 'sr')`

## Floodplain Data
The areas represent total inundated area in acres.
```{r}
battle_creek_floodplain <- scale_fp_flow_area('Battle Creek')

knitr::kable(align = 'c', head(battle_creek_floodplain, 5), 
             caption = "Header Descriptions: flow_cfs = flow in cubic feet per second,
             FR_floodplain_acres = Fall Run Chinook floodplain acres, 
             SR_floodplain_acres = Spring Run Chinook floodplain acres, 
             ST_floodplain_acres = Steelhead floodplain acres")
# usethis::use_data(battle_creek_floodplain, overwrite = TRUE)
```

*...with 25 more rows*

## Floodplain Plot
```{r}
battle_creek_floodplain %>% 
  select(flow_cfs, FR_floodplain_acres, ST_floodplain_acres, watershed) %>%
  rename(`Fall Run Chinook` = FR_floodplain_acres, 
         `Steelhead and Spring Run Chinook` = ST_floodplain_acres) %>% 
  gather(Species, acres, -flow_cfs, -watershed) %>% 
  ggplot(aes(flow_cfs, acres, color = Species)) %>% +
  geom_line() +
  theme_minimal() +
  scale_color_brewer(palette = 'Dark2') +
  labs(x = 'Flow (cfs)', y = 'Total Inundated Acres') + 
  theme(legend.justification = c(1,0), legend.position = c(.95,.2))
```
