---
title: "Battle Creek"
author: "[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com)"
date: "July 2020"
output: html_document
---
```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, out.width = '100%')
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
```
## Future Data Improvements
None planned. 

## Instream Spawning and Rearing Habitat
**Data Source:** [Thomas R. Payne & Associates 1995](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/Payne1995_BattleCreekIFIM.pdf){target="_blank"} (pg C-1)

The instream spawning and rearing habitat for Fall Run Chinook Salmon, Spring Run Chinook Salmon, and Steelhead in Battle Creek are based on data from Thomas R. Payne & Associates' instream flow evaluations. The flow evaluations were conducted for the California Department of Fish and Wildlife. They uncovered relationships between flow and both suitable spawning habitat and suitable rearing habitat from seven study reaches in Battle Creek. For both species, the weighted usable area (WUA) data were consolidated from all mainstem reaches. 

### Combine Spawning and Rearing Data
The instream spawning and rearing habitat data described above for Fall Run Chinook Salmon (FR), Spring Run Chinook Salmon (SR), and Steelhead (ST) are combined for use in the Decision Support Model (DSM) in the following format. Units are in square feet per 1000 feet. 
```{r}
battle_ST <- read.csv('data-raw/watershed/battle/data/Weighted_composite_Battle.csv', skip = 1) 

battle_creek_instream <- battle_ST %>% 
  select(flow_cfs = Flow, FR_spawn_wua = 'Spawning.FR',
         FR_fry_wua = 'Fry.Rearing.FR', FR_juv_wua = 'Juv.Rearing.FR',
         ST_adult_wua = 'Adult.Trout.ST', ST_spawn_wua = 'Spawning.ST',
         ST_fry_wua = 'Fry.Rearing.ST', ST_juv_wua = 'Juv.Rearing.ST', 
         SR_spawn_wua = 'Spawning.SR',SR_fry_wua = 'Fry.Rearing.SR', SR_juv_wua = 'Juv.Rearing.SR') %>%
  mutate(watershed = 'Battle Creek')

knitr:: kable(head(battle_creek_instream, 5), align = 'c', 
              caption = "Header Descriptions: flow_cfs = flow in cubic feet per second,
              FR_spawn_wua = Fall Run Chinook Spawning WUA,  FR_fry_wua = 
              Fall Run Chinook Fry WUA, FR_juv_wua = Fall Run Chinook Juvenile WUA, 
              ST_adult_wua = Steelhead Adult WUA, ST_spawn_wua = Steelhead Spawning WUA,
              ST_fry_wua = Steelhead Fry WUA, ST_juv_wua = Steelhead Juvenile WUA,
              SR_spawn_wua = Spring Run Chinook Spawning WUA,  SR_fry_wua = 
              Spring Run Chinook Fry WUA, SR_juv_wua = Spring Run Chinook Juvenile WUA,
              watershed = section of stream modeled for CVPIA SDM") 

# usethis::use_data(battle_creek_instream, overwrite = TRUE)
```

*...with 30 more rows.* 

### Spawning WUA    
The following plot shows the spawning WUA in square feet per thousand feet for Fall Run Chinook Salmon, Steelhead, and Spring Run Chinook Salmon. These area per length rates are multiplied by the total spawning reach length mapped by the Science Integration Team (SIT) to estimate available habitat.    
```{r, echo=FALSE}
battle_creek_instream %>% 
  select(flow_cfs, `Fall Run Chinook` = FR_spawn_wua, `Steelhead` = ST_spawn_wua, `Spring Run Chinook` = SR_spawn_wua) %>% 
  gather(Species, WUA, -flow_cfs) %>% 
  filter(!is.na(WUA)) %>% 
  ggplot(aes(flow_cfs, WUA, color = Species)) +
  geom_line() +
  labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') + 
  theme_minimal() + 
  scale_color_brewer(palette = 'Dark2') + 
  theme(legend.justification = c(1,0), legend.position = c(.9,.7))

```

### Rearing WUA
The following plot shows the weighted usable rearing area in square feet per thousand feet for Fall Run Chinook Salmon and Steelhead fry and juvenile. These rates are multiplied by the total rearing reach length mapped by the SIT to estimate available habitat. 
``` {r}
battle_creek_instream %>% 
  select(SR_juv_wua, SR_fry_wua, ST_juv_wua, ST_fry_wua, flow_cfs, watershed) %>%
  gather(species_lifestage, acres, -flow_cfs, -watershed) %>% 
  mutate(
    Species = case_when( 
      str_detect(species_lifestage, 'SR') ~ 'Spring and Fall Run',
      str_detect(species_lifestage, 'ST') ~ 'Steelhead'),
      lifestage = ifelse(str_detect(species_lifestage, 'juv'), 'Juvenile', 'Fry')) %>% 
  ggplot(aes(flow_cfs, acres, color = Species)) +
  geom_line() +
  theme_minimal() +
  facet_wrap(~lifestage, ncol = 1, scales = 'free_y') +
  scale_color_brewer(palette = 'Dark2') +
  labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') + 
  theme(legend.justification = c(1,0), legend.position = c(.95,.7))
```
```{r, include = FALSE}
source('data-raw/watershed/floodplain_utils.R')
```

## Floodplain Rearing Habitat
**Data Source**: [Scaled from a Deer Creek flow to floodplain area relationship generated with a 1D HEC-RAS hydraulic model](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/DRAFT+Deer+Creek+Hydraulic+Models+Tech+Memo+6-08-07.pdf){target="_blank"}

### Fall Run
`r print_model_details('Battle Creek', 'fr')`

### Spring Run and Steelhead 
`r print_model_details('Battle Creek', 'sr')`

## Floodplain Data
The areas represent total inundated area in acres.
```{r}
battle_creek_floodplain <- scale_fp_flow_area('Battle Creek')

knitr::kable(align = 'c', head(battle_creek_floodplain, 5), 
             caption = "Header Descriptions: flow_cfs = flow in cubic feet per second,
             FR_floodplain_acres = Fall Run Chinook floodplain acres,
             SR_floodplain_acres = Spring Run Chinook floodplain acres,
             ST_floodplain_acres = Steelhead floodplain acres,
             watershed = section of stream modeled for CVPIA SDM") 
# usethis::use_data(battle_creek_floodplain, overwrite = TRUE) 
```

*...with 25 more rows*

## Floodplain Plot
```{r}
battle_creek_floodplain %>% 
  select(flow_cfs, FR_floodplain_acres, ST_floodplain_acres, watershed) %>%
  rename(`Fall Run Chinook` = FR_floodplain_acres, 
         `Steelhead and Spring Run Chinook` = ST_floodplain_acres) %>% 
  gather(Species, acres, -flow_cfs, -watershed) %>% 
  ggplot(aes(flow_cfs, acres, color = Species)) %>% +
  geom_line() +
  theme_minimal() +
  scale_color_brewer(palette = 'Dark2') +
  labs(x = 'Flow (cfs)', y = 'Total Inundated Acres') + 
  theme(legend.justification = c(1,0), legend.position = c(.95,.2))
```
