---
title: "Calaveras River"
author: "[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com)" 
date: "July 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, out.width = '100%')
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
```
## Future Data Improvements
None planned. 

## Instream Spawning and Rearing Habitat
**Data Source:** [FISHBIO Environmental and Thomas R. Payne & Associates 2009](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/FishBio_Payne2009_CalaverasInstreamFlowStudy.pdf){target="_blank"} (pgs. 18-19)

There is no habitat modeling available for Fall Run Chinook Salmon on the Calaveras River. Instream spawning and rearing habitat for Steelhead in the Calaveras River is used as a proxy for Chinook. The data are based on an instream flow study conducted by FISHBIO and Thomas Payne for the Stockton East Water District. They used a Physical Habitat Simulation model on four reaches to find a relationship between stream flow and suitable Steelhead habitat. Graphs showing the original data are on pages 18-19 of the study report linked above. The  Steelhead spawning habitat data from [Mark Gard](mailto:mark_gard@fws.gov) were cross-referenced with the original FISHBIO data and amended to include the original FISHBIO data. The data were digitized from the flow : area curves using this [website](https://automeris.io/WebPlotDigitizer/). Values from the four reaches were first compiled into a spreadsheet. For each reach, the data were then weighted by reach length to find the weighted average of the Steelhead spawning and rearing weighted usable area (WUA) in Calaveras River. This approach was developed in coordination with [Mark Gard](mailto:mark_gard@fws.gov) from the U.S. Fish and Wildlife Service. Cells A16:D26 were used with a reach length of 18,282 ft, A30:D40 with a reach length of 25,438 ft, and A43:D43 with a reach length of 27,867 ft. 

### Combine Spawning and Rearing Data 
The instream spawning and rearing habitat data described above for Steelhead (ST) are combined for use in the Decision Support Model (DSM) in the following format. Units are in square feet per 1000 feet.
```{r, echo =FALSE}
total <- 71632
calaveras_reach_1and2 <- read_excel('data-raw/watershed/calaveras/data/Calaveras_all_reaches.xlsx', range = "A16:D26") %>%
  mutate(ft = 18282)%>%
  rename(flow_cfs = "Flow", ST_fry_wua = 'Fry Rearing', ST_juv_wua = 'Juv Rearing', ST_spawn_wua = 'Spawning') 
calaveras_reach_3 <- read_excel('data-raw/watershed/Calaveras/data/Calaveras_all_reaches.xlsx', range = "A30:D40") %>%
  mutate(ft = 25438)%>%
  rename(flow_cfs = "Flow", ST_fry_wua = 'Fry Rearing', ST_juv_wua = 'Juv Rearing', ST_spawn_wua = 'Spawning') 
calaveras_reach_4 <- read_excel('data-raw/watershed/calaveras/data/Calaveras_all_reaches.xlsx', range = "A43:D53") %>%
  mutate(ft = 27867)%>%
  rename(flow_cfs = "Flow", ST_fry_wua = 'Fry Rearing', ST_juv_wua = 'Juv Rearing', ST_spawn_wua = 'Spawning') 

calaveras_river_instream <- bind_rows(calaveras_reach_1and2, calaveras_reach_3, calaveras_reach_4) %>% 
   gather(lifestage, sq_ft, -flow_cfs, -ft) %>% 
   group_by(lifestage, flow_cfs ) %>% 
   summarise(wua = sum(sq_ft * ft/total, na.rm = TRUE)) %>% 
   spread(lifestage, wua) %>% 
   select(flow_cfs, ST_spawn_wua, ST_fry_wua, ST_juv_wua) %>%
   mutate(watershed = 'Calaveras River')
 
knitr::kable(head(calaveras_river_instream, 5), align = "c", 
             caption = "Header Descriptions: flow_cfs = flow in cubic feet per second,
             ST_spawn_wua = Steelhead spawning WUA, ST_fry_wua = Steelhead Fry WUA, 
             ST_juv_wua = Steelhead Juvenile WUA,
             watershed = section of stream modeled for CVPIA SDM")
 
# usethis::use_data(calaveras_river_instream, overwrite = TRUE)
```

*...with 5 more rows.* 

### Spawning WUA 
The following plot shows the weighted usable spawning area (WUA) in square feet per thousand feet for Steelhead. These area per length rates are multiplied by the total spawning reach length mapped by the Science Integration Team (SIT) to estimate available habitat. 
``` {r,echo=FALSE}
calaveras_river_instream %>% 
    filter(!is.na(ST_spawn_wua)) %>%
    ggplot(aes(x = flow_cfs , y = ST_spawn_wua)) +
    geom_line(color = '#7570b3') +
    labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') +
    theme_minimal()
```

### Rearing WUA
The following plot shows the rearing WUA in square feet per thousand feet for Steelhead fry and juvenile. These rates are multiplied by the total rearing reach length mapped by the SIT to estimate available habitat. 
``` {r,echo=FALSE}
calaveras_river_instream %>% 
  gather(Lifestage, wua, -flow_cfs, -watershed)  %>% 
  filter(!is.na(wua), Lifestage != 'ST_spawn_wua') %>%
  mutate(Lifestage = ifelse(Lifestage == 'ST_fry_wua', 'Fry', 'Juvenile')) %>% 
  ggplot(aes(x = flow_cfs , y = wua, color = Lifestage)) +
  geom_line() +
  labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') +
  theme_minimal() + 
  scale_color_manual(values = c('#d95f02','#7570b3')) +
  theme(legend.justification = c(1,0), legend.position = c(.9,.5))
```
```{r, include = FALSE}
source('data-raw/watershed/floodplain_utils.R')
```

## Floodplain Rearing Habitat 

**Data Source:** [TUFLOW hydraulic model with 1D channel and 2D overbank components](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/Tuolumne_W-AR_21__Study+Report.pdf){target="_blank"}

`r print_model_details('Calaveras River', 'fr')`

## Floodplain Data
The areas represent total inundated area in acres.
```{r, echo=FALSE}
calaveras_river_floodplain <- scale_fp_flow_area('Calaveras River')

knitr::kable(align = 'c', head(calaveras_river_floodplain, 5), 
             caption = "Header Descriptions: flow_cfs = flow in cubic feet per second,
             FR_floodplain_acres = Fall Run Chinook floodplain acres, 
             ST_floodplain_acres = Steelhead floodplain acres,
             watershed = section of stream modeled for CVPIA SDM")
# usethis::use_data(calaveras_river_floodplain, overwrite = TRUE)
```

*...with 28 more rows*

## Floodplain Plot
```{r, echo=FALSE}
calaveras_river_floodplain %>% 
  select(FR_floodplain_acres, flow_cfs, watershed) %>%
  rename(`Fall Run Chinook and Steelhead` = FR_floodplain_acres) %>% 
  gather(Species, acres, -flow_cfs, -watershed) %>% 
  ggplot(aes(flow_cfs, acres, color = Species)) +
  geom_line() +
  theme_minimal() +
  scale_color_brewer(palette = 'Dark2') +
  labs(x = 'Flow (cfs)', y = 'Total Inundated Acres') +
  theme(legend.justification = c(1,0), legend.position = c(.95,.05))
```
