---
title: "Yuba River"
author: "[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com)"
date: "June 11, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
library(scales)
```

## Future Data Improvements
none 

## Instream Spawning and Rearing Habitat

**Data Source:**  [TODO
 <https://flowwest.github.io/cvpiaHabitat/reference/yuba_river_instream.html>]

[TODO modeling description]

[on engle to daguerre segment use 400cfs as the min cfs
then use linear interpolation to fill in gaps]
[TODO check with mark about what to do for flows under 400 cfs]

Instream spawning and rearing habitat for Fall Run Chinook Salmon, Spring Run Chinook Salmon, and Steelhead in the Yuba River is based on data from [data source description]. The [DWR FERC instream spawning and rearing habitat data] was provided by [Mark Gard](mailto:mark_gard@fws.gov) from the U.S. Fish and Wildlife Service in a spreadsheet (see '~/cvpiaHabitat/data-raw/mark_gard_data/IFIMWUA.xlsx').Mark Gard instructed us to use cells A4:H33 and A39:H68 of the 'Yuba' tab within his 'IFIMWUA.xlsx', representing Englebright to Daguerre Segment (12.7 miles) and Daguerre to Feather Segment (11.4 miles) respectively.For adult Steelhead Mark Gard instructed us to use cells A73:B100 of the 'Yuba' tab
within his 'IFIMWUA.xlsx'.

### Combine Spawning and Rearing Data
The instream spawning and rearing habitat data described above for Fall Run Chinook Salmon (FR), Spring Run Chinook Salmon (SR), and Steelhead (ST) is combined for use in the DSM in the following format. Units are in square feet per 1000 feet.
```{r, echo =FALSE}

eng_dag <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', 
                      sheet = 'Yuba', range = 'A4:H33', 
                      col_names = c('flow', 'SR_spawn', 'FR_spawn', 'ST_spawn', 
                                    'FR_fry', 'ST_fry', 'FR_juv', 'ST_juv'))

dag_fea <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', 
                      sheet = 'Yuba', range = 'A39:H68',
                      col_names = c('flow', 'SR_spawn', 'FR_spawn', 'ST_spawn', 
                                    'FR_fry', 'ST_fry', 'FR_juv', 'ST_juv'))

Flow <- sort(unique(c(eng_dag$flow, dag_fea$flow)))

tot_study_len = 12.7 + 11.4

eng_dag_approx <- tibble(
  flow_cfs = Flow,
  SR_spawn_wua = approx(eng_dag$flow, eng_dag$SR_spawn, xout = Flow, rule = 1:2)$y,
  FR_spawn_wua = approx(eng_dag$flow, eng_dag$FR_spawn, xout = Flow, rule = 1:2)$y,
  ST_spawn_wua = approx(eng_dag$flow, eng_dag$ST_spawn, xout = Flow, rule = 1:2)$y,
  FR_fry_wua = approx(eng_dag$flow, eng_dag$FR_fry, xout = Flow, rule = 1:2)$y,
  ST_fry_wua = approx(eng_dag$flow, eng_dag$ST_fry, xout = Flow, rule = 1:2)$y,
  FR_juv_wua = approx(eng_dag$flow, eng_dag$FR_juv, xout = Flow, rule = 1:2)$y,
  ST_juv_wua = approx(eng_dag$flow, eng_dag$ST_juv, xout = Flow, rule = 1:2)$y,
  mi = 12.7
)

# convert from sqft to sqft/1000ft
dag_fea_approx <- tibble(
  flow_cfs = Flow,
  SR_spawn_wua = approx(dag_fea$flow, dag_fea$SR_spawn, xout = Flow, rule = 2)$y,
  FR_spawn_wua = approx(dag_fea$flow, dag_fea$FR_spawn, xout = Flow, rule = 2)$y,
  ST_spawn_wua = approx(dag_fea$flow, dag_fea$ST_spawn, xout = Flow, rule = 2)$y,
  FR_fry_wua = approx(dag_fea$flow, dag_fea$FR_fry, xout = Flow, rule = 2)$y,
  ST_fry_wua = approx(dag_fea$flow, dag_fea$ST_fry, xout = Flow, rule = 2)$y,
  FR_juv_wua = approx(dag_fea$flow, dag_fea$FR_juv, xout = Flow, rule = 2)$y,
  ST_juv_wua = approx(dag_fea$flow, dag_fea$ST_juv, xout = Flow, rule = 2)$y,
  mi = 11.4
)

# combine two reaches
rearing <- eng_dag_approx %>% 
  bind_rows(dag_fea_approx) %>% 
  gather(lifestage, sq_ft, -flow_cfs, -mi) %>% 
  group_by(lifestage, flow_cfs) %>% 
  summarise(wua = sum(sq_ft/mi/5.28 * mi/tot_study_len, na.rm = TRUE)) %>% 
  ungroup() %>% 
  spread(lifestage, wua) %>% 
  select(flow_cfs, FR_spawn_wua, SR_spawn_wua, ST_spawn_wua,
         FR_fry_wua, ST_fry_wua, FR_juv_wua, ST_juv_wua)

```

```{r, echo=FALSE}
rear_len <- cvpiaHabitat::watershed_lengths %>% 
  filter(watershed == 'Yuba River', lifestage == 'rearing') %>% 
  pull(miles)

adult_trout <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', 
                      sheet = 'Yuba', range = 'A73:B100',
                      col_names = c('flow_cfs', 'sq_ft')) %>% 
  mutate(ST_adult_wua = sq_ft/rear_len/5.28) %>% 
  select(-sq_ft)

yuba_river_instream <- rearing %>% 
  full_join(adult_trout) %>% 
  mutate(watershed = 'Yuba River') %>% 
  select(flow_cfs, FR_spawn_wua, FR_fry_wua, FR_juv_wua, SR_spawn_wua,
         ST_spawn_wua, ST_fry_wua, ST_juv_wua, ST_adult_wua, watershed)

knitr::kable(align = 'c', head(yuba_river_instream, 5))

# devtools::use_data(yuba_river_instream, overwrite = TRUE)
```


*...with 45 more rows.*

### Spawning WUA 
The instream spawning habitat weighted usable areas for Fall Run Chinook Salmon and Steelhead in the Yuba River.

#### Fall Run and Spring Run Chinook Salmon Spawning WUA Plot
The following plot shows the weighted usable spawning area in square feet per thousand feet of river for Fall Run Chinook Salmon and Spring Run Chinook Salmon. These area per length rates are multiplied by the total spawning reach length mapped by the SIT. 

``` {r,echo=FALSE}
yuba_river_instream %>% 
    rename(`Fall Run` = FR_spawn_wua,  `Steelhead` = ST_spawn_wua, `Spring Run` = SR_spawn_wua) %>% 
    gather(Species, wua, -flow_cfs, -watershed)  %>% 
    filter(!is.na(wua), Species != 'FR_fry_wua',Species != 'FR_juv_wua',Species != 'ST_adult_wua',Species != 'ST_juv_wua',Species != 'ST_fry_wua', Species != 'Steelhead') %>% 
    ggplot(aes(x = flow_cfs , y = wua, color = Species)) +
    geom_line() +
    labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') +
    theme_minimal() + 
    scale_color_manual(values = c('#d95f02','#7570b3','#56B4E9'))
  

```

#### Steelhead Spawning WUA Plot
The following plot shows the weighted usable spawning area in square feet per thousand feet of river for Steelhead. These area per length rates are multiplied by the total spawning reach length mapped by the SIT. 

``` {r,echo=FALSE}
yuba_river_instream %>% 
    rename(`Fall Run` = FR_spawn_wua,  `Steelhead` = ST_spawn_wua, `Spring Run` = SR_spawn_wua) %>% 
    gather(Species, wua, -flow_cfs, -watershed)  %>% 
    filter(!is.na(wua), Species != 'FR_fry_wua',Species != 'FR_juv_wua',Species != 'ST_adult_wua',Species != 'ST_juv_wua',Species != 'ST_fry_wua',Species != 
             'Fall Run',Species != 'Spring Run') %>% 
    ggplot(aes(x = flow_cfs , y = wua, color = Species)) +
    geom_line() +
    labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') +
    theme_minimal() + 
    scale_color_manual(values = c('#d95f02','#7570b3','#56B4E9'))
  

```

### Rearing WUA
The fry and juvenile instream rearing habitat weighted usable areas for Fall Run Chinook Salmon in the Yuba River. 

#### Fall Run and Spring Run Chinook Salmon Rearing WUA Plot

The following plot shows the weighted usable rearing area in square feet per thousand feet of river for Fall Run Chinook Salmon and Spring Run Salmon fry and juvenile. These rates are multiplied by the total rearing reach length mapped by the SIT. 
``` {r,echo=FALSE}
yuba_river_instream %>% 
  rename(`Fall Run/Spring Run (Fry)` = FR_fry_wua, `Fall Run/Spring Run (Juvenile)` = FR_juv_wua, `Steelhead (Fry)` = ST_fry_wua, `Steelhead (Juvenile)` = ST_juv_wua) %>% 
  gather(Lifestage, wua, -flow_cfs, -watershed)  %>% 
  filter(!is.na(wua), Lifestage != 'FR_spawn_wua', Lifestage != 'ST_spawn_wua', Lifestage != 'ST_adult_wua',Lifestage != 'SR_spawn_wua', Lifestage != 'Steelhead (Fry)', Lifestage != 'Steelhead (Juvenile)') %>%
  #mutate(lifestage = ifelse(lifestage == 'FR_fry_wua', 'fry', 'juvenile')) %>% 
  ggplot(aes(x = flow_cfs , y = wua, color = Lifestage)) +
  geom_line() +
  labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') + 
  theme_minimal() + 
  scale_color_manual(values = c('#d95f02','#7570b3','#56B4E9','#009E73'))
```

#### Steelhead Rearing WUA Plot


The following plot shows the weighted usable rearing area in square feet per thousand feet of river for Steelhead fry and juvenile. These rates are multiplied by the total rearing reach length mapped by the SIT.   
``` {r,echo=FALSE}
yuba_river_instream %>% 
  rename(`Fall Run/Spring Run (Fry)` = FR_fry_wua, `Fall Run/Spring Run (Juvenile)` = FR_juv_wua, `Steelhead (Fry)` = ST_fry_wua, `Steelhead (Juvenile)` = ST_juv_wua) %>% 
  gather(Lifestage, wua, -flow_cfs, -watershed)  %>% 
  filter(!is.na(wua), Lifestage != 'FR_spawn_wua', Lifestage != 'ST_spawn_wua', Lifestage != 'ST_adult_wua',Lifestage != 'SR_spawn_wua', Lifestage != 'Fall Run/Spring Run (Fry)', Lifestage !='Fall Run/Spring Run (Juvenile)') %>%
  #mutate(lifestage = ifelse(lifestage == 'FR_fry_wua', 'fry', 'juvenile')) %>% 
  ggplot(aes(x = flow_cfs , y = wua, color = Lifestage)) +
  geom_line() +
  labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') + 
  theme_minimal() + 
  scale_color_manual(values = c('#d95f02','#7570b3','#56B4E9','#009E73'))
```
