---
title: "Yuba River"
author: "sadie"
date: "December 12, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
library(scales)
```

## Instream Spawning and Rearing Habiat

> **Data Source:**  TODO
> <https://flowwest.github.io/cvpiaHabitat/reference/yuba_river_instream.html>

TODO modeling description

on engle to daguerre segment use 400cfs as the min cfs
then use linear interpolation to fill in gaps

Mark Gard instructed us to use cells A4:H33 and A39:H68 of the 'Yuba' tab
within his 'IFIMWUA.xlsx', representing Englebright to Daguerre Segment (12.7 miles) 
and Daguerre to Feather Segment (11.4 miles) respectively.

units in square feet

TODO check with mark about what to do for flows under 400 cfs
```{r}

eng_dag <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', 
                      sheet = 'Yuba', range = 'A4:H33', 
                      col_names = c('flow', 'SR_spawn', 'FR_spawn', 'ST_spawn', 
                                    'FR_fry', 'ST_fry', 'FR_juv', 'ST_juv'))

dag_fea <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', 
                      sheet = 'Yuba', range = 'A39:H68',
                      col_names = c('flow', 'SR_spawn', 'FR_spawn', 'ST_spawn', 
                                    'FR_fry', 'ST_fry', 'FR_juv', 'ST_juv'))

Flow <- sort(unique(c(eng_dag$flow, dag_fea$flow)))

tot_study_len = 12.7 + 11.4

eng_dag_approx <- tibble(
  flow_cfs = Flow,
  SR_spawn_wua = approx(eng_dag$flow, eng_dag$SR_spawn, xout = Flow, rule = 1:2)$y,
  FR_spawn_wua = approx(eng_dag$flow, eng_dag$FR_spawn, xout = Flow, rule = 1:2)$y,
  ST_spawn_wua = approx(eng_dag$flow, eng_dag$ST_spawn, xout = Flow, rule = 1:2)$y,
  FR_fry_wua = approx(eng_dag$flow, eng_dag$FR_fry, xout = Flow, rule = 1:2)$y,
  ST_fry_wua = approx(eng_dag$flow, eng_dag$ST_fry, xout = Flow, rule = 1:2)$y,
  FR_juv_wua = approx(eng_dag$flow, eng_dag$FR_juv, xout = Flow, rule = 1:2)$y,
  ST_juv_wua = approx(eng_dag$flow, eng_dag$ST_juv, xout = Flow, rule = 1:2)$y,
  mi = 12.7
)

# convert from sqft to sqft/1000ft
dag_fea_approx <- tibble(
  flow_cfs = Flow,
  SR_spawn_wua = approx(dag_fea$flow, dag_fea$SR_spawn, xout = Flow, rule = 2)$y,
  FR_spawn_wua = approx(dag_fea$flow, dag_fea$FR_spawn, xout = Flow, rule = 2)$y,
  ST_spawn_wua = approx(dag_fea$flow, dag_fea$ST_spawn, xout = Flow, rule = 2)$y,
  FR_fry_wua = approx(dag_fea$flow, dag_fea$FR_fry, xout = Flow, rule = 2)$y,
  ST_fry_wua = approx(dag_fea$flow, dag_fea$ST_fry, xout = Flow, rule = 2)$y,
  FR_juv_wua = approx(dag_fea$flow, dag_fea$FR_juv, xout = Flow, rule = 2)$y,
  ST_juv_wua = approx(dag_fea$flow, dag_fea$ST_juv, xout = Flow, rule = 2)$y,
  mi = 11.4
)

# combine two reaches
rearing <- eng_dag_approx %>% 
  bind_rows(dag_fea_approx) %>% 
  gather(lifestage, sq_ft, -flow_cfs, -mi) %>% 
  group_by(lifestage, flow_cfs) %>% 
  summarise(wua = sum(sq_ft/mi/5.28 * mi/tot_study_len, na.rm = TRUE)) %>% 
  ungroup() %>% 
  spread(lifestage, wua) %>% 
  select(flow_cfs, FR_spawn_wua, SR_spawn_wua, ST_spawn_wua,
         FR_fry_wua, ST_fry_wua, FR_juv_wua, ST_juv_wua)

```

## Adult Trout
Mark Gard instructed us to use cells A73:B100 of the 'Yuba' tab
within his 'IFIMWUA.xlsx'.

Units in square feet.

TODO confirm with mark tompkins its okay to call trout steelhead in pkg
```{r}
rear_len <- cvpiaHabitat::watershed_lengths %>% 
  filter(watershed == 'Yuba River', lifestage == 'rearing') %>% 
  pull(miles)

adult_trout <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', 
                      sheet = 'Yuba', range = 'A73:B100',
                      col_names = c('flow_cfs', 'sq_ft')) %>% 
  mutate(ST_adult_wua = sq_ft/rear_len/5.28) %>% 
  select(-sq_ft)

yuba_river_instream <- rearing %>% 
  full_join(adult_trout) %>% 
  mutate(watershed = 'Yuba River') 

yuba_river_instream

# devtools::use_data(yuba_river_instream, overwrite = TRUE)
```
