---
title: "Clear Creek"
author: "[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com)"
date: "July 2020"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, out.width = '100%')
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
library(scales)
```
## Future Data Improvements
None planned.  

## Instream Spawning and Rearing Habiat
**Data Source:** [FWS 2007](https://www.fws.gov/lodi/instream-flow/Documents/Clear%20Creek%20Whiskeytown%20to%20Clear%20Creek%20Br%20spawning%20final%20report.pdf){target="_blank"} (pgs 113-117) and [FWS Rearing 2011](https://www.fws.gov/lodi/instream-flow/Documents/Clear%20Creek%20Whiskeytown%20to%20Clear%20Creek%20Br%20rearing%20final%20report.pdf){target="_blank"} (pgs 144-157) and [FWS Spawning 2011](https://www.fws.gov/lodi/instream-flow/Documents/Clear%20Creek%20Lower%20Spawning%20Final%20Report.pdf){target="_blank"} (pgs 107-109) and [FWS 2013](https://www.fws.gov/lodi/instream-flow/Documents/Clear%20Creek%20Lower%20Rearing%20Final%20Report-1.pdf){target="_blank"}(pgs 150-160)    


Instream spawning and rearing habitat for Fall Run Chinook Salmon, Spring Run Chinook Salmon, and Steelhead in Clear Creek are based on data from four U.S. Fish and Wildlife Service (USFWS) instream flow evaluations. The data were collected between 2007 and 2013 for the Central Valley Project Improvement Act (CVPIA). These evaluations uncovered relationships between flow and suitable spawning and rearing habitat area for "upper" (Whiskeytown Dam to Clear Creek Road) and "lower" (Clear Creek Road to the Sacramento River) Clear Creek. The original data can be found on pages 113 - 117 of FWS 2007 (Upper Clear Creek spawning), pages 144 - 157 of FWS Rearing 2011 (Upper Clear Creek rearing, pages 107 - 109 of FWS Spawning 2011 (Lower Clear Creek spawning), and pages 150 - 160 of FWS 2013 (Lower Clear Creek rearing). [Mark Gard](mailto:mark_gard@fws.gov) from the U.S. Fish and Wildlife Service combined the habitat data from these four studies into a [spreadsheet](https://cvpiahabitat-r-package.s3-us-west-2.amazonaws.com/IFIMWUA.xlsx) for the entire Clear Creek watershed. Mark Gard instructed us to use A3:G26, A30:G53, and A58:G81 of the 'Clear' tab within the spreadsheet. These cells represent the weighted usable area (WUA) in square feet for the lower alluvial (8.81 mi), canyon (7.33 mi), and upper alluvial (2.27 mi) segments. Fall Run are only present in the lower alluvial segment. The Fall Run spawning data are an estimate provided by Clear Creek watershed experts upon receiving feedback that the FWS estimates had been too low.

### Combine Spawning and Rearing Data
The instream spawning and rearing habitat data described above for Fall Run Chinook Salmon (FR), Spring Run Chinook Salmon (SR), and Steelhead (ST) is combined for use in the Decision Support Model (DSM) in the following format. Units are in square feet per 1000 feet.   
```{r, echo=FALSE}
clear_raw_ua <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', 
                           range = "A4:G26", sheet = 'Clear', 
                           col_names = c('flow_cfs', 'SR_spawn_wua', 'ST_spawn_wua', 
                                        'SR_fry_wua', 'ST_fry_wua', 'SR_juv_wua', 
                                          'ST_juv_wua')) %>% 
   mutate(miles = 2.27)
 
clear_raw_c <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', 
                           range = "A31:G53", sheet = 'Clear',
                           col_names = c('flow_cfs', 'SR_spawn_wua', 'ST_spawn_wua', 
                                        'SR_fry_wua', 'ST_fry_wua', 'SR_juv_wua', 
                                         'ST_juv_wua')) %>% 
   mutate(miles = 7.33)
 
clear_raw_la <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', 
                            range = "A59:I81", sheet = 'Clear',
                            col_names = c('flow_cfs', 'FR_spawn_wua', 'ST_spawn_wua', 
                                          'SR_fry_wua', 'ST_fry_wua','FR_fry_wua', 
                                          'SR_juv_wua', 'ST_juv_wua', 'FR_juv_wua')) %>% 
   mutate(miles = 8.81)
 
total <- 8.81 + 7.33 + 2.27
 
fr <- clear_raw_la %>% 
   select(flow_cfs, FR_spawn_wua, FR_fry_wua, FR_juv_wua, miles) %>% 
   gather(lifestage, sq_ft, -flow_cfs, -miles) %>% 
   group_by(lifestage, flow_cfs) %>% 
   summarise(wua = sum(sq_ft/miles/5.28, na.rm = TRUE)) %>% 
   spread(lifestage, wua) %>% 
   select(flow_cfs, FR_spawn_wua, FR_fry_wua, FR_juv_wua)%>%
   mutate(FR_spawn_wua = 13000)
 
st <- bind_rows(clear_raw_la, clear_raw_c, clear_raw_ua) %>% 
   select(flow_cfs, ST_spawn_wua, ST_fry_wua, ST_juv_wua, miles) %>% 
   gather(lifestage, sq_ft, -flow_cfs, -miles) %>% 
   group_by(lifestage, flow_cfs) %>% 
   summarise(wua = sum(sq_ft/miles/5.28 * miles/total, na.rm = TRUE)) %>% 
   spread(lifestage, wua) %>% 
   select(flow_cfs, ST_spawn_wua, ST_fry_wua, ST_juv_wua) 
 
sr_juv <- bind_rows(clear_raw_la, clear_raw_c, clear_raw_ua) %>% 
   select(flow_cfs, SR_fry_wua, SR_juv_wua, miles) %>% 
   gather(lifestage, sq_ft, -flow_cfs, -miles) %>% 
   group_by(lifestage, flow_cfs) %>% 
   summarise(wua = sum(sq_ft/miles/5.28 * miles/total, na.rm = TRUE)) %>% 
   spread(lifestage, wua) %>% 
   select(flow_cfs, SR_fry_wua, SR_juv_wua) 
 
sr_spawn <- bind_rows(clear_raw_la, clear_raw_c, clear_raw_ua) %>% 
   select(flow_cfs, SR_spawn_wua, miles) %>% 
   gather(lifestage, sq_ft, -flow_cfs, -miles) %>%
   group_by(lifestage, flow_cfs) %>%
   summarise(wua = sum(sq_ft/miles/5.28 * miles/(total-8.81), na.rm = TRUE)) %>% 
   spread(lifestage, wua) %>% 
   select(flow_cfs, SR_spawn_wua) 
 
sr <- full_join(sr_spawn, sr_juv)
 
clear_creek_instream <- bind_cols(fr, sr, st) %>% 
   select(-flow_cfs1, -flow_cfs2) %>% 
  mutate(watershed = 'Clear Creek')

knitr::kable(align = 'c', head(clear_creek_instream, 5), 
             caption = "Header Descriptions: flow_cfs = flow in cubic feet per second,
             FR_spawn_wua = Fall Run Chinook Spawning WUA, 
             FR_fry_wua = Fall Run Chinook Fry wua, FR_juv_wua = Fall Run Chinook Juvenile WUA,
             SR_spawn_wua = Spring Run Chinook Spawning WUA, SR_fry_wua = 
             Spring Run Chinook Fry WUA, SR_juv_wua = Spring Run Chinook Juvenile WUA,
             ST_spawn_wua = Steelhead Spawning WUA, ST_fry_wua = Steelhead Fry WUA, 
             ST_juv_wua = Steelhead Juvenile WUA,
             watershed = section of stream modeled for CVPIA SDM") 
# usethis::use_data(clear_creek_instream, overwrite = TRUE)
```

*...with 18 more rows.* 

### Spawning WUA 
The following plot shows the weighted usable spawning area (WUA) in square feet per thousand feet for Fall Run Chinook salmon, Spring Run Chinook Salmon, and Steelhead. These area per length rates are multiplied by the total spawning reach length mapped by the Science Integration Team (SIT) to estimate available habitat. 
```{r, echo=FALSE}

clear_creek_instream %>% 
  select(flow_cfs, `Fall Run Chinook` = FR_spawn_wua, `Steelhead` = ST_spawn_wua,
         `Spring Run Chinook` = SR_spawn_wua) %>% 
  gather(Species, WUA, -flow_cfs) %>% 
  filter(!is.na(WUA)) %>% 
  ggplot(aes(flow_cfs, WUA, color = Species)) +
  geom_line() +
  labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') + 
  theme_minimal() + 
  scale_color_manual(values = c('#d95f02','#7570b3', '#56B4E9')) +
  theme(legend.justification = c(1,0), legend.position = c(.9,.55))

```

### Rearing WUA
The following plot shows the rearing WUA in square feet per thousand feet for Fall Run Chinook Salmon, Spring Run Chinook Salmon, and Steelhead fry and juvenile. The juvenile rearing data are the same for Spring Run and Steelhead. These rates are multiplied by the total rearing reach length mapped by the SIT to estimate available habitat.
```{r, echo=FALSE}
clear_creek_instream %>% 
  select(FR_juv_wua, FR_fry_wua, ST_juv_wua, ST_fry_wua, SR_juv_wua, SR_fry_wua, 
         flow_cfs, watershed) %>%
  gather(species_lifestage, acres, -flow_cfs, -watershed) %>% 
  mutate(
    species = case_when( 
      str_detect(species_lifestage, 'ST') ~ 'Steelhead', 
      str_detect(species_lifestage, 'FR') ~ 'Fall Run Chinook',
      str_detect(species_lifestage, 'SR') ~ 'Spring Run Chinook'),
      lifestage = ifelse(str_detect(species_lifestage, 'juv'), 'Juvenile', 'Fry')) %>% 
  ggplot(aes(flow_cfs, acres, color = species)) +
  geom_line() +
  theme_minimal() +
  facet_wrap(~lifestage, ncol = 1, scales = 'free_y') +
  scale_color_brewer(palette = 'Dark2') +
  labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') + 
  theme(legend.justification = c(1,0), legend.position = c(.95,.05))
```

```{r, include = FALSE}
source('data-raw/watershed/floodplain_utils.R')
```

## Floodplain Rearing Habitat

**Data Source:** [Scaled USFWS / FEMA 1D HEC-RAS hydraulic model](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/CVPIA+Annual+Progress+Report+Fiscal+Year+2017.pdf){target="_blank"}

### Fall Run
`r print_model_details('Clear Creek', 'fr')`

### Spring Run and Steelhead
`r print_model_details('Clear Creek', 'sr')`

## Floodplain Data    
The areas represent total inundated area in acres.    
```{r, echo=FALSE}
clear_creek_floodplain <- scale_fp_flow_area('Clear Creek')

knitr::kable(align = 'c', head(clear_creek_floodplain, 5), 
             caption = "Header Descriptions: flow_cfs = flow in cubic feet per second,
             FR_floodplain_acres = Fall Run Chinook floodplain acres, 
             SR_floodplain_acres = Spring Run Chinook floodplain acres, 
             ST_floodplain_acres = Steelhead floodplain acres,
             watershed = section of stream modeled for CVPIA SDM") 
# usethis::use_data(clear_creek_floodplain, overwrite = TRUE)
```

*...with 27 more rows*

## Floodplain Plot
```{r, echo = FALSE}
clear_creek_floodplain %>% 
  select(flow_cfs, FR_floodplain_acres, ST_floodplain_acres, watershed)%>%
  rename(`Fall Run Chinook` = FR_floodplain_acres, 
         `Steelhead and Spring Run Chinook` = ST_floodplain_acres) %>% 
  gather(Species, acres, -flow_cfs, -watershed) %>% 
  ggplot(aes(flow_cfs, acres, color = Species)) +
  geom_line() +
  theme_minimal() +
  scale_color_brewer(palette = 'Dark2') +
  labs(x = 'Flow (cfs)', y = 'Total Inundated Acres') +
  theme(legend.justification = c(1,0), legend.position = c(.9,.1))
```
