---
title: "Clear Creek"
author: "sadie"
date: "December 12, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
library(scales)
```

## Upcoming:
* Scale spawning wua, examine emails from mark gard and charile chamberline to try scaling exercise of spawning habitat

## Instream Spawning and Rearing Habiat

> **Data Source:** FWS 2007-2013 TODO
> <https://flowwest.github.io/cvpiaHabitat/reference/clear_creek_instream.html>

describe modeling TODO

Mark Gard instructed us to use A3:G26, A30:G53, and A58:G81 of the 'Clear' tab
within his 'IFIMWUA.xlsx'. These cells represent the WUA in sqft for the lower
alluvial (8.81 mi), canyon (7.33 mi), and upper alluvial (2.27 mi) segments.
Fall run are only in the lower alluvial segment.


### WUA raw data
Combined the data from three reaches TODO
where is the spring run spawning in the lower alluvial segment?

```{r, echo=FALSE}
cvpiaHabitat::clear_creek_instream
clear_raw_ua <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', range = "A3:G26", sheet = 'Clear', 
                           col_names = c('Flow', 'SR Spawning', 'ST Spawning', 'SR fry', 'ST fry', 'SR juv', 'ST juv'))
clear_raw_c <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', range = "A30:G53", sheet = 'Clear',
                          col_names = c('Flow', 'SR Spawning', 'ST Spawning', 'SR fry', 'ST fry', 'SR juv', 'ST juv'))
clear_raw_la <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', range = "A58:I81", sheet = 'Clear',
                           col_names = c('Flow', 'FR Spawning', 'ST Spawning', 'SR fry', 'ST fry',
                                         'FR fry', 'SR juv', 'STjuv', 'FR juv'))

clear <- mutate(clear_raw_la, miles = 8.81) %>%
  bind_rows(
    mutate(clear_raw_c, miles = 7.33),
    mutate(clear_raw_ua, miles = 2.27)
  )
total <- 8.81 + 7.33 + 2.27
prop_ua <- 2.27 / total
prop_c <- 7.33 / total
prop_la <- 8.81 / total

clear_wua <- clear %>%
  gather(species_lifestage, sqft, -Flow, -miles) %>%
  mutate(sqft_per_1000ft = sqft/miles/5.28, species_lifestage = paste (species_lifestage, miles)) %>%
  select(Flow, species_lifestage, sqft_per_1000ft) %>%
  spread(species_lifestage, sqft_per_1000ft) %>%
  mutate(SR_juv = prop_ua * `SR juv 2.27` + prop_c * `SR juv 7.33` + prop_la * `SR juv 8.81`,
         ST_juv = prop_ua * `ST juv 2.27` + prop_c * `ST juv 7.33` + prop_la * `ST juv 8.81`,
         SR_fry = prop_ua * `SR fry 2.27` + prop_c * `SR fry 7.33` + prop_la * `SR fry 8.81`,
         ST_fry = prop_ua * `ST fry 2.27` + prop_c * `ST fry 7.33` + prop_la * `ST fry 8.81`,
         SR_spawning = (2.27 / (7.33 + 2.27)) * `SR Spawning 2.27` + (7.33 / (7.33 + 2.27)) * `SR Spawning 7.33`,
         ST_spawning = prop_ua * `ST Spawning 2.27` + prop_c * `ST Spawning 7.33` + prop_la * `ST Spawning 8.81`) %>%
  select(Flow, FR_spawning = `FR Spawning 8.81`, SR_spawning, ST_spawning,
         FR_fry = `FR fry 8.81`, SR_fry, ST_fry, FR_juv = `FR juv 8.81`, SR_juv, ST_juv)

clear_wua
```


### Spawning WUA from raw data

```{r, echo=FALSE}
clear_wua %>%
  select(Flow, FR_spawning:ST_spawning) %>%
  gather(species, wua, -Flow) %>%
  ggplot(aes(Flow, wua, color = species)) +
  geom_line() +
  labs(x = 'flow (cfs)', y = 'spawning WUA (sqft/1000ft)') +
  scale_y_continuous(labels = comma)
```

### Spawning WUA from cvpiaHabitat package

```{r, echo=FALSE}
cvpiaHabitat::clear_creek_instream %>%
  select(flow_cfs, FR_spawn_wua, SR_spawn_wua, ST_spawn_wua) %>%
  gather(species, wua, -flow_cfs) %>%
  ggplot(aes(flow_cfs, wua, color = species)) +
  geom_line() +
  labs(x = 'flow (cfs)', y = 'spawning WUA (sqft/1000ft)') +
  scale_y_continuous(labels = comma)
```


### Spawning Habitat in Acres
Using the habitat package `cvpiaHabitat::set_spawning_habitat`

```{r, echo=FALSE}
spawn_calc <- tibble(
  Flow = seq(25, 1000, length.out = 100),
  FR_acres = square_meters_to_acres(cvpiaHabitat::set_spawning_habitat('Clear Creek', 'fr', Flow)),
  SR_acres = square_meters_to_acres(cvpiaHabitat::set_spawning_habitat('Clear Creek', 'sr', Flow)),
  ST_acres = square_meters_to_acres(cvpiaHabitat::set_spawning_habitat('Clear Creek', 'st', Flow)))

spawn_calc %>%
  gather(species, acres, -Flow) %>%
  ggplot(aes(Flow, acres, color = species)) +
  geom_line()

```

These results are excessively low compared to measured spawning habitat used

### Spawning habitat used
```{r, echo=FALSE}
used <- read_csv('clear_creek_mapped_spawning.csv', skip = 1)

used %>%
  ggplot(aes(x = as.character(Year), y = acres)) +
  geom_col() +
  labs(x = 'year')
```



### Rearing WUA Plot
```{r, echo=FALSE}
clear_wua %>%
  select(-FR_spawning, -SR_spawning, -ST_spawning) %>%
  gather(species, wua, -Flow) %>%
  ggplot(aes(Flow, wua, color = species)) +
  geom_line() +
  labs(x = 'flow (cfs)', y = 'spawning WUA (sqft/1000ft)') +
  scale_y_continuous(labels = comma)
```

Steelhead and spring run juv rearing habitat is the same, why?
ASK MARK GARD- is criteria the same?
small sample size for both steelhead and spring run, criteria is the same so lumped the data together TODO

```{r}
# Old code
# clear creek instream cleaning-----------------------------------
# in units sqft want in units sqft/1000ft
clear_creek <- read_csv('data-raw/instream/clear_creek_instream.csv')
glimpse(clear_creek)
View(clear_creek)

clear_creek_instream <- clear_creek %>%
  gather(species_stage, WUA, -flow_cfs, -Reach:-WUA_units) %>%
  filter(!is.na(WUA)) %>%
  group_by(species_stage, flow_cfs) %>%
  mutate(total_reach_length = sum(Reach_length)) %>%
  summarise(WUA = sum(WUA)/ max(total_reach_length) / 5.28) %>%
  ungroup() %>%
  spread(species_stage, WUA) %>%
  mutate(watershed = 'Clear Creek') %>%
  select(flow_cfs,
         FR_fry_wua = FR_fry,
         FR_juv_wua = FR_juv,
         FR_spawn_wua = FR_spawning,
         SR_fry_wua = SR_fry,
         SR_juv_wua = SR_juvenile,
         SR_spawn_wua = SR_spawning,
         ST_fry_wua = ST_fry,
         ST_juv_wua = ST_juvenile,
         ST_spawn_wua = ST_spawning,
         watershed) # add new naming convention to cols

# devtools::use_data(clear_creek_instream, overwrite = TRUE)
```


