---
title: "Clear Creek"
author: "sadie"
date: "December 12, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
library(scales)
```

## Upcoming:
* Scale spawning wua, examine emails from mark gard and charile chamberline to try scaling exercise of spawning habitat

## Instream Spawning and Rearing Habiat

> **Data Source:** FWS 2007-2013 TODO
> <https://flowwest.github.io/cvpiaHabitat/reference/clear_creek_instream.html>

describe modeling TODO

Mark Gard instructed us to use A3:G26, A30:G53, and A58:G81 of the 'Clear' tab
within his 'IFIMWUA.xlsx'. These cells represent the WUA in sqft for the lower
alluvial (8.81 mi), canyon (7.33 mi), and upper alluvial (2.27 mi) segments.
Fall run are only in the lower alluvial segment.

### WUA raw data
Combined the data from three reaches TODO
where is the spring run spawning in the lower alluvial segment?

```{r, echo=FALSE}
clear_raw_ua <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', 
                           range = "A4:G26", sheet = 'Clear', 
                           col_names = c('flow_cfs', 'SR_spawn_wua', 'ST_spawn_wua', 
                                         'SR_fry_wua', 'ST_fry_wua', 'SR_juv_wua', 
                                         'ST_juv_wua')) %>% 
  mutate(miles = 2.27)

clear_raw_c <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', 
                          range = "A31:G53", sheet = 'Clear',
                          col_names = c('flow_cfs', 'SR_spawn_wua', 'ST_spawn_wua', 
                                        'SR_fry_wua', 'ST_fry_wua', 'SR_juv_wua', 
                                        'ST_juv_wua')) %>% 
  mutate(miles = 7.33)

clear_raw_la <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', 
                           range = "A59:I81", sheet = 'Clear',
                           col_names = c('flow_cfs', 'FR_spawn_wua', 'ST_spawn_wua', 
                                         'SR_fry_wua', 'ST_fry_wua','FR_fry_wua', 
                                         'SR_juv_wua', 'ST_juv_wua', 'FR_juv_wua')) %>% 
  mutate(miles = 8.81)

total <- 8.81 + 7.33 + 2.27

clear_wua <- bind_rows(clear_raw_la, clear_raw_c, clear_raw_ua) %>% 
  gather(lifestage, sq_ft, -flow_cfs, -miles) %>% 
  group_by(lifestage, flow_cfs) %>% 
  summarise(wua = sum(sq_ft/miles/5.28 * miles/total, na.rm = TRUE)) %>% 
  spread(lifestage, wua) %>% 
  select(flow_cfs, FR_spawn_wua, FR_fry_wua, FR_juv_wua, SR_spawn_wua, SR_fry_wua, 
         SR_juv_wua, ST_spawn_wua, ST_fry_wua, ST_juv_wua)

fr <- clear_raw_la %>% 
  select(flow_cfs, FR_spawn_wua, FR_fry_wua, FR_juv_wua, miles) %>% 
  gather(lifestage, sq_ft, -flow_cfs, -miles) %>% 
  group_by(lifestage, flow_cfs) %>% 
  summarise(wua = sum(sq_ft/miles/5.28, na.rm = TRUE)) %>% 
  spread(lifestage, wua) %>% 
  select(flow_cfs, FR_spawn_wua, FR_fry_wua, FR_juv_wua) %>% 
  mutate(FR_spawn_wua = 13000)

st <- bind_rows(clear_raw_la, clear_raw_c, clear_raw_ua) %>% 
  select(flow_cfs, ST_spawn_wua, ST_fry_wua, ST_juv_wua, miles) %>% 
  gather(lifestage, sq_ft, -flow_cfs, -miles) %>% 
  group_by(lifestage, flow_cfs) %>% 
  summarise(wua = sum(sq_ft/miles/5.28 * miles/total, na.rm = TRUE)) %>% 
  spread(lifestage, wua) %>% 
  select(flow_cfs, ST_spawn_wua, ST_fry_wua, ST_juv_wua) 

sr_juv <- bind_rows(clear_raw_la, clear_raw_c, clear_raw_ua) %>% 
  select(flow_cfs, SR_fry_wua, SR_juv_wua, miles) %>% 
  gather(lifestage, sq_ft, -flow_cfs, -miles) %>% 
  group_by(lifestage, flow_cfs) %>% 
  summarise(wua = sum(sq_ft/miles/5.28 * miles/total, na.rm = TRUE)) %>% 
  spread(lifestage, wua) %>% 
  select(flow_cfs, SR_fry_wua, SR_juv_wua) 

sr_spawn <- bind_rows(clear_raw_la, clear_raw_c, clear_raw_ua) %>% 
  select(flow_cfs, SR_spawn_wua, miles) %>% 
  gather(lifestage, sq_ft, -flow_cfs, -miles) %>%
  group_by(lifestage, flow_cfs) %>%
  summarise(wua = sum(sq_ft/miles/5.28 * miles/(total-8.81), na.rm = TRUE)) %>% 
  spread(lifestage, wua) %>% 
  select(flow_cfs, SR_spawn_wua) 

sr <- full_join(sr_spawn, sr_juv)

clear_creek_instream <- bind_cols(fr, sr, st) %>% 
  select(-flow_cfs1, -flow_cfs2)

clear_creek_instream
# devtools::use_data(clear_creek_instream, overwrite = TRUE)
```


### Spawning WUA from raw data

```{r, echo=FALSE}
feet <- cvpiaHabitat::watershed_lengths %>% 
  filter(watershed == 'Clear Creek', lifestage == 'spawning') %>% 
  pull(feet)

names(feet) <- c('FR_spawn_wua', 'SR_spawn_wua', 'ST_spawn_wua')

clear_creek_instream %>%
  select(flow_cfs, FR_spawn_wua, SR_spawn_wua, ST_spawn_wua) %>%
  gather(species, wua, -flow_cfs) %>% 
  mutate(feet = feet[species], acres = wua* feet/1000/43560) %>% 
  ggplot(aes(flow_cfs, acres, color = species)) +
  geom_line() +
  labs(x = 'flow (cfs)', y = 'spawning (acres)') +
  scale_y_continuous(labels = comma)
```



### Spawning habitat used
```{r, echo=FALSE}
used <- read_csv('data-raw/instream/clear/clear_creek_mapped_spawning.csv', skip = 1)

used %>%
  ggplot(aes(x = as.character(Year), y = acres)) +
  geom_col() +
  labs(x = 'year')
```

The spawning wua are excessively low compared to measured spawning habitat used
Will look into scaling modeling, for now fixing spawning habitat value.

### Rearing WUA Plot
```{r, echo=FALSE}
clear_creek_instream %>% 
  select(-FR_spawn_wua, -SR_spawn_wua, -ST_spawn_wua) %>%
  gather(species, wua, -flow_cfs) %>%
  ggplot(aes(flow_cfs, wua, color = species)) +
  geom_line() +
  labs(x = 'flow (cfs)', y = 'rearing WUA (sqft/1000ft)') +
  scale_y_continuous(labels = comma)
```

Steelhead and spring run juv rearing habitat are the same, because small sample size for both steelhead and spring run, criteria is the same so lumped the data together TODO
