---
title: "Clear Creek"
author: "sadie"
date: "December 12, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
library(scales)
```

## Instream Spawning and Rearing Habiat

> **Data Source:** FWS 2007-2013
> <https://flowwest.github.io/cvpiaHabitat/reference/clear_creek_instream.html>

Mark Gard instructed us to use A3:G26, A30:G53, and A58:G81 of the 'Clear' tab
within his 'IFIMWUA.xlsx'. These cells represent the WUA in sqft for the lower
alluvial (8.81 mi), canyon (7.33 mi), and upper alluvial (2.27 mi) segments.
Fall run are only in the lower alluvial segment.



#### WUA data in cvpiaHabitat package compared to mark gard raw data
```{r, echo=FALSE}
cvpiaHabitat::clear_creek_instream
clear_raw_ua <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', range = "A3:G26", sheet = 'Clear', 
                           col_names = c('Flow', 'SR Spawning', 'ST Spawning', 'SR fry', 'ST fry', 'SR juv', 'ST juv'))
clear_raw_c <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', range = "A30:G53", sheet = 'Clear',
                          col_names = c('Flow', 'SR Spawning', 'ST Spawning', 'SR fry', 'ST fry', 'SR juv', 'ST juv'))
clear_raw_la <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', range = "A58:I81", sheet = 'Clear',
                           col_names = c('Flow', 'FR Spawning', 'ST Spawning', 'SR fry', 'ST fry',
                                         'FR fry', 'SR juv', 'STjuv', 'FR juv'))

clear <- mutate(clear_raw_la, miles = 8.81) %>%
  bind_rows(
    mutate(clear_raw_c, miles = 7.33),
    mutate(clear_raw_ua, miles = 2.27)
  )
total <- 8.81 + 7.33 + 2.27
prop_ua <- 2.27 / total
prop_c <- 7.33 / total
prop_la <- 8.81 / total

clear_wua <- clear %>%
  gather(species_lifestage, sqft, -Flow, -miles) %>%
  mutate(sqft_per_1000ft = sqft/miles/5.28, species_lifestage = paste (species_lifestage, miles)) %>%
  select(Flow, species_lifestage, sqft_per_1000ft) %>%
  spread(species_lifestage, sqft_per_1000ft) %>%
  mutate(SR_juv = prop_ua * `SR juv 2.27` + prop_c * `SR juv 7.33` + prop_la * `SR juv 8.81`,
         ST_juv = prop_ua * `ST juv 2.27` + prop_c * `ST juv 7.33` + prop_la * `ST juv 8.81`,
         SR_fry = prop_ua * `SR fry 2.27` + prop_c * `SR fry 7.33` + prop_la * `SR fry 8.81`,
         ST_fry = prop_ua * `ST fry 2.27` + prop_c * `ST fry 7.33` + prop_la * `ST fry 8.81`,
         SR_spawning = (2.27 / (7.33 + 2.27)) * `SR Spawning 2.27` + (7.33 / (7.33 + 2.27)) * `SR Spawning 7.33`,
         ST_spawning = prop_ua * `ST Spawning 2.27` + prop_c * `ST Spawning 7.33` + prop_la * `ST Spawning 8.81`) %>%
  select(Flow, FR_spawning = `FR Spawning 8.81`, SR_spawning, ST_spawning,
         FR_fry = `FR fry 8.81`, SR_fry, ST_fry, FR_juv = `FR juv 8.81`, SR_juv, ST_juv)

clear_wua
```
where is the spring run spawning in the lower alluvial segment?

  ### Spawning WUA from raw data

  ```{r, echo=FALSE}
clear_wua %>%
  select(Flow, FR_spawning:ST_spawning) %>%
  gather(species, wua, -Flow) %>%
  ggplot(aes(Flow, wua, color = species)) +
  geom_line() +
  labs(x = 'flow (cfs)', y = 'spawning WUA (sqft/1000ft)') +
  scale_y_continuous(labels = comma)
```

### Spawning WUA from cvpiaHabitat package

```{r, echo=FALSE}
cvpiaHabitat::clear_creek_instream %>%
  select(flow_cfs, FR_spawn_wua, SR_spawn_wua, ST_spawn_wua) %>%
  gather(species, wua, -flow_cfs) %>%
  ggplot(aes(flow_cfs, wua, color = species)) +
  geom_line() +
  labs(x = 'flow (cfs)', y = 'spawning WUA (sqft/1000ft)') +
  scale_y_continuous(labels = comma)
```


### Spawning Habitat in Acres
Using the habitat package `cvpiaHabitat::set_spawning_habitat`

```{r, echo=FALSE}
spawn_calc <- tibble(
  Flow = seq(25, 1000, length.out = 100),
  FR_acres = square_meters_to_acres(cvpiaHabitat::set_spawning_habitat('Clear Creek', 'fr', Flow)),
  SR_acres = square_meters_to_acres(cvpiaHabitat::set_spawning_habitat('Clear Creek', 'sr', Flow)),
  ST_acres = square_meters_to_acres(cvpiaHabitat::set_spawning_habitat('Clear Creek', 'st', Flow)))

spawn_calc %>%
  gather(species, acres, -Flow) %>%
  ggplot(aes(Flow, acres, color = species)) +
  geom_line()

```

These results are excessively low compared to measured spawning habitat used

### Spawning habitat used
```{r, echo=FALSE}
used <- read_csv('clear_creek_mapped_spawning.csv', skip = 1)

used %>%
  ggplot(aes(x = as.character(Year), y = acres)) +
  geom_col() +
  labs(x = 'year')
```

examine emails from mark gard and charile chamberline to try scaling exercise of spawning habitat


### Rearing WUA from raw data
```{r, echo=FALSE}
clear_wua %>%
  select(-FR_spawning, -SR_spawning, -ST_spawning) %>%
  gather(species, wua, -Flow) %>%
  ggplot(aes(Flow, wua, color = species)) +
  geom_line() +
  labs(x = 'flow (cfs)', y = 'spawning WUA (sqft/1000ft)') +
  scale_y_continuous(labels = comma)
```

Steelhead and spring run juv rearing habitat is the same, why?
  ASK MARK GARD- is criteria the same?
  small sample size for both steelhead and spring run, criteria is the same so lumped the data together


### Rearing Habitat in Acres
Using the habitat package `cvpiaHabitat::set_instream_habitat`

```{r, echo=FALSE}
rear_calc <- tibble(
  Flow = seq(100, 1000, length.out = 100),
  fr_fry = cvpiaHabitat::square_meters_to_acres(
    cvpiaHabitat::set_instream_habitat('Clear Creek', 'fr', 'fry', Flow)),
  fr_juv = cvpiaHabitat::square_meters_to_acres(
    cvpiaHabitat::set_instream_habitat('Clear Creek', 'fr', 'juv', Flow)),
  sr_fry = cvpiaHabitat::square_meters_to_acres(
    cvpiaHabitat::set_instream_habitat('Clear Creek', 'sr', 'fry', Flow)),
  sr_juv = cvpiaHabitat::square_meters_to_acres(
    cvpiaHabitat::set_instream_habitat('Clear Creek', 'sr', 'juv', Flow)),
  st_fry = cvpiaHabitat::square_meters_to_acres(
    cvpiaHabitat::set_instream_habitat('Clear Creek', 'st', 'fry', Flow)),
  st_juv = cvpiaHabitat::square_meters_to_acres(
    cvpiaHabitat::set_instream_habitat('Clear Creek', 'st', 'juv', Flow))) %>%
  gather(species, acres, -Flow)

rear_calc %>%
  ggplot(aes(Flow, acres, color = species)) +
  geom_line()

```

# mark gard scaling

