---
title: "Merced River"
author: "[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com), [Erin Cain](ecain@flowwest.com)"
date: "June 11, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
```
## Future Data Improvements
none

## Instream Spawning and Rearing Habitat

**Data Source:**
[FWS 1996 and MID 2013](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/FWS1996_AmericanRiverSpawningFinalReport.pdf)

Instream spawning and rearing habitat for Fall Run Chinook Salmon and Steelhead in the Merced River is based on data from [data source description]. The [DWR FERC instream spawning and rearing habitat data] was provided by [Mark Gard](mailto:mark_gard@fws.gov) from the U.S. Fish and Wildlife Service in a spreadsheet (see '~/cvpiaHabitat/data-raw/mark_gard_data/IFIMWUA.xlsx'). Mark Gard instructed us to use cells A18:E48 in the 'Merced' tab within his 'IFIMWUA.xlsx'.

### Combine Spawning and Rearing Data
The instream spawning and rearing habitat data described above for Fall Run Chinook Salmon (FR) and Steelhead (ST) is combined for use in the DSM in the following format. Units are in square feet per 1000 feet.


```{r,echo=FALSE}
merced_river_instream <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', 
            range = "A18:E48", sheet = 'Merced') %>% 
   rename(flow_cfs = Flow, FR_spawn_wua = Spawning, FR_fry_wua = `Fry Rearing`, 
          FR_juv_wua = `Juv Rearing`, ST_adult_wua = `adult steelhead`) %>% 
   mutate(watershed = 'Merced River')
 
knitr::kable(align = 'c', head(merced_river_instream, 5))
# devtools::use_data(merced_river_instream, overwrite = TRUE)
```

###### ...with 25 more rows. 
### Spawning WUA 
The instream spawning habitat weighted usable areas for Fall Run Chinook Salmon in the Merced River.

#### Spawning WUA Plot
The following plot shows the weighted usable spawning area in square feet per thousand feet of river for Fall Run Chinook Salmon. These area per length rates are multiplied by the total spawning reach length mapped by the SIT. 

``` {r,echo=FALSE}
merced_river_instream %>% 
    filter(!is.na(FR_spawn_wua)) %>%
    ggplot(aes(x = flow_cfs , y = FR_spawn_wua)) +
    geom_line(color = '#7570b3') +
    labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') + 
    theme_minimal()

```

### Rearing WUA
The fry and juvenile instream rearing habitat weighted usable areas for Fall Run Chinook Salmon in the Merced River. 

#### Rearing WUA Plot

The following plot shows the weighted usable rearing area in square feet per thousand feet of river for Fall Run Chinook Salmon fry and juvenile. These rates are multiplied by the total rearing reach length mapped by the SIT. 

``` {r,echo=FALSE}
merced_river_instream %>% 
  gather(Lifestage, wua, -flow_cfs, -watershed)  %>% 
  filter(!is.na(wua), Lifestage != 'FR_spawn_wua', Lifestage != 'ST_adult_wua') %>%
  mutate(Lifestage = ifelse(Lifestage == 'FR_fry_wua', 'Fry', 'Juvenile')) %>% 
  ggplot(aes(x = flow_cfs , y = wua, color = Lifestage)) +
  geom_line() +
  labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') + 
  theme_minimal() + 
  scale_color_manual(values = c('#d95f02','#7570b3'))
```


