---
title: "Cosumnes River"
author: "[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com), [Erin Cain](ecain@flowwest.com)"
date: "June 18, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
library(scales)

```
## Future Data Improvements
None Planned 

## Instream Spawning and Rearing Habitat

**Data Source:**   
[Calaveras, FISHBIO Environmental and Thomas R. Payne & Associates 2009](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/FishBio_Payne2009_CalaverasInstreamFlowStudy.pdf)    
[Mokelumne, CDFW 1991](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/Lower+Mokelumne+RIver+Fisheries+Management+Plan%2C+CDFG+1991.pdf)

No watershed specific salmonid habitat data was available for Consumnes River. A regional weighted usable area (WUA) relationship with flow was derived for Consumnes River by averaging the WUA values on Calaveras River and Mokelumne River. The geomorphic and hydrologic conditions in Consumnes River (watershed area = X; active channel width = Y; 2-year flow = Z) are similar to those on Calaveras River (watershed area = X; active channel width = Y; 2-year flow = Z), Mokelumne River (watershed area = X; active channel width = Y; 2-year flow = Z). The regional WUA relationships for Calaveras River were multiplied by the length of spawning and rearing extents mapped by the Science Integration Team (SIT). 

### Combine Spawning and Rearing Data 
The instream spawning and rearing habitat data described above for Fall Run Chinook Salmon (FR) is combined for use in the DSM in the following format. Units are in square feet per 1000 feet.

```{r,echo=FALSE}
 #cosumnes -----
 region <- cvpiaHabitat::modeling_exist %>%
   filter(Region == 'South Delta') %>%
   select(Watershed, Region, FR_fry, FR_juv)
 
 flows <- seq(100, 1000, by = 100)
 
 #spawn
 cal_st_spwn <- approx(calaveras_river_instream$flow_cfs, 
                       calaveras_river_instream$ST_spawn_wua, 
                       rule = 2, xout = flows)$y

 mok_fr_spwn <- approx(mokelumne_river_instream$flow_cfs, 
                       mokelumne_river_instream$FR_spawn_wua, 
                       rule = 2, xout = flows)$y
 
 cos_spwn <- (cal_st_spwn + mok_fr_spwn) / 2
 
 # fry
 cal_st_fry <- approx(calaveras_river_instream$flow_cfs, 
                       calaveras_river_instream$ST_fry_wua, 
                       rule = 2, xout = flows)$y
 
 mok_fr_fry <- approx(mokelumne_river_instream$flow_cfs, 
                       mokelumne_river_instream$FR_fry_wua, 
                       rule = 2, xout = flows)$y
 
 cos_fry <- (cal_st_fry + mok_fr_fry) / 2
 
 # juv
 cal_st_juv <- approx(calaveras_river_instream$flow_cfs, 
                       calaveras_river_instream$ST_juv_wua, 
                       rule = 2, xout = flows)$y
 
 mok_fr_juv <- approx(mokelumne_river_instream$flow_cfs, 
                       mokelumne_river_instream$FR_juv_wua, 
                       rule = 2, xout = flows)$y
 
 cos_juv <- (cal_st_juv + mok_fr_juv) / 2
 
 
 
 cosumnes_river_instream <- tibble(flow_cfs = flows, FR_spawn_wua = cos_spwn, 
                                   FR_fry_wua = cos_fry, FR_juv_wua = cos_juv,
                                   watershed = 'Cosumnes River')
 
knitr::kable(align = 'c', head(cosumnes_river_instream, 5))
# devtools::use_data(cosumnes_river_instream, overwrite = TRUE)
```

*...with 5 more rows.* 

### Spawning WUA 
The instream spawning habitat weighted usable areas for Fall Run Chinook Salmon and Steelhead in the Cosumnes River.

#### Spawning WUA Plot
The following plot shows the weighted usable spawning area in square feet per thousand feet of river for Fall Run Chinook Salmon. These area per length rates are multiplied by the total spawning reach length mapped by the SIT. 

``` {r,echo=FALSE}
cosumnes_river_instream %>% 
    filter(!is.na(FR_spawn_wua)) %>%
    ggplot(aes(x = flow_cfs , y = FR_spawn_wua)) +
    geom_line(color= '#7570b3') +
    labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') + 
    theme_minimal()
  
```

### Rearing WUA
The fry and juvenile instream rearing habitat weighted usable areas for Fall Run Chinook Salmon in the Consumnes River.

#### Rearing WUA Plot

The following plot shows the weighted usable rearing area in square feet per thousand feet of river for Fall Run Chinook Salmon fry and juvenile. These rates are multiplied by the total rearing reach length mapped by the SIT. 
``` {r,echo=FALSE}
cosumnes_river_instream %>% 
  gather(Lifestage, wua, -flow_cfs, -watershed)  %>% 
  filter(!is.na(wua), Lifestage != 'FR_spawn_wua') %>%
  mutate(Lifestage = ifelse(Lifestage == 'FR_fry_wua', 'Fry', 'Juvenile')) %>% 
  ggplot(aes(x = flow_cfs , y = wua, color = Lifestage)) +
  geom_line() +
  labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') + 
  theme_minimal() + 
  scale_color_manual(values = c('#d95f02','#7570b3'))
```

```{r, include = FALSE}
source('data-raw/floodplain/floodplain_utils.R')
```

## Floodplain Modeling Details

**Data Source:**
[TUFLOW hydraulic model with 1D channel and 2D overbank components](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/Tuolumne_W-AR_21__Study+Report.pdf)

`r print_model_details('Cosumnes River', 'fr')`

## Floodplain Data

```{r, echo=FALSE}
cosumnes_river_floodplain <- scale_fp_flow_area('Cosumnes River')

knitr::kable(align = 'c', head(cosumnes_river_floodplain, 5))
# use_data(cosumnes_river_floodplain, overwrite = TRUE)
```

*...with 28 more rows*

## Floodplain Plot
```{r, echo = FALSE}
cosumnes_river_floodplain %>% 
  select(FR_floodplain_acres, flow_cfs, watershed) %>%
  rename(`Fall Run Chinook and Steelhead` = FR_floodplain_acres) %>% 
  gather(Species, acres, -flow_cfs, -watershed) %>% 
  ggplot(aes(flow_cfs, acres, color = Species)) +
  geom_line() +
  theme_minimal() +
  scale_color_brewer(palette = 'Dark2') +
  labs(x = 'Flow (cfs)', y = 'Total Inundated Acres')
```

