---
title: "Yuba River"
author: "[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com), [Erin Cain](mailto:ecain@flowwest.com)"
date: "August 19, 2019"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, out.width = '100%')
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
library(scales)
```

## Future Data Improvements
Update with new VA data 

## Instream Spawning and Rearing Habitat
**Data Source:** [Spawning FWS 2010](https://www.fws.gov/lodi/instream-flow/Documents/Yuba%20River%20Spawning%20Final%20Report.pdf){target="_blank"} (pgs 59-61) and  [Rearing FWS 2010](https://www.fws.gov/lodi/instream-flow/Documents/Yuba%20River%20Rearing%20Final%20Report.pdf){target="_blank"} (pgs 74-76)

Instream spawning and rearing habitat for Fall Run Chinook Salmon, Spring Run Chinook Salmon, and Steelhead in the Yuba River are based on data from the Fish and Wildlife Service flow-habitat relationship final reports for spawning and rearing. River2D, a 2-dimensional hydraulic and habitat model, was used to model available spawning and rearing habitat. The FWS instream spawning and rearing data was provided by [Mark Gard](mailto:mark_gard@fws.gov) from the U.S. Fish and Wildlife Service in a spreadsheet (see '~/cvpiaHabitat/data-raw/mark_gard_data/IFIMWUA.xlsx'). Mark Gard instructed us to use cells A4:H33 and A39:H68 of the 'Yuba' tab within his 'IFIMWUA.xlsx', representing Englebright to Daguerre Segment (12.7 miles) and Daguerre to Feather Segment (11.4 miles) respectively. For adult Steelhead Mark Gard instructed us to use cells A73:B100 of the 'Yuba' tab
within his 'IFIMWUA.xlsx'.

### Combine Spawning and Rearing Data
The instream spawning and rearing habitat data described above for Fall Run Chinook Salmon (FR), Spring Run Chinook Salmon (SR), and Steelhead (ST) is combined for use in the DSM in the following format. Units are in square feet per 1000 feet.
```{r, echo =FALSE}

eng_dag <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', 
                      sheet = 'Yuba', range = 'A4:H33', 
                      col_names = c('flow', 'SR_spawn', 'FR_spawn', 'ST_spawn', 
                                    'FR_fry', 'ST_fry', 'FR_juv', 'ST_juv'))

dag_fea <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', 
                      sheet = 'Yuba', range = 'A39:H68',
                      col_names = c('flow', 'SR_spawn', 'FR_spawn', 'ST_spawn', 
                                    'FR_fry', 'ST_fry', 'FR_juv', 'ST_juv'))

Flow <- sort(unique(c(eng_dag$flow, dag_fea$flow)))

tot_study_len = 12.7 + 11.4

eng_dag_approx <- tibble(
  flow_cfs = Flow,
  SR_spawn_wua = approx(eng_dag$flow, eng_dag$SR_spawn, xout = Flow, rule = 1:2)$y,
  FR_spawn_wua = approx(eng_dag$flow, eng_dag$FR_spawn, xout = Flow, rule = 1:2)$y,
  ST_spawn_wua = approx(eng_dag$flow, eng_dag$ST_spawn, xout = Flow, rule = 1:2)$y,
  FR_fry_wua = approx(eng_dag$flow, eng_dag$FR_fry, xout = Flow, rule = 1:2)$y,
  ST_fry_wua = approx(eng_dag$flow, eng_dag$ST_fry, xout = Flow, rule = 1:2)$y,
  FR_juv_wua = approx(eng_dag$flow, eng_dag$FR_juv, xout = Flow, rule = 1:2)$y,
  ST_juv_wua = approx(eng_dag$flow, eng_dag$ST_juv, xout = Flow, rule = 1:2)$y,
  mi = 12.7
)

# convert from sqft to sqft/1000ft
dag_fea_approx <- tibble(
  flow_cfs = Flow,
  SR_spawn_wua = approx(dag_fea$flow, dag_fea$SR_spawn, xout = Flow, rule = 2)$y,
  FR_spawn_wua = approx(dag_fea$flow, dag_fea$FR_spawn, xout = Flow, rule = 2)$y,
  ST_spawn_wua = approx(dag_fea$flow, dag_fea$ST_spawn, xout = Flow, rule = 2)$y,
  FR_fry_wua = approx(dag_fea$flow, dag_fea$FR_fry, xout = Flow, rule = 2)$y,
  ST_fry_wua = approx(dag_fea$flow, dag_fea$ST_fry, xout = Flow, rule = 2)$y,
  FR_juv_wua = approx(dag_fea$flow, dag_fea$FR_juv, xout = Flow, rule = 2)$y,
  ST_juv_wua = approx(dag_fea$flow, dag_fea$ST_juv, xout = Flow, rule = 2)$y,
  mi = 11.4
)

# combine two reaches
rearing <- eng_dag_approx %>% 
  bind_rows(dag_fea_approx) %>% 
  gather(lifestage, sq_ft, -flow_cfs, -mi) %>% 
  group_by(lifestage, flow_cfs) %>% 
  summarise(wua = sum(sq_ft/mi/5.28 * mi/tot_study_len, na.rm = TRUE)) %>% 
  ungroup() %>% 
  spread(lifestage, wua) %>% 
  select(flow_cfs, FR_spawn_wua, SR_spawn_wua, ST_spawn_wua,
         FR_fry_wua, ST_fry_wua, FR_juv_wua, ST_juv_wua)

```

```{r, echo=FALSE}
rear_len <- cvpiaHabitat::watershed_lengths %>% 
  filter(watershed == 'Yuba River', lifestage == 'rearing') %>% 
  pull(miles)

adult_trout <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', 
                      sheet = 'Yuba', range = 'A73:B100',
                      col_names = c('flow_cfs', 'sq_ft')) %>% 
  mutate(ST_adult_wua = sq_ft/rear_len/5.28) %>% 
  select(-sq_ft)

yuba_river_instream <- rearing %>% 
  mutate(watershed = 'Yuba River') %>% 
  select(flow_cfs, FR_spawn_wua, FR_fry_wua, FR_juv_wua, SR_spawn_wua,
         ST_spawn_wua, ST_fry_wua, ST_juv_wua, watershed)

knitr::kable(align = 'c', head(yuba_river_instream, 5), 
             caption = "Header Descriptions: flow_cfs = flow in cubic feet per second,
             FR_spawn_wua = Fall Run Chinook Spawning WUA, FR_fry_wua = Fall Run Chinook Fry wua,
             FR_juv_wua = Fall Run Chinook Juvenile wua, SR_spawn_wua = 
             Spring Run Chinook Spawning WUA, ST_spawn_wua = Steelhead Spawning WUA, 
             ST_fry_wua = Steelhead Fry wua, ST_juv_wua = Steelhead Juvenile wua")

# usethis::use_data(yuba_river_instream, overwrite = TRUE)
```

*...with 29 more rows.* 

### Spawning WUA    
The following plot shows the weighted usable spawning area in square feet per thousand feet of river for Fall Run Chinook Salmon, Spring Run Chinook Salmon, and Steelhead. These area per length rates are multiplied by the total spawning reach length mapped by the SIT. 
``` {r,echo=FALSE}
yuba_river_instream %>% 
    rename(`Fall Run` = FR_spawn_wua,  `Steelhead` = ST_spawn_wua, 
           `Spring Run` = SR_spawn_wua) %>% 
    gather(Species, wua, -flow_cfs, -watershed)  %>% 
    filter(!is.na(wua), Species != 'FR_fry_wua', Species != 'FR_juv_wua',
           Species != 'ST_adult_wua',Species != 'ST_juv_wua',
           Species != 'ST_fry_wua') %>% 
    ggplot(aes(x = flow_cfs , y = wua, color = Species)) +
    geom_line() +
    labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') +
    theme_minimal() + 
    scale_color_manual(values = c('#d95f02','#7570b3','#56B4E9')) +
  theme(legend.justification = c(1,0), legend.position = c(.95,.7))
```

### Rearing WUA   
The following plot shows the weighted usable rearing area in square feet per thousand feet of river for Fall Run Chinook Salmon and Steelhead fry and juvenile. These rates are multiplied by the total rearing reach length mapped by the SIT. 
``` {r,echo=FALSE}
yuba_river_instream %>% 
  filter(flow_cfs < 5000) %>% 
  select(FR_juv_wua, FR_fry_wua, ST_juv_wua, ST_fry_wua, flow_cfs, watershed) %>%
  gather(species_lifestage, acres, -flow_cfs, -watershed) %>% 
  mutate(
    species = case_when( 
      str_detect(species_lifestage, 'ST') ~ 'Steelhead', 
      str_detect(species_lifestage, 'FR') ~ 'Fall Run'),
      lifestage = ifelse(str_detect(species_lifestage, 'juv'), 'Juvenile', 'Fry')) %>% 
  ggplot(aes(flow_cfs, acres, color = species)) +
  geom_line() +
  theme_minimal() +
  facet_wrap(~lifestage, ncol = 1, scales = 'free_y') +
  scale_color_brewer(palette = 'Dark2') +
  labs(x = 'Flow (cfs)', y = 'Suitable Acres') + 
  theme(legend.justification = c(1,0), legend.position = c(.95,.45))
```

```{r, include=FALSE}
source('data-raw/watershed/floodplain_utils.R')
```

## Floodplain Rearing Habitat 
**Data Source:** [SRH-2D hydraulic model developed by Dr. Greg Pasternack at the University of California, Davis](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/CVPIA_Annual_Progress_Report_Fiscal_Year_2014_Yuba_SRH2D_Page18.pdf){target="_blank"}

`r print_model_details('Yuba River', 'fr')`

## Floodplain Data
The areas represent total inundated area in acres.    
```{r, echo= FALSE}
yuba_fp <- read_excel('data-raw/watershed/CVPIA_FloodplainAreas.xlsx',
                                    sheet = 'YubaRiver') %>%
  mutate(watershed = 'Yuba River',
         SR_floodplain_acres = modeled_floodplain_area_acres,
         ST_floodplain_acres = modeled_floodplain_area_acres) %>%
  select(flow_cfs,
         FR_floodplain_acres = modeled_floodplain_area_acres,
         SR_floodplain_acres, ST_floodplain_acres,
         watershed)

threshold <- yuba_fp %>% 
  filter(FR_floodplain_acres == 0) %>% 
  summarise(max = max(flow_cfs)) %>% 
  pull(max)

yuba_river_floodplain <- yuba_fp %>% 
  filter(flow_cfs >= threshold)

knitr::kable(align = 'c', head(yuba_river_floodplain, 5), 
             caption = "Header Descriptions: flow_cfs = flow in cubic feet per second,
             FR_floodplain_acres = Fall Run Chinook floodplain acres, 
             SR_floodplain_acres = Spring Run Chinook floodplain acres, 
             ST_floodplain_acres = Steelhead floodplain acres")
# usethis::use_data(yuba_river_floodplain, overwrite = TRUE)
```

*...with 15 more rows*

## Floodplain Plot
```{r, echo = FALSE}
yuba_river_floodplain %>% 
  select(FR_floodplain_acres, watershed, flow_cfs) %>%
  rename(`Fall Run Chinook, 
Spring Run Chinook,
and Steelhead` = FR_floodplain_acres) %>% 
  gather(Species, acres, -flow_cfs, -watershed) %>% 
  ggplot(aes(flow_cfs, acres, color = Species)) +
  geom_line() +
  theme_minimal() +
  scale_color_brewer(palette = 'Dark2') +
  labs(x = 'Flow (cfs)', y = 'Total Inundated Acres') +
  theme(legend.justification = c(1,0), legend.position = c(.95,.1))
```

