---
title: "Bear River"
author: "[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com)"
date: "December 13, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
```
## Future Data Improvements
[Future data improvement bullets]

## Instream Spawning and Rearing Habiat

**Data Sources:**
[add more sources?] 
[Holton 1985](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/Holton1985_BearRiverIFIM.PDF) and 
<https://flowwest.github.io/cvpiaHabitat/reference/bear_river_instream.html>

Instream spawning and rearing habitat for Fall Run Chinook Salmon and Steelhead in the Bear River is based on data from California Department of Water Resources and Thomas R. Payne & Associates instream flow evaluations for the relicensing of Oroville facilities. [These evaluations developed relationships between flow and suitable spawning and rearing habitat for 23.25 miles of the Feather River between the Fish Barrier Dam and Honcut Creek, consisting of two river segments. The DWR FERC instream spawning and rearing habitat data] was provided by [Mark Gard](mailto:mark_gard@fws.gov) from the U.S. Fish and Wildlife Service in a spreadsheet (see '~/cvpiaHabitat/data-raw/mark_gard_data/IFIMWUA.xlsx').Mark Gard instructed us to use A1:D5 of the 'BearR' tab within his 'IFIMWUA.xlsx'.

Modeling for spawning and juvenile rearing.

Units are in square feet per 1000 feet.

### Combine Spawning and Rearing Data
[The instream spawning and rearing habitat data described above for Fall Run Chinook Salmon (FR) and Steelhead (ST) is combined for use in the DSM in the following format.]

```{r,echo=FALSE}
bear <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', 
                           range = "A1:D5", sheet = 'BearR')

bear_river_instream <- bear %>% 
  select(flow_cfs = Flow, FR_spawn_wua = Spawning, FR_juv_wua = `Juv Rearing`) %>% 
  mutate(watershed = 'Bear River')

bear_river_instream
# devtools::use_data(bear_river_instream, overwrite = TRUE)
```

