---
title: "Butte Creek"
author: "[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com)"
date: "July 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, out.width = '100%')
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
```
## Future Data Improvements
None planned. 

## Instream Spawning and Rearing Habiat

**Data Source:** [FWS 2003](https://www.fws.gov/lodi/instream-flow/Documents/Butte%20Creek%20Spring-run%20chinook%20salmon%20spawning%208-29-2003.pdf){target="_blank"} (77-78) and [FERC relicensing of DeSabla](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/DeSabla2008ButteIFIM.pdf){target="_blank"} (APPENDIX E6.3.2.6-J1) 

Instream spawning and rearing habitat in Butte Creek is based on data from the FWS instream 2003 study and the FERC relicensing of DeSabla 2008 study. Spawning habitat data are from the FWS study (p. 77-78). For Spring Run only, spawning data were generated using a RIVER2D 2-dimensional hydraulic and habitat model. The spawning data were compiled in a spreadsheet and scaled by reach length. Spring Run Chinook Salmon Spawning data are used as a proxy for Fall Run Chinook Salmon Spawning. The rearing data come from DeSabla 2008 in the lower butte section of the appendix (APPENDIX E6.3.2.6-J1). Instream rearing habitat data were provided in a [spreadsheet](https://cvpiahabitat-r-package.s3-us-west-2.amazonaws.com/IFIMWUA.xlsx). Mark Gard instructed us to use cells A71:C101 for the Fall Run Chinook rearing data.  Units are in square feet per 1000 feet.

```{r, echo=FALSE}
# Spring run spawning data 
butte_raw <- read.csv('data-raw/watershed/butte/data/butte_creek_data.csv', 
                     skip = 1) 

# Combine two reaches of spawning data with different flow values using aproxfun   
flow <- butte_raw %>% 
  pull(flow_cfs) %>%
  unique() %>%
  sort()

make_approx_fun <- function(df) {
  flow <- df$flow_cfs
  return(approxfun(flow, df$SR_spawn, rule = 2))
}
reach_miles <- c(6.5, 9)
names(reach_miles) <- 1:2
total <- sum(reach_miles)

Reach_1 <- butte_raw %>%
  filter(length == 6.5) 

Reach_2 <- butte_raw %>%
  filter(length == 9)

# Steelhead spawning Data 
butte_creek_st_spawn <-  read_excel('data-raw/watershed/butte/data/butte_creek_ST_spawn_data.xlsx', 
                   range = "A2:B32") %>% 
  select(flow_cfs,ST_spawn_wua)


butte_creek_spawning <- tibble(
  flow_cfs = flow,
  SR_spawn_wua =
    make_approx_fun(Reach_1)(flow) * reach_miles[1]/total +
    make_approx_fun(Reach_1)(flow) * reach_miles[2]/total
  )

# Rearing Data 
butte_creek_rearing <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', 
                   range = "A71:C101", sheet = 'Butte') %>% 
  rename(flow_cfs = Flow)
  
# Combine spawning and rearing data 
butte_creek_instream <- butte_creek_spawning %>% 
  full_join(butte_creek_rearing) %>%
  full_join(butte_creek_st_spawn) %>%
  mutate(FR_spawn_wua = SR_spawn_wua, watershed = 'Butte Creek') %>% 
  select(flow_cfs, FR_spawn_wua, SR_spawn_wua, ST_spawn_wua, FR_fry_wua = `Fry Rearing`, 
         FR_juv_wua = `Juv Rearing`, 
         watershed)

#usethis::use_data(butte_creek_instream, overwrite = TRUE)
```

### Spawning Data
```{r, echo = FALSE}
knitr::kable(align = 'c', head(butte_creek_instream %>% select(flow_cfs,ST_spawn_wua,SR_spawn_wua) %>% 
                                 mutate(watershed = 'Butte Creek'), 5), 
             caption = "Header Descriptions: flow_cfs = flow in cubic feet per second, ST_spawn_wua = Steelhead Spawning WUA,
             SR_spawn_wua = Spring Run Chinook Spawning WUA, watershed = section of stream modeled for CVPIA SDM")
```

*...with 36 more rows* 

### Rearing Data 
```{r, echo = FALSE}
knitr::kable(align = 'c', head(butte_creek_rearing %>% 
                                 mutate(watershed = 'Butte Creek') %>% 
                                  rename(SR_fry_wua = 'Fry Rearing', SR_juv_wua = 
                                          'Juv Rearing'), 5), caption = 
                                  "Header Descriptions: flow_cfs = flow in cubic feet per second, 
                                  SR_juv_wua = Spring Run Chinook Juvenile WUA,  
                                  SR_fry_wua = Spring Run Chinook Fry WUA, 
                                  watershed = section of stream modeled for CVPIA SDM")
```

*...with 25 more rows*

### Spawning WUA 
The following plot shows the weighted usable spawning area (WUA) in square feet per thousand feet for Spring Run Chinook Salmon. These area per length rates are multiplied by the total spawning reach length mapped by the Science Integration Team (SIT) to estimate available habitat. 
``` {r,echo=FALSE,warning = FALSE}
butte_creek_instream %>% 
    filter(!is.na(FR_spawn_wua)) %>%
    ggplot(aes(x = flow_cfs)) +
    geom_line(aes(y = FR_spawn_wua,color = '#d95f02')) +
    geom_line(aes(y = ST_spawn_wua,color = '#7570b3')) +
    labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)',color = 'Species') +
    theme_minimal() +
    scale_color_manual(labels = c('Fall and Spring Run','Steelhead'),values = c('#d95f02','#7570b3')) +
    theme(legend.justification = c(1,0), legend.position = c(.9,.75))
   
```

### Rearing WUA
The following plot shows the rearing WUA in square feet per thousand feet for Spring Run Chinook Salmon fry and juvenile. These rates are multiplied by the total rearing reach length mapped by the SIT to estimate available habitat. 

``` {r,echo=FALSE,warning = FALSE}
butte_creek_rearing %>% 
  ggplot(aes(x = flow_cfs)) +
  geom_line(aes(y = `Fry Rearing`,color = '#d95f02',Lifestage = 'Fry Rearing')) +
  geom_line(aes(y = `Juv Rearing`,color = '#7570b3',Lifestage = 'Juv Rearing')) + 
  labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)',color = 'Lifestage') +
  theme_minimal() + 
  scale_color_manual(labels = c('Fry Rearing','Juv Rearing'),values = c('#d95f02','#7570b3')) +
  theme(legend.justification = c(1,0), legend.position = c(.9,.75))
```

```{r, include =FALSE}
source('data-raw/watershed/floodplain_utils.R')
```

## Floodplain Rearing Habitat
**Data Source:** [Central Valley Floodplain Evaluation and Delineation (CVFED) HEC-RAS hydraulic model](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/CombinedTM_IQAR_Final-FULL-REPORT_20140206.pdf){target="_blank"}

### Fall Run
`r print_model_details('Butte Creek', 'fr')`

### Spring Run and Steelhead
`r print_model_details('Butte Creek', 'sr')`

## Floodplain Data    
The areas represent total inundated area in acres.    
```{r, echo=FALSE}
butte_fp <- read_excel('data-raw/watershed/CVPIA_FloodplainAreas.xlsx', sheet = 'ButteCreek') %>%
  mutate(watershed = 'Butte Creek')

threshold_flow <- butte_fp %>% 
  filter(modeled_floodplain_area_acres == 0) %>% 
  summarise(max_flow = max(flow_cfs)) %>% 
  pull(max_flow)

butte_creek_floodplain <- scale_fp_flow_area_partial_model(ws = 'Butte Creek', 
                                                           df = filter(butte_fp, flow_cfs >= threshold_flow))

knitr::kable(align = 'c', head(butte_creek_floodplain, 5), 
             caption = "Header Descriptions: flow_cfs = flow in cubic feet per second,
             FR_floodplain_acres = Fall Run Chinook floodplain acres, 
             SR_floodplain_acres = Spring Run Chinook floodplain acres, 
             ST_floodplain_acres = Steelhead floodplain acres,
             watershed = section of stream modeled for CVPIA SDM") 
# usethis::use_data(butte_creek_floodplain, overwrite = TRUE)
```

*...with 26 more rows*

## Floodplain Plot
```{r, echo = FALSE}
butte_creek_floodplain %>% 
  select(FR_floodplain_acres, SR_floodplain_acres, watershed, flow_cfs) %>%
  rename(`Fall Run Chinook` = FR_floodplain_acres,
         `Spring Run Chinook and Steelhead` = SR_floodplain_acres) %>% 
  gather(Species, acres, -flow_cfs, -watershed) %>% 
  ggplot(aes(flow_cfs, acres, color = Species)) +
  geom_line() +
  theme_minimal() +
  scale_color_brewer(palette = 'Dark2') +
  labs(x = 'Flow (cfs)', y = 'Total Inundated Acres') +
  theme(legend.justification = c(1,0), legend.position = c(1,.1))

```
