---
title: "Deltas"
author: "[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com)"
date: "July 2, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, out.width = '100%')
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(scales)
library(lubridate)
```

## Modeling Details
**Data Source:**
Central Valley Floodplain Evaluation and Delineation (CVFED) HEC-RAS hydraulic model refined for use in the [NOAA-NMFS Winter Run Chinook Salmon life cycle model](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/HendrixEtAl2014_Winter_Run_Model_Tech_Memo.pdf)

The National Oceanographic and Atmospheric Administration (NOAA) NMFS model outputs habitat area estimates for the months of December-May. Habitat area estimates for the months of June-Novemeber are represented by the yearly average habitat. Units are in square meters.

High quality habitat defined by:

* Channel depth > 0.2 m and < 1.5 m
* Velocity <= 0.15 m/s
* Cover Vegetated edge

## North and South Deltas Rearing Habitat

```{r}
south_yolo <- read_excel('data-raw/correigh_greene_data/delta habitat By Region.xlsx', sheet = 3, skip = 1) %>%
  select(X__1, south_yolo = ppp)

north_delta <- read_excel('data-raw/correigh_greene_data/delta habitat By Region.xlsx', sheet = 4, skip = 1)  %>%
  select(X__1, north_delta = ppp)

south_delta <- read_excel('data-raw/correigh_greene_data/delta habitat By Region.xlsx', sheet = 5, skip = 1) %>%
  select(X__1, south_delta = ppp)

north_delta_habitat <- north_delta %>%
  left_join(south_yolo) %>%
  separate(col = X__1, into = c('year', 'month'), sep = ' ') %>%
  mutate(`North Delta` = north_delta + south_yolo,
         date = mdy(paste0(month, ' 1, ', year))) %>%
  select(date, `North Delta`)

nd_year_means <- north_delta_habitat %>%
  group_by(year = year(date)) %>%
  summarise(mean_area = mean(`North Delta`))

nd_means <- tibble(
  date = seq(as.Date('1980-06-01'), as.Date('2010-11-01'), by = 'month')) %>%
  mutate(year = year(date)) %>%
  filter(between(month(date), 6, 11)) %>%
  left_join(nd_year_means) %>%
  select(date, `North Delta` = mean_area)

nd_hab <- north_delta_habitat %>%
  bind_rows(nd_means) %>%
  arrange(date)

south_delta_habitat <- south_delta %>%
  separate(col = X__1, into = c('year', 'month'), sep = ' ') %>%
  mutate(date = mdy(paste0(month, ' 1, ', year))) %>%
  select(date, `South Delta` = south_delta)

sd_year_means <- south_delta_habitat %>%
  group_by(year = year(date)) %>%
  summarise(mean_area = mean(`South Delta`))

sd_means <- tibble(
  date = seq(as.Date('1980-06-01'), as.Date('2010-11-01'), by = 'month')) %>%
  mutate(year = year(date)) %>%
  filter(between(month(date), 6, 11)) %>%
  left_join(sd_year_means) %>%
  select(date, `South Delta` = mean_area)

sd_hab <- south_delta_habitat %>%
  bind_rows(sd_means) %>%
  arrange(date)

delta_habitat <- left_join(nd_hab, sd_hab)

# use_data(delta_habitat, overwrite = TRUE)

knitr::kable(align = 'c', head(delta_habitat, 5))
```

*... with 357 more rows*

## North and South Deltas Rearing Habitat Plot

```{r}

delta_habitat %>% 
  gather(region, square_meters, -date) %>% 
  ggplot(aes(date, square_meters / 4046.86, color = region)) +
  geom_line() +
  theme_minimal() +
  scale_color_brewer(palette = 'Dark2') +
  labs(y = 'highly suitable habitat area (acres)')
```
