---
title: "Stanislaus River"
author: "sadie"
date: "December 12, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
library(scales)
```

## Instream Spawning and Rearing Habiat

> **Data Source:** Spawning Aceituno 1993 (FWS) PHABSIM model and the rearing from the FWS River2D model.
> Need to add links to souce documents
> <https://flowwest.github.io/cvpiaHabitat/reference/stanislaus_river_instream.html>

Mark Gard instructed us to use I4:N60 of the 'Stanislaus' tab within his 'IFIMWUA.xlsx'
These cells represent the sums of the rearing habitat in the canyon segment,
knights ferry to orange blossom, orange blossom to jacomb myers, and jacob myers
to san joaquin and then converted from sqft to sqft/1000ft. Spawning is modeled
as one 50 mi segment.

I saved this tab as 'stanislaus_IFIM.xlsx' to explore here.

#### WUA data in cvpiaHabitat package compared to mark gard raw data
```{r, echo=FALSE}
cvpiaHabitat::stanislaus_river_instream
stanislaus_raw <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', range = "I4:N60", sheet = 'Stanislaus',
                             col_names = c('flow_cfs', 'FR_fry_wua', 'ST_fry_wua', 'FR_juv_wua', 'ST_juv_wua', 'FR_spawn_wua'))
stanislaus_raw

```


### Spawning WUA from raw data

```{r, echo=FALSE}
stanislaus_raw %>%
  filter(!is.na(FR_spawn_wua)) %>%
  ggplot(aes(flow_cfs, FR_spawn_wua)) +
  geom_point() +
  labs(x = 'flow (cfs)', y = 'spawning WUA (sqft/1000ft)') +
  scale_y_continuous(labels = comma)
```

### Rearing WUA from raw data
```{r, echo=FALSE}
stanislaus_raw %>%
  filter(is.na(FR_spawn_wua)) %>%
  select(-FR_spawn_wua) %>%
  gather(species, wua, -flow_cfs) %>%
  ggplot(aes(flow_cfs, wua, color = species)) +
  geom_line() +
  labs(x = 'flow (cfs)', y = 'rearing WUA (sqft/1000ft)') +
  scale_y_continuous(labels = comma)
```

### Rearing Habitat in Acres
Using the habitat package `cvpiaHabitat::set_instream_habitat`

```{r, echo=FALSE}
stanislaus_river_instream <- stanislaus_raw %>% 
  mutate(watershed = 'Stanislaus River') %>% 
  select(flow_cfs, FR_spawn_wua, FR_fry_wua, FR_juv_wua, ST_fry_wua, ST_juv_wua, watershed)

devtools::use_data(stanislaus_river_instream, overwrite = TRUE)

```

