---
title: "Stanislaus River"
author: "sadie"
date: "December 12, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
library(scales)
```

## Instream Spawning and Rearing Habiat

> **Data Source:** Spawning Aceituno 1993 (FWS) PHABSIM model and the rearing from the FWS River2D model. TODO
> Need to add links to souce documents TODO
> <https://flowwest.github.io/cvpiaHabitat/reference/stanislaus_river_instream.html>

Description of modeling TODO

Mark Gard instructed us to use I4:N60 of the 'Stanislaus' tab within his 'IFIMWUA.xlsx'
These cells represent the sums of the rearing habitat in the canyon segment,
knights ferry to orange blossom, orange blossom to jacomb myers, and jacob myers
to san joaquin and then converted from sqft to sqft/1000ft. Spawning is modeled
as one 50 mi segment.

## Mark Gard WUA data 
```{r}
stanislaus_raw <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', range = "I4:N60", sheet = 'Stanislaus',
                             col_names = c('flow_cfs', 'FR_fry_wua', 'ST_fry_wua', 'FR_juv_wua', 'ST_juv_wua', 'FR_spawn_wua'))
stanislaus_raw

```


### Spawning WUA Plot

```{r, echo=FALSE}
stanislaus_raw %>%
  filter(!is.na(FR_spawn_wua)) %>%
  ggplot(aes(flow_cfs, FR_spawn_wua)) +
  geom_point() +
  labs(x = 'flow (cfs)', y = 'spawning WUA (sqft/1000ft)') +
  scale_y_continuous(labels = comma)
```

### Rearing WUA Plot
```{r}
stanislaus_raw %>%
  filter(is.na(FR_spawn_wua)) %>%
  select(-FR_spawn_wua) %>%
  gather(species, wua, -flow_cfs) %>%
  ggplot(aes(flow_cfs, wua, color = species)) +
  geom_line() +
  labs(x = 'flow (cfs)', y = 'rearing WUA (sqft/1000ft)') +
  scale_y_continuous(labels = comma)
```

## Format Data
Prepare data in proper form to be used in habitat setting functions.


```{r, echo=FALSE}
stanislaus_river_instream <- stanislaus_raw %>% 
  mutate(watershed = 'Stanislaus River') %>% 
  select(flow_cfs, FR_spawn_wua, FR_fry_wua, FR_juv_wua, ST_fry_wua, ST_juv_wua, watershed)

# devtools::use_data(stanislaus_river_instream, overwrite = TRUE)

```

