---
title: "Stanislaus River"
author: "[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com), [Erin Cain](ecain@flowwest.com)"
date: "June 20, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
library(scales)
```

## Future Data Improvements
None Planned 

## Instream Spawning and Rearing Habitat

**Data Source:** [Aceituno 1993](https://cvpiahabitat-r-package.s3-us-west-2.amazonaws.com/cvpia-sit-model-inputs/Aceituno1993_StanislausRiverIFIMStudy.pdf) (pg 19)
[FWS 2012](https://cvpiahabitat-r-package.s3-us-west-2.amazonaws.com/cvpia-sit-model-inputs/CVPIA_Annual_Progress_Report_Fiscal_Year_2012.pdf) (pg 6)


Instream spawning and rearing habitat for Fall Run Chinook Salmon and Steelhead in the Stanislaus River is based on data from Aceituno 1993 and FWS 2012. The Aceituno 1993 study uses the PHABSIM model which was developed in 1989 to look at the relationship between physical habitat availability and streamflow for Chinook Salmon Spawning in approximately 24 miles of the Stanislaus.THe original spawning data is on pg 19 of the report. THe rearing data comes from the 2012 FWS CVPIA Annual progress report.The FWS used four sites on the river and conducted 2-D hydraulic modeling of all the sites a graph representing the rearing totals for all the sites is on pg 6 of the report. The data was provided to us by [Mark Gard](mailto:mark_gard@fws.gov) from the U.S. Fish and Wildlife Service in a spreadsheet (see '~/cvpiaHabitat/data-raw/mark_gard_data/IFIMWUA.xlsx'). Mark Gard instructed us to use I4:N60 of the 'Stanislaus' tab within his 'IFIMWUA.xlsx' These cells represent the sums of the rearing habitat in the canyon segment, knights ferry to orange blossom, orange blossom to jacomb myers, and jacob myers to san joaquin and then converted from sqft to sqft/1000ft. Spawning is modeled
as one 50 mi segment (REPORT SAYS @$ MILES***THIS MIGHT BE THE REASON FOR THE DATA CONFUSION.

###Combine Spawning and Rearing Data 
The instream spawning and rearing habitat data described above for Fall Run Chinook Salmon (FR) and Steelhead (ST) is combined for use in the DSM in the following format. Data for spawning was taken at differnt csf then rearing data and these values can be found at the bottom of the table. Units are in square feet per 1000 feet.

```{r,include=FALSE}
 stanislaus_raw <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', range = "I4:N60", sheet = 'Stanislaus',
                              col_names = c('flow_cfs', 'FR_fry_wua', 'ST_fry_wua', 'FR_juv_wua', 'ST_juv_wua', 'FR_spawn_wua'))
 stanislaus_raw

```


```{r, echo=FALSE}
stanislaus_river_instream <- stanislaus_raw %>% 
   mutate(watershed = 'Stanislaus River') %>% 
   select(flow_cfs, FR_spawn_wua, FR_fry_wua, FR_juv_wua, ST_fry_wua, ST_juv_wua, watershed)
 
knitr::kable( align = 'c', head(stanislaus_river_instream, 5))
 # devtools::use_data(stanislaus_river_instream, overwrite = TRUE)
 
```

###### ...with 52 more rows.

### Spawning WUA 
The instream spawning habitat weighted usable areas for Fall Run Chinook Salmon in the Stanislaus River.

#### Spawning WUA Plot
The following plot shows the weighted usable spawning area in square feet per thousand feet of river for Fall Run Chinook Salmon. These area per length rates are multiplied by the total spawning reach length mapped by the SIT. 

```{r, echo=FALSE}
stanislaus_river_instream %>%
  filter(!is.na(FR_spawn_wua)) %>%
  ggplot(aes(flow_cfs, FR_spawn_wua)) +
  geom_line(color = '#7570b3' ) +
  labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') +
  theme_minimal() 
```

### Rearing WUA
The fry and juvenile instream rearing habitat weighted usable areas for Fall Run Chinook Salmon in the Stanislaus River. 

#### Salmon Rearing WUA Plot

The following plot shows the weighted usable rearing area in square feet per thousand feet of river for Fall Run Chinook Salmon fry and juvenile. These rates are multiplied by the total rearing reach length mapped by the SIT. 



``` {r,echo=FALSE}
stanislaus_river_instream %>% 
  gather(lifestage, wua, -flow_cfs, -watershed)  %>% 
  filter(!is.na(wua), lifestage != 'FR_spawn_wua', lifestage != "ST_fry_wua", lifestage != "ST_juv_wua") %>%
  mutate(lifestage = ifelse(lifestage == 'FR_fry_wua', 'fry', 'juvenile')) %>% 
  ggplot(aes(x = flow_cfs , y = wua, color = lifestage)) +
  geom_line() +
  labs(x = 'flow (cfs)', y = 'WUA (sqft/1000ft)') + theme_minimal() +
  scale_color_manual(values = c('#d95f02','#7570b3'))
```


#### Steelhead Rearing WUA Plot

The following plot shows the weighted usable rearing area in square feet per thousand feet of river for Steelhead fry and juvenile. These rates are multiplied by the total rearing reach length mapped by the SIT. 

``` {r,echo=FALSE}
stanislaus_river_instream %>% 
  gather(lifestage, wua, -flow_cfs, -watershed)  %>% 
  filter(!is.na(wua), lifestage != 'FR_spawn_wua', lifestage != 'ST_adult_wua',  lifestage != "FR_juv_wua", lifestage != "FR_fry_wua", lifestage != "ST_spawn_wua") %>%
  mutate(lifestage = ifelse(lifestage == 'ST_fry_wua', 'fry', 'juvenile')) %>% 
  ggplot(aes(x = flow_cfs , y = wua, color = lifestage)) +
  geom_line() +
  labs(x = 'flow (cfs)', y = 'WUA (sqft/1000ft)') + theme_minimal() +
  scale_color_manual(values = c('#d95f02','#7570b3'))
```


