---
title: "American River"
author: "sadie"
date: "December 11, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
library(scales)

```

## Instream Spawning and Rearing Habiat
**Data Source:** FWS 1985 Mark Gard IFIM study + Lower American River Biological Rationale,
Development and Performance of the Modified Flow Management Standard 2017 study

Need link to 1985 FWS study

Mark Gard instructed us to use A35:D46 of the 'American' tab within his 'IFIMWUA.xlsx'
These cells represent the sums of the reaches Sailor Bar, Ancil Hoffman, Watt, 
and H Street divided by 4. The units are sqft/1000ft. 

#### mark gard raw data
```{r}
american_raw <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', range = "A35:D46", sheet = 'American')
american_raw 

```

Chris Hammersmark has provided updated spawning WUA

## Chris Hammersmark New Data
Attachment B, Table B-4 and Figure B-8 on pdf page 91 and 92
<https://www.waterboards.ca.gov/waterrights/water_issues/programs/bay_delta/california_waterfix/exhibits/docs/PCWA/part2/ARWA-702.pdf>
saved here:
<https://s3-us-west-2.amazonaws.com/cvpia-reference-docs/ARWA-702.pdf>

pg 89 table b-3 contains the study reach lengths

WUA curves old and new

```{r, echo=FALSE}
spawn <- read_csv('data-raw/instream/american/american_river_spawning_wua.csv')

study_len <- 800+377+1553+522+3142+3103+1000+754+3442+1500

cs <- spawn %>%
  mutate(sqft_per_1000ft = wua_ft2/(study_len/1000)) %>% 
  select(Flow = flow_cfs, Spawning = sqft_per_1000ft) %>% 
  mutate(source = 'Chris')

american_raw %>% 
  select(Flow, Spawning) %>% 
  mutate(source = 'Mark') %>% 
  bind_rows(cs) %>% 
  # group_by(source) %>% mutate(max = max(Spawning)) %>% filter(max == Spawning)
  ggplot(aes(Flow, Spawning, color = source)) +
  geom_line()

```

Chris data converted to acres

```{r, echo=FALSE}
amer_len <- cvpiaHabitat::watershed_lengths %>% 
  filter(watershed == 'American River', lifestage == 'spawning') %>% 
  pull(feet)

cs %>% 
  mutate(acres = Spawning * amer_len / 1000 / 43560) %>% 
  ggplot(aes(Flow, acres)) +
  geom_line()

```

## Combine Mark Gard and Chris Hammersmark data

```{r, echo=FALSE}
rearing <- american_raw %>% 
  select(-Spawning) %>% 
  gather(lifestage, WUA, -Flow)
  

spawning <- cs %>% 
  select(-source) %>% 
  rename(WUA = Spawning) %>% 
  mutate(lifestage = 'Spawning')

american_river_instream <- rearing %>% 
  bind_rows(spawning) %>% 
  spread(lifestage, WUA) %>% 
  select(flow_cfs = Flow,
         FR_spawn_wua = Spawning,
         FR_fry_wua = `Fry Rearing`,
         FR_juv_wua = `Juv Rearing`) %>% 
  mutate(watershed = 'American River')

american_river_instream
# devtools::use_data(american_river_instream, overwrite = TRUE)
```
