---
title: "San Joaquin River"
author: "[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com)"
date: "December 17, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
library(scales)

```
## Future Data Improvements
[Future data improvements bullets]

## Instream Spawning and Rearing Habiat

**Data Source:** 
[add data source]

Instream spawning and rearing habitat for Fall Run Chinook Salmon in the San Joaquin River is based on data from [data source description]. The [DWR FERC instream spawning and rearing habitat data] was provided by [Mark Gard](mailto:mark_gard@fws.gov) from the U.S. Fish and Wildlife Service in a spreadsheet (see '~/cvpiaHabitat/data-raw/mark_gard_data/IFIMWUA.xlsx'). 


[There is no habitat modeling, using san joaquin tribs]


```{r,echo=FALSE}
get_approx_fry <- function(df) {approxfun(df$flow_cfs, df$FR_fry_wua, rule = 2)}
get_approx_juv <- function(df) {approxfun(df$flow_cfs, df$FR_juv_wua, rule = 2)}

# san joaq -----
cvpiaHabitat::merced_river_instream$flow_cfs %>% range()
cvpiaHabitat::stanislaus_river_instream$flow_cfs %>% range()
cvpiaHabitat::tuolumne_river_instream$flow_cfs %>% range()

san_joaq_juv <- list(get_approx_juv(cvpiaHabitat::merced_river_instream),
                     get_approx_juv(cvpiaHabitat::stanislaus_river_instream),
                     get_approx_juv(cvpiaHabitat::tuolumne_river_instream))

san_joaq_fry <- list(get_approx_fry(cvpiaHabitat::merced_river_instream),
                     get_approx_fry(cvpiaHabitat::stanislaus_river_instream),
                     get_approx_fry(cvpiaHabitat::tuolumne_river_instream))

flows <- cvpiaHabitat::stanislaus_river_instream$flow_cfs

san_joaquin_river_instream <- purrr::map_df(flows, function(flow) {
  wua_fry <- mean(purrr::map_dbl(1:length(san_joaq_juv), function(i){san_joaq_fry[[i]](flow)}))
  wua_juv <- mean(purrr::map_dbl(1:length(san_joaq_juv), function(i){san_joaq_juv[[i]](flow)}))
  tibble(flow_cfs = flow, FR_fry_wua = wua_fry, FR_juv_wua = wua_juv,
         watershed = 'San Joaquin River')
})
```

``` {r,echo=FALSE}
bind_rows(cvpiaHabitat::merced_river_instream,
          cvpiaHabitat::stanislaus_river_instream,
          cvpiaHabitat::tuolumne_river_instream,
          san_joaquin_river_instream) %>%
  ggplot(aes(x = flow_cfs, y= FR_fry_wua, color = watershed)) +
  geom_line() +
  theme_minimal()
```

``` {r,echo=FALSE}
bind_rows(cvpiaHabitat::merced_river_instream,
          cvpiaHabitat::stanislaus_river_instream,
          cvpiaHabitat::tuolumne_river_instream,
          san_joaquin_river_instream) %>%
  ggplot(aes(x = flow_cfs, y= FR_juv_wua, color = watershed)) +
  geom_line() +
  theme_minimal()


# san_joaquin_river_instream
# devtools::use_data(san_joaquin_river_instream, overwrite = TRUE)
```


### Rearing WUA
The fry and juvenile instream rearing habitat weighted usable areas for Fall Run Chinook Salmon in the San Joaquin River provided by Mark Gard (see '~/cvpiaHabitat/data-raw/mark_gard_data/IFIMWUA.xlsx') are used directly in the DSM. 

#### Rearing WUA Plot

The following plot shows the weighted usable rearing area in square feet per thousand feet of river for Fall Run Chinook Salmon fry and juvenile. These rates are multiplied by the total rearing reach length mapped by the SIT and applied for the flow in each month in the DSM.
``` {r,echo=FALSE}
san_joaquin_river_instream %>% 
  gather(lifestage, wua, -flow_cfs, -watershed)  %>% 
  filter(!is.na(wua), lifestage != 'FR_spawn_wua', lifestage != 'ST_adult_wua') %>%
  mutate(lifestage = ifelse(lifestage == 'FR_fry_wua', 'fry', 'juvenile')) %>% 
  ggplot(aes(x = flow_cfs , y = wua, color = lifestage)) +
  geom_line() +
  labs(x = 'flow (cfs)', y = 'WUA (sqft/1000ft)')
```

### Rearing Data
The rearing habitat data described above for Fall Run Chinook Salmon (FR) is for use in the DSM in the following format.
```{r,echo=FALSE}
san_joaquin_river_instream
```
