---
title: "Cow Creek"
author: "[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com)"
date: "July 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, out.width = '100%')
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
```
## Future Data Improvements
None planned. 

## Instream Spawning and Rearing Habitat
**Data Source:** [FWS 2011](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/FWS2011_SouthCowrpt.pdf){target="_blank"} (pg 19) and [Calculated with a relationship between flow and mean weighted usable area from watersheds in the region with modeled habitat](http://cvpia-habitat-docs-markdown.s3-website-us-west-2.amazonaws.com/watershed/Regional_Approximation.html){target="_blank"}

There were no watershed specific instream spawning habitat data available for Cow Creek. Instream spawning data for Cow Creek were developed using a regional approximation of similar watersheds in the Upper-mid Sacramento River Region. The data were then used in the Decision Support Model (DSM). 

The instream rearing habitat for Fall Run Chinook Salmon in South Cow Creek is based on data from the Fish and Wildlife Service South Cow Creek habitat assessment. The study gathered rearing data from the Valley Floor Segment and the Boero Segment of South Cow Creek. The rearing data from both stretches are provided in figures 9 and 10 on page 19. The rearing habitat data were provided to us by [Mark Gard](mailto:mark_gard@fws.gov) from the U.S. Fish and Wildlife Service in a [spreadsheet](https://cvpiahabitat-r-package.s3-us-west-2.amazonaws.com/IFIMWUA.xlsx). Mark Gard instructed us to the 'SouthCow' tab within the spreadsheet. The cells A2:C32 represent the Valley Floor Reach (5.11 miles) and cells A36:C66 represent the Boero Reach (1.68 miles). Units are in square feet per 1000 feet. 

### Rearing Data
```{r,echo=FALSE}
valley <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', 
                      range = "A2:C32", sheet = 'SouthCow') %>% 
   mutate(mi = 5.11)
 
boero <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', 
                      range = "A36:C66", sheet = 'SouthCow') %>% 
   mutate(mi = 1.68)
 
tot_len <- 5.11 + 1.68
 
cow_creek_instream <- bind_rows(valley, boero) %>% 
   gather(lifestage, sqft_per_mile, -Flows, -mi) %>% 
   group_by(Flows, lifestage) %>% 
   summarise(wua = sum(sqft_per_mile * mi/tot_len)/5.28) %>% 
   spread(lifestage, wua) %>% 
   ungroup() %>% 
   rename(flow_cfs = Flows, FR_fry_wua = `fall-run fry`, 
          FR_juv_wua = `fall-run juvenile`) %>% 
   mutate(watershed = 'Cow Creek')
knitr::kable(align = 'c', head(cow_creek_instream, 5), 
             caption = "Header Descriptions: flow_cfs = flow in cubic feet per second, 
             FR_fry_wua = Fall Run Chinook Fry WUA,
             FR_juv_wua = Fall Run Chinook Juvenile WUA, 
             watershed = section of stream modeled for CVPIA SDM")

# usethis::use_data(cow_creek_instream, overwrite = TRUE)
```

*...with 25 more rows.*

### Rearing WUA
The following plot shows the weighted usable rearing area (WUA) in square feet per thousand feet for Fall Run Chinook Salmon fry and juvenile. These rates are multiplied by the total rearing reach length mapped by the Science Integration Team (SIT) to estimate available habitat. 
``` {r,echo=FALSE}
cow_creek_instream %>% 
  gather(Lifestage, wua, -flow_cfs, -watershed)  %>% 
  filter(!is.na(wua), Lifestage != 'FR_spawn_wua') %>%
  mutate(Lifestage = ifelse(Lifestage == 'FR_fry_wua', 'Fry', 'Juvenile')) %>% 
  ggplot(aes(x = flow_cfs , y = wua, color = Lifestage)) +
  geom_line() +
  labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') + 
  theme_minimal() + 
  scale_color_manual(values = c('#d95f02','#7570b3')) +
  theme(legend.justification = c(1,0), legend.position = c(1,.1))
```

```{r, include = FALSE}
source('data-raw/watershed/floodplain_utils.R')
```

## Floodplain Rearing Habitat
**Data Source:** [Scaled from a Deer Creek flow to floodplain area relationship generated with a 1D HEC-RAS hydraulic model](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/DRAFT+Deer+Creek+Hydraulic+Models+Tech+Memo+6-08-07.pdf){target="_blank"}

### Fall Run
`r print_model_details('Cow Creek', 'fr')`

### Steelhead
`r print_model_details('Cow Creek', 'st')`

## Floodplain Data    
The areas represent total inundated area in acres.    
```{r, echo = FALSE}
# TODO Ask mark about the difference in outcome from his sheet
cow_creek_floodplain <- scale_fp_flow_area('Cow Creek')

knitr::kable(align = 'c', head(cow_creek_floodplain, 5), 
             caption = "Header Descriptions: flow_cfs = flow in cubic feet per second,
             FR_floodplain_acres = Fall Run Chinook floodplain acres, 
             ST_floodplain_acres = Steelhead floodplain acres,
             watershed = section of stream modeled for CVPIA SDM") 
# usethis::use_data(cow_creek_floodplain, overwrite = TRUE)
```

*...with 17 more rows*

## Floodplain Plot
```{r, echo = FALSE}
cow_creek_floodplain %>% 
  rename(`Fall Run Chinook` = FR_floodplain_acres,
         `Steelhead` = ST_floodplain_acres) %>% 
  gather(Species, acres, -flow_cfs, -watershed) %>% 
  ggplot(aes(flow_cfs, acres, color = Species)) +
  geom_line() +
  theme_minimal() +
  scale_color_brewer(palette = 'Dark2') +
  labs(x = 'Flow (cfs)', y = 'Total Inundated Acres') +
  theme(legend.justification = c(1,0), legend.position = c(1,.1))
```

