---
title: "Cow Creek"
author: "[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com), [Erin Cain](ecain@flowwest.com)"
date: "June 11, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
```
## Future Data Improvement
Instream habitat suitability relationships provided for the Voluntary Agreement process.

## Instream Rearing Habitat

**Data Source:**
[FWS 2011](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/FWS2011_SouthCowrpt.pdf) (pg 19)

Instream rearing habitat for Fall Run Chinook Salmon in South Cow Creek is based on data from the Fish and Wildlidfe Service South Cow Creek habitat asessment. This study gathered rearing data from the Valley Floor Segment and the Boero Segment of South Cow Creek. The rearing data from both stretches is provided in figures 9 and 10 on page 19. The rearing habitat data was provided to us by [Mark Gard](mailto:mark_gard@fws.gov) from the U.S. Fish and Wildlife Service in a spreadsheet (see '~/cvpiaHabitat/data-raw/mark_gard_data/IFIMWUA.xlsx'). Mark Gard instructed us to the 'SouthCow' tab within his 'IFIMWUA.xlsx'.The cells A2:C32 represent the Valley Floor Reach (5.11 miles) and cells A36:C66 represent the Boero Reach (1.68 miles). 

No watershed specific salmonid spawning habitat data was available for Cow Creek. A regional weighted usable area (WUA) relationship with flow was derived for Cow Creek by averaging the WUA values on Battle Creek, Butte Creek, Clear Creek, and Cottonwood Creek. The geomorphic and hydrologic conditions in Cow Creek (watershed area = 1107 sqkm; active channel width = 8.5 meters) are similar to those on Battle Creek (watershed area = 957 sqkm; active channel width = 22.5 m), Butte Creek (watershed area = 2123 sqkm; active channel width = 15.2 m), Clear Creek (watershed area = 645 sqkm; active channel width = 18 m), and Cottonwood Creek (watershed area = 2444 sqkm; active channel width = 28.25 m). The regional WUA relationships for Cow Creek were multiplied by the length of spawning and rearing extents mapped by the Science Integration Team (SIT).

## Instream Spawning and Rearing Habiat

### Combine Spawning and Rearing Data 
The instream spawning and rearing habitat data described above for Fall Run Chinook Salmon (FR) is combined for use in the DSM in the following format.

```{r, echo = FALSE}
#Spawning data
mean_wua_spawn <- upper_mid_sac_region_instream %>% 
  mutate(watershed = 'Cow Creek') %>% 
  select(flow_cfs, FR_spawn_wua) 

#Rearing data
valley <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', 
                      range = "A2:C32", sheet = 'SouthCow') %>% 
   mutate(mi = 5.11)
 
boero <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', 
                      range = "A36:C66", sheet = 'SouthCow') %>% 
   mutate(mi = 1.68)
 
tot_len <- 5.11 + 1.68
 
cow_creek_instream <- bind_rows(valley, boero) %>% 
   gather(lifestage, sqft_per_mile, -Flows, -mi) %>% 
   group_by(Flows, lifestage) %>% 
   summarise(wua = sum(sqft_per_mile * mi/tot_len)/5.28) %>% 
   spread(lifestage, wua) %>% 
   ungroup() %>% 
   rename(flow_cfs = Flows, FR_fry_wua = `fall-run fry`, 
          FR_juv_wua = `fall-run juvenile`) %>% 
   mutate(watershed = 'Cow Creek')

#combine spawning and rearing data 
cow_creek_instream <- cow_creek_instream %>% 
  left_join(mean_wua_spawn) %>% 
  select(flow_cfs, FR_spawn_wua, FR_fry_wua, FR_juv_wua, watershed)

knitr::kable(align = 'c', head(cow_creek_instream, 5))

#usethis::use_data(cow_creek_instream, overwrite = TRUE)
```

*...with 25 more rows*

### Spawning WUA 
The instream spawning habitat weighted usable areas for Fall Run Chinook Salmon in the Antelope Creek.

#### Spawning WUA Plot
The following plot shows the weighted usable spawning area in square feet per thousand feet of river for Fall Run Chinook Salmon. These area per length rates are multiplied by the total spawning reach length mapped by the SIT. 

```{r, echo=FALSE}
cow_creek_instream %>%
  filter(!is.na(FR_spawn_wua)) %>%
  ggplot(aes(flow_cfs, FR_spawn_wua)) +
  geom_line(color = '#7570b3' ) +
  labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') +
  theme_minimal() 
```

## Instream Rearing Habiat

### Rearing WUA
The fry and juvenile instream rearing habitat weighted usable areas for Fall Run Chinook Salmon in the Cow River. Units are in square feet per 1000 feet.

#### Rearing WUA Plot

The following plot shows the weighted usable rearing area in square feet per thousand feet of river for Fall Run Chinook Salmon fry and juvenile. These rates are multiplied by the total rearing reach length mapped by the SIT. 

``` {r,echo=FALSE}
cow_creek_instream %>% 
  gather(Lifestage, wua, -flow_cfs, -watershed)  %>% 
  filter(!is.na(wua), Lifestage != 'FR_spawn_wua') %>%
  mutate(Lifestage = ifelse(Lifestage == 'FR_fry_wua', 'Fry', 'Juvenile')) %>% 
  ggplot(aes(x = flow_cfs , y = wua, color = Lifestage)) +
  geom_line() +
  labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') + 
  theme_minimal() + 
  scale_color_manual(values = c('#d95f02','#7570b3'))
```

```{r, include = FALSE}
source('data-raw/floodplain/floodplain_utils.R')
```

## Floodplain Modeling Details

### Fall Run
`r print_model_details('Cow Creek', 'fr')`

### Steelhead
`r print_model_details('Cow Creek', 'st')`


## Floodplain Data

```{r, echo = FALSE}
# TODO Ask mark about the difference in outcome from his sheet
cow_creek_floodplain <- scale_fp_flow_area('Cow Creek')

knitr::kable(align = 'c', head(cow_creek_floodplain, 5)) 
# devtools::use_data(cow_creek_floodplain, overwrite = TRUE)
```

*...with 17 more rows*

## Floodplain Plot
```{r, echo = FALSE}
cow_creek_floodplain %>% 
  rename(`Fall Run Chinook` = FR_floodplain_acres,
         `Steelhead` = ST_floodplain_acres) %>% 
  gather(Species, acres, -flow_cfs, -watershed) %>% 
  ggplot(aes(flow_cfs, acres, color = Species)) +
  geom_line() +
  theme_minimal() +
  scale_color_brewer(palette = 'Dark2') +
  labs(x = 'Flow (cfs)', y = 'Total Inundated Acres')
```

