---
title: "Merced River"
author: "[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com)"
date: "July 2020"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, out.width = '100%')
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
```
## Future Data Improvements
None planned. 
 
## Instream Spawning and Rearing Habitat
**Data Source:** [Merced Irrigation District 2013](https://cvpiahabitat-r-package.s3-us-west-2.amazonaws.com/cvpia-sit-model-inputs/MID2013_MercedIFIMStudy.pdf) (pg 67-75)

Instream spawning and rearing habitat for Fall Run Chinook Salmon and Steelhead in the Merced River are based on data from a 2011-2012 Merced Irrigation District flow-habitat study. The study took place along the 19.2 mile section of the Merced River from the Crocker-Huffman Diversion Dam at river mile (RM) 52.01 to the Shaffer Bridge at river mile 32.8. Habitat data from this study are summarized on pages 67 - 75 of the MID 2013 study. The instream Fall Run Chinook Salmon spawning and rearing habitat data were provided to us by [Mark Gard](mailto:mark_gard@fws.gov) from the U.S. Fish and Wildlife Service in a [spreadsheet](https://cvpiahabitat-r-package.s3-us-west-2.amazonaws.com/IFIMWUA.xlsx). Mark Gard instructed us to use A51:D81, A84:D114, and A117:D147 of the 'Merced' tab within the spreadsheet. Steelhead data were also compiled from the Merced Irrigation District 2013 study. The three study reaches for both Fall Run Chinook Salmon and Steelhead were combined and weighted by reach length. 

### Combine Spawning and Rearing Data 
The instream spawning and rearing habitat data described above for Fall Run Chinook Salmon (FR) and Steelhead (ST) are combined for use in the Decision Support Model (DSM) in the following format. Units are in square feet per 1000 feet.    
```{r,echo=FALSE}
reach_miles <- c(9.2, 4.4, 5.6)
names(reach_miles) <- 1:3

# fall run data
fr_reach_1 <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', 
                         range = "A51:D81", sheet = 'Merced') %>% 
  mutate(reach = 1, miles = reach_miles[reach]) %>% 
  rename(flow_cfs = Flow, spawn = Spawning, fry = `Fry Rearing`, 
         juv = `Juv Rearing`) 

fr_reach_2 <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', 
                         range = "A84:D114", sheet = 'Merced') %>% 
  mutate(reach = 2, miles = reach_miles[reach]) %>% 
  rename(flow_cfs = Flow, spawn = Spawning, fry = `Fry Rearing`, 
         juv = `Juv Rearing`) 

fr_reach_3 <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', 
                         range = "A117:D147", sheet = 'Merced') %>% 
  mutate(reach = 3, miles = reach_miles[reach]) %>% 
  rename(flow_cfs = Flow, spawn = Spawning, fry = `Fry Rearing`, 
         juv = `Juv Rearing`) 

fr <- bind_rows(fr_reach_1, fr_reach_2, fr_reach_3) %>% 
  gather(life_stage, wua, -flow_cfs, -reach, -miles) %>% 
  mutate(species = 'FR')

make_approx_funs <- function(df) {
  
  flow <- df$flow_cfs
  spawn <- approxfun(flow, df$spawn, rule = 2)
  fry <- approxfun(flow, df$fry, rule = 2)
  juv <- approxfun(flow, df$juv, rule = 2)
  
  result <- list(spawn = spawn, fry = fry, juv = juv)

  if ('adult' %in% names(df)) {
    adult <- approxfun(flow, df$adult, rule = 2)
    result$adult = adult
  }
  
  return(result)
}

fr_approx_1 <- make_approx_funs(fr_reach_1)
fr_approx_2 <- make_approx_funs(fr_reach_2)
fr_approx_3 <- make_approx_funs(fr_reach_3)

# steelhead data
st <- read_csv('data-raw/watershed/merced/data/merced_steelhead.csv', 
               skip = 1) %>% 
  rename(spawn = ST_spawn_wua, fry = ST_fry_wua, juv = ST_juv_wua, 
         adult = ST_adult_wua) %>% 
  gather(life_stage, wua, -flow_cfs, -reach, -miles) %>% 
  mutate(species = 'ST')

st_reach_1 <- st %>% 
  filter(reach == 1) %>% 
  select(flow_cfs, life_stage, wua) %>% 
  spread(life_stage, wua)

st_reach_2 <- st %>% 
  filter(reach == 2) %>% 
  select(flow_cfs, life_stage, wua) %>% 
  spread(life_stage, wua)

st_reach_3 <- st %>% 
  filter(reach == 3) %>% 
  select(flow_cfs, life_stage, wua) %>% 
  spread(life_stage, wua)

st_approx_1 <- make_approx_funs(st_reach_1)
st_approx_2 <- make_approx_funs(st_reach_2)
st_approx_3 <- make_approx_funs(st_reach_3)

total <- sum(reach_miles)

flows <- bind_rows(fr, st) %>% 
  pull(flow_cfs) %>% 
  unique() %>% 
  sort()

# combine reaches weighted by length

merced_river_instream <- tibble(
  flow_cfs = flows,
  FR_spawn_wua = 
    fr_approx_1$spawn(flow_cfs) * reach_miles[1] / total + 
    fr_approx_2$spawn(flow_cfs) * reach_miles[2] / total + 
    fr_approx_3$spawn(flow_cfs) * reach_miles[3] / total,
  FR_fry_wua = 
    fr_approx_1$fry(flow_cfs) * reach_miles[1] / total +
    fr_approx_2$fry(flow_cfs) * reach_miles[2] / total +
    fr_approx_3$fry(flow_cfs) * reach_miles[3] / total,
  FR_juv_wua = 
    fr_approx_1$juv(flow_cfs) * reach_miles[1] / total +
    fr_approx_2$juv(flow_cfs) * reach_miles[2] / total +
    fr_approx_3$juv(flow_cfs) * reach_miles[3] / total,
  ST_spawn_wua =
    st_approx_1$spawn(flow_cfs) * reach_miles[1] / total + 
    st_approx_2$spawn(flow_cfs) * reach_miles[2] / total + 
    st_approx_3$spawn(flow_cfs) * reach_miles[3] / total,
  ST_fry_wua =
    st_approx_1$fry(flow_cfs) * reach_miles[1] / total + 
    st_approx_2$fry(flow_cfs) * reach_miles[2] / total + 
    st_approx_3$fry(flow_cfs) * reach_miles[3] / total,
  ST_juv_wua =
    st_approx_1$juv(flow_cfs) * reach_miles[1] / total + 
    st_approx_2$juv(flow_cfs) * reach_miles[2] / total + 
    st_approx_3$juv(flow_cfs) * reach_miles[3] / total,
  ST_adult_wua =
    st_approx_1$adult(flow_cfs) * reach_miles[1] / total + 
    st_approx_2$adult(flow_cfs) * reach_miles[2] / total + 
    st_approx_3$adult(flow_cfs) * reach_miles[3] / total,
  watershed = 'Merced River'
)

knitr::kable(align = 'c', head(merced_river_instream, 5), 
             caption = "Header Descriptions: flow_cfs = flow in cubic feet per second,
             FR_spawn_wua = Fall Run Chinook Spawning WUA, 
             FR_fry_wua = Fall Run Chinook Fry WUA, 
             FR_juv_wua = Fall Run Chinook Juvenile WUA, 
             ST_spawn_wua = Steelhead Spawning WUA,
             ST_fry_wua = Steelhead Fry WUA, 
             ST_juv_wua = Steelhead Juvenile WUA, 
             ST_adult_wua = Steelhead Adult WUA,
             watershed = section of stream modeled for CVPIA SDM")
# usethis::use_data(merced_river_instream, overwrite = TRUE)
```

*...with 50 more rows.* 

### Spawning WUA 
The following plot shows the weighted usable spawning area (WUA) in square feet per thousand feet for Fall Run Chinook Salmon and Steelhead. These area per length rates are multiplied by the total spawning reach length mapped by the Science Integration Team (SIT) to estimate available habitat. 
``` {r,echo=FALSE}
merced_river_instream %>% 
  select(flow_cfs, `Fall Run Chinook` = FR_spawn_wua, `Steelhead` = ST_spawn_wua) %>% 
  gather(Species, WUA, -flow_cfs) %>% 
  filter(!is.na(WUA)) %>% 
  ggplot(aes(flow_cfs, WUA, color = Species)) +
  geom_line() +
  labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') + 
  theme_minimal() + 
  scale_color_manual(values = c('#d95f02','#7570b3')) +
  theme(legend.justification = c(1,0), legend.position = c(.9,.55))

```

### Rearing WUA
The following plot shows the rearing WUA in square feet per thousand feet for Fall Run Chinook Salmon and Steelhead fry and juvenile. These rates are multiplied by the total rearing reach length mapped by the SIT to estimate available habitat. 
``` {r,echo=FALSE}
merced_river_instream %>% 
  select(FR_juv_wua, FR_fry_wua, ST_juv_wua, ST_fry_wua, flow_cfs, watershed) %>%
  gather(species_lifestage, acres, -flow_cfs, -watershed) %>% 
  mutate(
    Species = case_when( 
      str_detect(species_lifestage, 'ST') ~ 'Steelhead', 
      str_detect(species_lifestage, 'FR') ~ 'Fall Run'),
      lifestage = ifelse(str_detect(species_lifestage, 'juv'), 'Juvenile', 'Fry')) %>% 
  ggplot(aes(flow_cfs, acres, color = Species)) +
  geom_line() +
  theme_minimal() +
  facet_wrap(~lifestage, ncol = 1, scales = 'free_y') +
  scale_color_brewer(palette = 'Dark2') +
  labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') + 
  theme(legend.justification = c(1,0), legend.position = c(.95,.05))
```

### Adult WUA 
The following plot shows the rearing WUA in square feet per thousand feet for adult Steelhead. These rates are multiplied by the total rearing reach length mapped by the SIT to estimate available habitat. 
``` {r,echo=FALSE}
merced_river_instream %>% 
    filter(!is.na(ST_adult_wua)) %>% 
    ggplot(aes(x = flow_cfs , y = ST_spawn_wua)) +
    geom_line(color = '#7570b3') +
    labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') + 
    theme_minimal()
```
```{r, include=FALSE}
source('data-raw/watershed/floodplain_utils.R')
```

## Floodplain Rearing Habitat

**Data Source:**
[Central Valley Floodplain Evaluation and Delineation (CVFED) HEC-RAS hydraulic model](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/TO25-ST8_Tech_Memo_SJR.pdf)

`r print_model_details('Merced River', 'fr')`

## Floodplain Data
The areas represent total inundated area in acres.    
```{r, echo=FALSE}
merced_fp <- read_excel('data-raw/watershed/CVPIA_FloodplainAreas.xlsx', sheet = 'MercedRiver') %>%
  mutate(watershed = 'Merced River')

threshold <- merced_fp %>% 
  filter(modeled_floodplain_area_acres == 0) %>% 
  pull(flow_cfs) %>% 
  max()

merced_river_floodplain <- scale_fp_flow_area_partial_model(ws = 'Merced River', 
                                                            df = filter(merced_fp, flow_cfs >= threshold))

knitr::kable(align = 'c', head(merced_river_floodplain, 5), 
             caption = "Header Descriptions: flow_cfs = flow in cubic feet per second,
             FR_floodplain_acres = Fall Run Chinook floodplain acres, 
             ST_floodplain_acres = Steelhead floodplain acres,
             watershed = section of stream modeled for CVPIA SDM")
# usethis::use_data(merced_river_floodplain, overwrite = TRUE)
```

*...with 17 more rows*

## Floodplain Plot
```{r, echo = FALSE}
merced_river_floodplain %>% 
  select(FR_floodplain_acres, flow_cfs, watershed) %>%
  rename(`Fall Run Chinook and Steelhead` = FR_floodplain_acres) %>% 
  gather(Species, acres, -flow_cfs, -watershed) %>% 
  ggplot(aes(flow_cfs, acres, color = Species)) +
  geom_line() +
  theme_minimal() +
  scale_color_brewer(palette = 'Dark2') +
  labs(x = 'Flow (cfs)', y = 'Total Inundated Acres') + 
  theme(legend.justification = c(1,0), legend.position = c(.95,.05))
```
