---
title: "Feather River"
author: "[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com), [Erin Cain](ecain@flowwest.com)"
date: "August 15, 2019"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, out.width = '100%')
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
library(scales)
```
## Future Data Improvements
None planned

## Instream Spawning and Rearing Habitat
**Data Source:** [California Department of Water Resources 2004](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/Feather_FERC_IFIM_Phase_2.pdf){target="_blank"} (pgs 35-36 ) and [Thomas R. Payne & Associates 2002](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/Payne2002_FeatherRiverIFIM+7-22-02.pdf){target="_blank"} (pgs 35-37)

Instream spawning and rearing habitat for Fall Run Chinook Salmon and Steelhead in the Feather River are based on data from the California Department of Water Resources and Thomas R. Payne & Associates instream flow evaluations for the relicensing of Oroville facilities. These evaluations developed relationships between flow and suitable spawning and rearing habitat for 23.25 miles of the Feather River between the Fish Barrier Dam and Honcut Creek, consisting of two river segments. The DWR FERC instream spawning and rearing habitat data was provided by [Mark Gard](mailto:mark_gard@fws.gov) from the U.S. Fish and Wildlife Service in a spreadsheet (see '~/cvpiaHabitat/data-raw/mark_gard_data/IFIMWUA.xlsx'). The Fall Run Chinook Salmon and Steelhead spawning habitat data from Gard was cross-referenced with the original FERC data and amended to include all of the original data in the FERC reports by adding values digitized from the flow area curves in the [FERC evaluation reports (pages 35-36)](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/Feather_FERC_IFIM_Phase_2.pdf) using <https://automeris.io/WebPlotDigitizer/>. The FERC instream flow evaluation produced flow area curves for "upper" and "lower" reaches. These values were combined into a single area weighted by study reach length for each flow value.

### Combine Spawning and Rearing Data
The instream spawning and rearing habitat data described above for Fall Run Chinook Salmon (FR) and Steelhead (ST) is combined for use in the DSM in the following format.
```{r, echo=FALSE}
feather_raw <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', range = "A1:D8", sheet = 'Feather')
# feather_raw
```
```{r, echo=FALSE}
upper_fr <- read_csv('data-raw/watershed/feather/data/alt_upper_feather_spawn_wua_rsi.csv') %>%
  mutate(location = 'upper')
lower_fr <- read_csv('data-raw/watershed/feather/data/alt_lower_feather_spawn_wua_rsi.csv') %>%
  mutate(location = 'lower')
upper_st <- read_csv('data-raw/watershed/feather/data/alt_upper_feather_spawn_wua_rsi_steelhead.csv') %>%
  mutate(location = 'upper')
lower_st <- read_csv('data-raw/watershed/feather/data/alt_lower_feather_spawn_wua_rsi_steelhead.csv') %>%
  mutate(location = 'lower')
spawning <- tibble(
  flow_cfs = c(200, 500, 750, 1000, 1250, 1500, 1750, 2000, 2500, 3000, 3500, 
               4000, 4500, 5000, 5500, 6000, 6500, 7000),
  upper_wua_fr = approx(upper_fr$flow, upper_fr$wua_rsi, xout = flow_cfs, rule = 2)$y,
  lower_wua_fr = approx(lower_fr$flow, lower_fr$wua_rsi, xout = flow_cfs, rule = 2)$y,
  upper_wua_st = approx(upper_st$flow, upper_st$wua_rsi, xout = flow_cfs, rule = 2)$y,
  lower_wua_st = approx(lower_st$flow, lower_st$wua_rsi, xout = flow_cfs, rule = 2)$y) %>% 
  mutate(FR_spawn_wua = upper_wua_fr * (9.75/18) + lower_wua_fr * (8.25/18),
         ST_spawn_wua = upper_wua_st * (9.75/18) + lower_wua_st * (8.25/18)) %>%
  select(flow_cfs, FR_spawn_wua, ST_spawn_wua)
```
```{r, echo=FALSE}
feather_river_instream <- feather_raw %>% 
  select(-Spawning) %>% 
  rename(flow_cfs = Flow, FR_fry_wua = `Fry Rearing`, FR_juv_wua = `Juv Rearing`) %>% 
  full_join(spawning) %>% 
  arrange(flow_cfs) %>% 
  mutate(watershed = 'Feather River') %>%
  select(flow_cfs,
         FR_spawn_wua,
         FR_fry_wua,
         FR_juv_wua,
         ST_spawn_wua,
         watershed)

knitr::kable(align ="c", head(feather_river_instream, 5), 
             caption = "Header Descriptions: flow_cfs = flow in cubic feet per second,
             FR_spawn_wua = Fall Run Chinook Spawning WUA, 
             FR_fry_wua = Fall Run Chinook Fry WUA, 
             FR_juv_wua = Fall Run Chinook Juvenile WUA, 
             ST_spawn_wua = Steelhead Spawning WUA")

# usethis::use_data(feather_river_instream, overwrite = TRUE)
```

*...with 13 more rows.*

### Spawning WUA
The following plot shows the weighted usable spawning area in square feet per thousand feet of river for Fall Run Chinook Salmon and Steelhead. These area per length rates are multiplied by the total spawning reach length mapped by the SIT and applied for the flow in each month in the DSM. 
```{r, echo = FALSE}
feather_river_instream %>% 
  select(flow_cfs, `Fall Run Chinook` = FR_spawn_wua, `Steelhead` = ST_spawn_wua) %>% 
  gather(Species, WUA, -flow_cfs) %>% 
  filter(!is.na(WUA)) %>% 
  ggplot(aes(flow_cfs, WUA, color = Species)) +
  geom_line() +
  labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') + 
  theme_minimal() + 
  scale_color_manual(values = c('#d95f02','#7570b3')) +
  theme(legend.justification = c(1,0), legend.position = c(.9,.7))
```
 
### Rearing WUA
The following plot shows the weighted usable rearing area in square feet per thousand feet of river for Fall Run Chinook Salmon fry and juvenile. These rates are multiplied by the total rearing reach length mapped by the SIT.  
```{r, echo=FALSE}
feather_river_instream %>% 
  gather(lifestage, wua, -flow_cfs, -watershed)  %>% 
  filter(!is.na(wua), lifestage != 'FR_spawn_wua', lifestage != 'ST_adult_wua', lifestage != "ST_spawn_wua", lifestage != "ST_juv_wua", lifestage != "ST_fry_wua") %>%
  mutate(lifestage = ifelse(lifestage == 'FR_fry_wua', 'fry', 'juvenile')) %>% 
  ggplot(aes(x = flow_cfs , y = wua, color = lifestage)) +
  geom_line() +
  labs(x = 'flow (cfs)', y = 'WUA (sqft/1000ft)') + theme_minimal() +
  scale_color_manual(values = c('#d95f02','#7570b3')) +
  theme(legend.justification = c(1,0), legend.position = c(.9,.5))
```
```{r, include=FALSE}
source('data-raw/watershed/floodplain_utils.R')
```

## Floodplain Rearing Habitat
**Data Source:** [Central Valley Floodplain Evaluation and Delineation (CVFED) HEC-RAS hydraulic model](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/CombinedTM_IQAR_Final-FULL-REPORT_20140206.pdf){target="_blank"}

`r print_model_details('Feather River', 'fr')`

## Floodplain Data
The areas represent total inundated area in acres.    
```{r, echo=FALSE}
feather_river_floodplain <- read_excel('data-raw/watershed/CVPIA_FloodplainAreas.xlsx',
                                       sheet = 'FeatherRiver') %>%
  mutate(watershed = 'Feather River',
         SR_floodplain_acres = modeled_floodplain_area_acres,
         ST_floodplain_acres = modeled_floodplain_area_acres) %>%
  select(flow_cfs,
         FR_floodplain_acres = modeled_floodplain_area_acres,
         SR_floodplain_acres, ST_floodplain_acres,
         watershed)

knitr::kable(align = 'c', head(feather_river_floodplain, 5), 
             caption = "Header Descriptions: flow_cfs = flow in cubic feet per second,
             FR_floodplain_acres = Fall Run Chinook floodplain acres, 
             SR_floodplain_acres = Spring Run Chinook floodplain acres, 
             ST_floodplain_acres = Steelhead floodplain acres")
# usethis::use_data(feather_river_floodplain, overwrite = TRUE)
```

*...with 35 more rows*

## Floodplain Plot
```{r, echo = FALSE}
feather_river_floodplain %>% 
  select(FR_floodplain_acres, flow_cfs, watershed) %>%
  rename(`Steelhead, 
Fall Run Chinook, 
and Spring Run Chinook` = FR_floodplain_acres) %>% 
  gather(Species, acres, -flow_cfs, -watershed) %>% 
  ggplot(aes(flow_cfs, acres, color = Species)) +
  geom_line() +
  theme_minimal() +
  scale_color_brewer(palette = 'Dark2') +
  labs(x = 'Flow (cfs)', y = 'Total Inundated Acres') +
  theme(legend.justification = c(1,0), legend.position = c(.95,.1))
```
