---
title: "Steelhead Adult Habitat Scaling"
author: "sadie"
date: "December 18, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
library(lubridate)
library(cvpiaFlow)
```


This document develops a method for scaling the available adult steelhead wua data
to the other watersheds

## Adult Trout and Steelhead WUAs

There is habitat modeling for adult trout or steelhead on 5 watersheds:    

**Adult Trout**

1. Battle Creek    
2. Butte Creek    

**Steelhead**

3. Merced River   
4. Tuolumne River   
5. Yuba River   


## Modeled WUA Exploration

```{r, echo=FALSE}
bac <- cvpiaHabitat::battle_creek_instream %>%
  select(flow_cfs, ST_adult_wua, watershed)
buc <- cvpiaHabitat::butte_creek_instream %>%
  select(flow_cfs, ST_adult_wua, watershed)
mr <- cvpiaHabitat::merced_river_instream %>%
  select(flow_cfs, ST_adult_wua, watershed)
tr <- cvpiaHabitat::tuolumne_river_instream %>%
  select(flow_cfs, ST_adult_wua, watershed)
yr <- cvpiaHabitat::yuba_river_instream %>%
  select(flow_cfs, ST_adult_wua, watershed)

bind_rows(bac, buc, mr, tr, yr) %>%
  filter(flow_cfs < 10000) %>%
  ggplot(aes(flow_cfs, ST_adult_wua, color = watershed)) +
  geom_line()

```


Yuba and Tuolumne WUA values are much higher than the other watersheds, they
appear to be proporitonately higher based on watershed size. 

Below are the mean values of flow in the monthes December through May. Tuolumne 
is about twice that of Merced and Yuba is about four times Merced. This is 
similar to the differences between these watershed's WUA curves.

```{r, echo=FALSE}
cvpiaFlow::flows_cfs %>%
  select(date, `Tuolumne River`, `Yuba River`, `Merced River`) %>%
  gather(watershed, flow, -date) %>%
  filter(month(date) %in% c(1:5, 12), year(date) >= 1980) %>%
  group_by(watershed) %>%
  summarise(mean = mean(flow, na.rm = TRUE)) %>%
  ungroup() %>% 
  knitr::kable(col.names = c('Watershed', 'Mean Flow (cfs)'))
```

In the following plot, the black line represents the ratio of Yuba/Merced mean flows 
and the red line represents the ration of Tuolumne/Merced mean flows. 

```{r, echo=FALSE}
bind_rows(bac, buc, mr, tr, yr) %>%
  filter(flow_cfs < 10000) %>%
  ggplot(aes(flow_cfs, ST_adult_wua, color = watershed)) +
  geom_line() +
  geom_line(data = mr, aes(flow_cfs, ST_adult_wua * 4.19), color = 'black') +
  geom_line(data = mr, aes(flow_cfs, ST_adult_wua * 1.78), color = 'red')

```

## Method

For unmodeled watersheds, we will take the annual mean flow for the watershed
and calculate the wua value using Merced's modeled relationship. We will then
scale the wua by the ratio of mean flow between the watershed and Merced.

Here is a plot of the ratios (the black verticle line is at 1)

```{r, echo=FALSE}
mean_flows <- cvpiaFlow::flows_cfs %>%
  filter(month(date) %in% c(1:5, 12), between(year(date), 1980, 1999)) %>%
  summarise_all(mean, na.rm=TRUE) %>%
  select(-date) %>%
  gather(watershed, mean_flow)

merced <- mean_flows %>%
  filter(watershed == 'Merced River') %>%
  pull(mean_flow)

prop_merced_flows <- mean_flows %>%
  mutate(prop_merced = mean_flow/merced)

prop_merced_flows %>%
  ggplot(aes(watershed, prop_merced)) +
  geom_col() +
  geom_hline(yintercept = 1) +
  coord_flip()
```

Below is the esitmated adult steelhead habitat in acres for 1980
```{r, echo=FALSE}

others <- approxfun(merced_river_instream$flow_cfs, merced_river_instream$ST_adult_wua, rule = 2)
other_st_adult <- prop_merced_flows %>% 
  mutate(ST_adult_wua = others(mean_flow) * prop_merced) %>% 
  full_join(cvpiaHabitat::watershed_lengths) %>% 
  filter(lifestage == 'rearing', species != 'sr') %>% 
  select(-source, -lifestage) %>% 
  group_by(watershed) %>% 
  filter(max(feet) == feet) %>% 
  ungroup() %>% 
  mutate(sq_meters = feet/1000 * ST_adult_wua / 10.7639,
         acres = square_meters_to_acres(sq_meters)) %>% 
  filter(!(watershed %in% c("Lower-mid Sacramento River", "North Delta", "South Delta",
                            "Tuolumne River", "Merced River", "Yuba River",
                            "Butte Creek", "Battle Creek"))) %>%
  select(watershed, sq_meters, order)

IChab.adult <- array(NA, dim = c(31, 12, 20))
for (i in 1:dim(other_st_adult)[1]) {
  IChab.adult[other_st_adult$order[i], , ] <- other_st_adult$sq_meters[i]
}

flows <- cvpiaFlow::flows_cfs %>%
  select(date, `Tuolumne River`, `Yuba River`, `Merced River`, `Battle Creek`, `Butte Creek`) %>% 
  filter(between(year(date), 1980, 1999))

fts <- cvpiaHabitat::watershed_lengths %>% 
  filter(watershed %in% c('Tuolumne River', 'Yuba River', 'Merced River', 
                          'Butte Creek', 'Battle Creek'), lifestage == 'rearing') %>% 
  group_by(watershed) %>% 
  summarise(feet = max(feet))

feet <- fts$feet
names(feet) <- fts$watershed


IChab.adult[30,,] <- approx(x = tuolumne_river_instream$flow_cfs, 
                            y = tuolumne_river_instream$ST_adult_wua,
                            xout = flows$`Tuolumne River`, rule = 2)$y * feet[['Tuolumne River']] / 1000 / 10.7639

IChab.adult[28,,] <- approx(x = merced_river_instream$flow_cfs, 
                            y = merced_river_instream$ST_adult_wua,
                            xout = flows$`Merced River`, rule = 2)$y * feet[['Merced River']] / 1000 / 10.7639

IChab.adult[20,,] <- approx(x = yuba_river_instream$flow_cfs, 
                            y = yuba_river_instream$ST_adult_wua,
                            xout = flows$`Yuba River`, rule = 2)$y * feet[['Yuba River']] / 1000 / 10.7639

IChab.adult[3,,] <- approx(x = battle_creek_instream$flow_cfs, 
                            y = battle_creek_instream$ST_adult_wua,
                            xout = flows$`Battle Creek`, rule = 2)$y * feet[['Battle Creek']] / 1000 / 10.7639

IChab.adult[6,,] <- approx(x = butte_creek_instream$flow_cfs, 
                            y = butte_creek_instream$ST_adult_wua,
                            xout = flows$`Butte Creek`, rule = 2)$y * feet[['Butte Creek']] / 1000 / 10.7639

#lower-mid sacramento scale wua using two flows
low_mid_flows <- cvpiaFlow::flows_cfs %>% 
  select(date, `Lower-mid Sacramento River1`, `Lower-mid Sacramento River2`) %>% 
  filter(between(year(date), 1980, 1999)) %>% 
  summarise_all(mean)

low_mid_ft <- watershed_lengths %>% 
  filter(watershed == 'Lower-mid Sacramento River') %>% pull(feet)

IChab.adult[21,,] <- (35.6/58 * others(low_mid_flows$`Lower-mid Sacramento River1`) + 
  22.4/58 * others(low_mid_flows$`Lower-mid Sacramento River2`)) * low_mid_ft / 1000 / 10.7639 

# devtools::use_data(IChab.adult)
IChab.adult[,,1]  %>% square_meters_to_acres()
```
