---
title: "Steelhead Adult Rearing Habitat"
author: "[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com)"
date: "December 18, 2018 (Updated July 2020)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, out.width = '100%')
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
library(lubridate)
library(cvpiaFlow)
options(usethis.quiet = TRUE)
```


This document develops a method for scaling the available adult steelhead WUA data
to the other watersheds.

## Adult Trout and Steelhead WUAs

There is habitat modeling for adult trout or steelhead on 5 watersheds:    

**Adult Trout**

1. Battle Creek    
2. Butte Creek    

**Steelhead**

3. Merced River   
4. Tuolumne River   
5. Yuba River   


## Modeled WUA Exploration

```{r}
bac <- cvpiaHabitat::battle_creek_instream %>%
  select(flow_cfs, ST_adult_wua, watershed)
buc <- cvpiaHabitat::butte_creek_instream %>%
  select(flow_cfs, ST_adult_wua, watershed)
mr <- cvpiaHabitat::merced_river_instream %>%
  select(flow_cfs, ST_adult_wua, watershed)
tr <- cvpiaHabitat::tuolumne_river_instream %>%
  select(flow_cfs, ST_adult_wua, watershed)
yr <- cvpiaHabitat::yuba_river_instream %>%
  select(flow_cfs, ST_adult_wua, watershed)

bind_rows(bac, buc, mr, tr, yr) %>% 
  filter(flow_cfs < 10000, !is.na(ST_adult_wua)) %>%
  ggplot(aes(flow_cfs, ST_adult_wua, color = watershed)) +
  geom_line() +
  theme_minimal() +
  labs(x = "Flow (cfs)", y = "WUA (sqft/1000ft)")

```


Yuba and Tuolumne WUA values are much higher than the other watersheds, they
appear to be proporitonately higher based on watershed size. 

Below are the mean values of flow in the monthes December through May. Tuolumne 
is about twice that of Merced and Yuba is about four times Merced. This is 
similar to the differences between these watershed's WUA curves.

```{r}
mean_flow_ratio <- cvpiaFlow::flows_cfs %>%
  select(date, `Tuolumne River`, `Yuba River`, `Merced River`) %>%
  gather(watershed, flow, -date) %>%
  filter(month(date) %in% c(1:5, 12), year(date) >= 1980) %>%
  group_by(watershed) %>%
  summarise(mean = mean(flow, na.rm = TRUE)) %>%
  ungroup() %>% 
  mutate(ratio = round(mean/min(mean), 2))

mean_flow_ratio %>% 
  knitr::kable(col.names = c('Watershed', 'Mean Flow (cfs)', 'Mean Flow Ratio (watershed:Merced)'))
```

The following plot shows:

* Yubas's WUA relationship along with Merced's WUA scaled using the ratio of Yuba/Merced mean flows and the ratio 

* Tuolumne's WUA relationship is plotted along with Merced's WUA scaled using the ratio of Tuolumne/Merced mean flows 

* Merced's WUA relationship

```{r}
yrm <- mr %>% 
  mutate(ST_adult_wua = ST_adult_wua * mean_flow_ratio$ratio[mean_flow_ratio$watershed == "Yuba River"],
         watershed = "Yuba:Merced", ratio = TRUE, watershed2 = "Yuba River")

trm <- mr %>% 
  mutate(ST_adult_wua = ST_adult_wua * mean_flow_ratio$ratio[mean_flow_ratio$watershed == "Tuolumne River"],
         watershed = "Tuolumne:Merced", ratio = TRUE, watershed2 = "Tuolumne River")

bind_rows(mr, tr, yr) %>%
  mutate(ratio = FALSE, watershed2 = watershed) %>% 
  bind_rows(yrm, trm) %>%
  filter(flow_cfs < 10000, !is.na(ST_adult_wua)) %>% 
  ggplot(aes(flow_cfs, ST_adult_wua, color = watershed2, linetype = ratio, group = watershed)) +
  geom_line() +
  theme_minimal() +
  labs(x = "Flow (cfs)", y = "WUA (sqft/1000ft)", color = "watershed", linetype = "scaled by merced flow ratio") +
  scale_y_continuous(labels = scales::comma)

```

## Scaling Method

For unmodeled watersheds, we will take the annual mean flow for the watershed
and calculate the wua value using Merced's modeled relationship. We will then
scale the wua by the ratio of mean flow between the watershed and Merced.

For the unmodeled watersheds, we will calculate the mean flow value from 1980-2000
during the months December-May. With these mean values we will calculate the 
ratio between the watershed and Merced. We will estimate the adult rearing habitat 
for the watersheduse this ratio to scale Merced's WUA values


Here is a plot of the ratios (the black verticle line is at 1)

```{r}
mean_flows <- cvpiaFlow::flows_cfs %>%
  filter(month(date) %in% c(1:5, 12), between(year(date), 1980, 2000)) %>%
  summarise_all(mean, na.rm = TRUE) %>%
  select(-date) %>%
  gather(watershed, mean_flow)

merced <- mean_flows %>%
  filter(watershed == 'Merced River') %>%
  pull(mean_flow)

prop_merced_flows <- mean_flows %>%
  mutate(prop_merced = mean_flow/merced)

prop_merced_flows %>%
  ggplot(aes(watershed, prop_merced)) +
  geom_col() +
  geom_hline(yintercept = 1) +
  coord_flip() +
  theme_minimal() +
  labs(y = "watershed mean flow : merced mean flow")
```


```{r}
merced_approx <- approxfun(merced_river_instream$flow_cfs, merced_river_instream$ST_adult_wua, rule = 2)

other_st_adult <- prop_merced_flows %>% 
  mutate(ST_adult_wua = merced_approx(mean_flow) * prop_merced) %>% 
  full_join(cvpiaHabitat::watershed_lengths) %>% 
  filter(lifestage == 'rearing', species != 'sr') %>% 
  select(-source, -lifestage) %>% 
  group_by(watershed) %>% 
  filter(max(feet) == feet) %>% 
  ungroup() %>% 
  mutate(sq_meters = feet/1000 * ST_adult_wua / 10.7639,
         acres = square_meters_to_acres(sq_meters)) %>% 
  filter(!(watershed %in% c("Lower-mid Sacramento River", "North Delta", "South Delta",
                            "Tuolumne River", "Merced River", "Yuba River",
                            "Butte Creek", "Battle Creek"))) %>%
  select(watershed, sq_meters, order)


inchannel_habitat_adult <- array(NA, dim = c(31, 12, 21))
for (i in 1:dim(other_st_adult)[1]) {
  inchannel_habitat_adult[other_st_adult$order[i], , ] <- other_st_adult$sq_meters[i]
}

flows <- cvpiaFlow::flows_cfs %>%
  select(date, `Tuolumne River`, `Yuba River`, `Merced River`, `Battle Creek`, `Butte Creek`) %>% 
  filter(between(year(date), 1980, 2000))

fts <- cvpiaHabitat::watershed_lengths %>% 
  filter(watershed %in% c('Tuolumne River', 'Yuba River', 'Merced River', 
                          'Butte Creek', 'Battle Creek'), lifestage == 'rearing') %>% 
  group_by(watershed) %>% 
  summarise(feet = max(feet))

feet <- fts$feet
names(feet) <- fts$watershed


inchannel_habitat_adult[30, , ] <- approx(x = tuolumne_river_instream$flow_cfs, 
                                          y = tuolumne_river_instream$ST_adult_wua,
                                          xout = flows$`Tuolumne River`, rule = 2)$y * feet[['Tuolumne River']] / 1000 / 10.7639

inchannel_habitat_adult[28, , ] <- approx(x = merced_river_instream$flow_cfs, 
                                          y = merced_river_instream$ST_adult_wua,
                                          xout = flows$`Merced River`, rule = 2)$y * feet[['Merced River']] / 1000 / 10.7639

inchannel_habitat_adult[20, , ] <- approx(x = yuba_river_instream$flow_cfs, 
                                          y = yuba_river_instream$ST_adult_wua,
                                          xout = flows$`Yuba River`, rule = 2)$y * feet[['Yuba River']] / 1000 / 10.7639

inchannel_habitat_adult[3, , ] <- approx(x = battle_creek_instream$flow_cfs, 
                                         y = battle_creek_instream$ST_adult_wua,
                                         xout = flows$`Battle Creek`, rule = 2)$y * feet[['Battle Creek']] / 1000 / 10.7639

inchannel_habitat_adult[6, , ] <- approx(x = butte_creek_instream$flow_cfs, 
                                         y = butte_creek_instream$ST_adult_wua,
                                         xout = flows$`Butte Creek`, rule = 2)$y * feet[['Butte Creek']] / 1000 / 10.7639

# lower-mid sacramento scale wua using two flows
low_mid_flows <- cvpiaFlow::flows_cfs %>% 
  select(date, `Lower-mid Sacramento River1`, `Lower-mid Sacramento River2`) %>% 
  filter(between(year(date), 1980, 2000)) %>% 
  summarise_all(mean)

low_mid_ft <- watershed_lengths %>% 
  filter(watershed == 'Lower-mid Sacramento River') %>% pull(feet)

inchannel_habitat_adult[21, , ] <- (35.6/58 * merced_approx(low_mid_flows$`Lower-mid Sacramento River1`) + 
                                      22.4/58 * merced_approx(low_mid_flows$`Lower-mid Sacramento River2`)) * low_mid_ft / 1000 / 10.7639 

usethis::use_data(inchannel_habitat_adult, overwrite = TRUE)
```
The estimated mean adult rearing habitat for unmodeled watersheds is below 
```{r}
other_st_adult %>% 
  select(watershed, `square meters` = sq_meters) %>% 
      knitr::kable(align = 'c')
```
