---
title: "Deer Creek"
author: "[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com)" 
date: "July 2020"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, out.width = '100%')
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
```    
```{r}
source('data-raw/watershed/floodplain_utils.R')
```
## Future Data Improvements
Update with new Flowwest 2019 data. 

## Instream Spawning and Rearing Habitat
**Data Source:** [Calculated with a relationship between flow and mean weighted usable area from watersheds in the region with modeled habitat](http://cvpia-habitat-docs-markdown.s3-website-us-west-2.amazonaws.com/watershed/Regional_Approximation.html){target="_blank"}
```{r, include=FALSE}
source('data-raw/watershed/regional_approximation/description_regional_approx.R')
```

`r print_regional_approx('Deer Creek')`

```{r, include=FALSE}
source('data-raw/watershed/floodplain_utils.R')
```

## Floodplain Rearing Habitat 
**Data Source:** [Deer Creek Watershed Conservancy (DCWC) 1D HEC-RAS model](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/DRAFT+Deer+Creek+Hydraulic+Models+Tech+Memo+6-08-07.pdf){target="_blank"}

### Fall Run
`r print_model_details('Deer Creek', 'fr')`

### Spring Run and Steelhead
`r print_model_details('Deer Creek', 'sr')`

## Floodplain Data    
The areas represent total inundated area in acres.    
```{r}
deer_fp <- read_excel('data-raw/watershed/CVPIA_FloodplainAreas.xlsx', sheet = 'DeerCreek') %>%
  mutate(watershed = 'Deer Creek')

threshold_flow <- deer_fp %>% 
  filter(modeled_floodplain_area_acres == 0) %>% 
  summarise(max_flow = max(flow_cfs)) %>% 
  pull(max_flow)

deer_creek_floodplain <- scale_fp_flow_area_partial_model(ws = 'Deer Creek', df = filter(deer_fp, flow_cfs >= threshold_flow))

knitr::kable(align = 'c', head(deer_creek_floodplain, 5), 
             caption = "Header Descriptions: flow_cfs = flow in cubic feet per second,
             FR_floodplain_acres = Fall Run Chinook floodplain acres, SR_floodplain_acres 
             = Spring Run Chinook floodplain acres, ST_floodplain_acres = 
             Steelhead floodplain acres, 
             watershed = section of stream modeled for CVPIA SDM")
# usethis::use_data(deer_creek_floodplain, overwrite = TRUE)
```

*...with 17 more rows.*

## Floodplain Plot
```{r}
deer_creek_floodplain %>% 
  select(FR_floodplain_acres, SR_floodplain_acres, watershed, flow_cfs) %>%
  rename(`Fall Run Chinook` = FR_floodplain_acres, `Spring Run Chinook and Steelhead` = SR_floodplain_acres) %>% 
  gather(Species, acres, -flow_cfs, -watershed) %>% 
  ggplot(aes(flow_cfs, acres, color = Species)) +
  geom_line() +
  theme_minimal() +
  scale_color_brewer(palette = 'Dark2') +
  labs(x = 'Flow (cfs)', y = 'Total Inundated Acres') +
  theme(legend.justification = c(1,0), legend.position = c(1,.1))
```
