---
title: "Battle Creek"
author: "[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com), [Erin Cain](mailto: ecain@flowwest.com)"
date: "June 6, 2019"
output: html_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo=FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
```
## Future Data Improvements
None

## Instream Spawning and Rearing Habitat

**Data Source:**   
[Thomas R. Payne & Associates 1995](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/Payne1995_BattleCreekIFIM.pdf) (pgs. C-1 to C-9)

Instream spawning and rearing habitat for Fall Run Chinook Salmon and Steelhead in Battle Creek is based on data from Thomas R. Payne & Associates instream flow evaluations conducted for the California Department of Fish and Wildlife. These evaluations developed relationships between flow and suitable spawning and rearing habitat for Battle Creek from seven study reaches in Battle Creek. See appendix on pgs C-1 to C-9 of linked report for original data. [Mark Gard](mailto:mark_gard@fws.gov) from the U.S. Fish and Wildlife Service consolidated the reach-scale date to the entire Battle Creek watershed, provided the consolidated data in a spreadsheet (see '~/cvpiaHabitat/data-raw/mark_gard_data/IFIMWUA.xlsx'), and instructed us to use A1:E36 of the 'Battle' tab within his 'IFIMWUA.xlsx'. The units for WUA in the raw data are in square feet per 1000 feet, and adult trout data are used for steelhead.


### Combine Spawning and Rearing Data
The instream spawning and rearing habitat data described above for Fall Run Chinook Salmon (FR) and Steelhead (ST) is combined for use in the DSM in the following format. Units are in square feet per 1000 feet. 


```{r}
battle <- readxl::read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx',
                           range = "A1:E36", sheet = 'Battle')
battle_creek_instream <- battle %>% 
  select(flow_cfs = Flow, FR_spawn_wua = Spawning,
         FR_fry_wua = `Fry Rearing`, FR_juv_wua = `Juv Rearing`,
         ST_adult_wua = `adult trout`) %>%
  mutate(watershed = 'Battle Creek')

knitr:: kable(head(battle_creek_instream, 5), align = 'c') 
# devtools::use_data(battle_creek_instream, overwrite = TRUE)
```

###### ...with 35 more rows. 

### Spawning WUA 
The instream spawning habitat weighted usable areas for Fall Run Chinook Salmon in Battle Creek.

#### Spawning WUA Plot
The following plot shows the weighted usable spawning area in square feet per thousand feet of river for Fall Run Chinook Salmon. These area per length rates are multiplied by the total spawning reach length mapped by the SIT. 

``` {r,echo=FALSE}
battle_creek_instream %>% 
    filter(!is.na(FR_spawn_wua)) %>%
    ggplot(aes(x = flow_cfs , y = FR_spawn_wua)) +
    geom_line(color='#7570b3') +
    labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') +
  theme_minimal()

```

### Rearing WUA
The fry and juvenile instream rearing habitat weighted usable areas for Fall Run Chinook Salmon in Battle Creek. 

#### Rearing WUA Plot

The following plot shows the weighted usable rearing area in square feet per thousand feet of river for Fall Run Chinook Salmon fry and juvenile. These rates are multiplied by the total rearing reach length mapped by the SIT. 
``` {r,echo=FALSE}
battle_creek_instream %>% 
  gather(Lifestage, wua, -flow_cfs, -watershed)  %>% 
  filter(!is.na(wua), Lifestage != 'FR_spawn_wua', Lifestage != 'ST_adult_wua') %>%
  mutate(Lifestage = ifelse(Lifestage == 'FR_fry_wua', 'Fry', 'Juvenile')) %>% 
  ggplot(aes(x = flow_cfs , y = wua, color = Lifestage)) +
  geom_line() +
  labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') + 
  theme_minimal() + 
  scale_color_manual(values = c('#d95f02','#7570b3'))
```


