---
title: "Battle Creek"
author: '[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com)'
date: "December 13, 2018"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, cache=TRUE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "../../..")
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
```
## Future Data Improvements
[Future data improvement bullets]

## Instream Spawning and Rearing Habiat

**Data Source:**   
[Payne 1995](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/Payne1995_BattleCreekIFIM.pdf) 

Instream spawning and rearing habitat for [Fall Run Chinook Salmon and Steelhead] in the Battle Creek is based on data from [California Department of Water Resources and Thomas R. Payne & Associates instream flow evaluations for the relicensing of Oroville facilities.These evaluations developed relationships between flow and suitable spawning and rearing habitat for 23.25 miles of the Feather River between the Fish Barrier Dam and Honcut Creek, consisting of two river segments. The DWR FERC instream spawning and rearing habitat data] was provided by [Mark Gard](mailto:mark_gard@fws.gov) from the U.S. Fish and Wildlife Service in a spreadsheet (see '~/cvpiaHabitat/data-raw/mark_gard_data/IFIMWUA.xlsx').Mark Gard instructed us to use A1:E36 of the 'Battle' tab within his 'IFIMWUA.xlsx'.

Units are in square feet per 1000 feet.

Adult trout are used for steelhead. 


```{r,include=TRUE}
knitr::opts_knit$set(root.dir = "../../..")
getwd()
# battle <- readxl::read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx',
#                            range = "A1:E36", sheet = 'Battle')
# battle_creek_instream <- battle %>%
#   select(flow_cfs = Flow, FR_spawn_wua = Spawning,
#          FR_fry_wua = `Fry Rearing`, FR_juv_wua = `Juv Rearing`,
#          ST_adult_wua = `adult trout`) %>%
#   mutate(watershed = 'Battle Creek')
# 
# battle_creek_instream
# devtools::use_data(battle_creek_instream, overwrite = TRUE)
```

### Spawning WUA 
[description]

#### Spawning WUA Plot
The following plot shows the weighted usable spawning area in square feet per thousand feet of river for Fall Run Chinook Salmon. These area per length rates are multiplied by the total spawning reach length mapped by the SIT and applied for the flow in each month in the DSM.

``` {r,echo=FALSE}
battle_creek_instream %>% 
    filter(!is.na(FR_spawn_wua)) %>%
    ggplot(aes(x = flow_cfs , y = FR_spawn_wua)) +
    geom_line() +
    labs(x = 'flow (cfs)', y = 'WUA (sqft/1000ft)')

```

### Rearing WUA
The fry and juvenile instream rearing habitat weighted usable areas for Fall Run Chinook Salmon in the Feather River provided by Mark Gard (see '~/cvpiaHabitat/data-raw/mark_gard_data/IFIMWUA.xlsx') are used directly in the DSM. 

#### Rearing WUA Plot

The following plot shows the weighted usable rearing area in square feet per thousand feet of river for Fall Run Chinook Salmon fry and juvenile. These rates are multiplied by the total rearing reach length mapped by the SIT and applied for the flow in each month in the DSM.
``` {r,echo=FALSE}
battle_creek_instream %>% 
  gather(lifestage, wua, -flow_cfs, -watershed)  %>% 
  filter(!is.na(wua), lifestage != 'FR_spawn_wua', lifestage != 'ST_adult_wua') %>%
  mutate(lifestage = ifelse(lifestage == 'FR_fry_wua', 'fry', 'juvenile')) %>% 
  ggplot(aes(x = flow_cfs , y = wua, color = lifestage)) +
  geom_line() +
  labs(x = 'flow (cfs)', y = 'WUA (sqft/1000ft)')
```

### Combine Spawning and Rearing Data
The instream spawning and rearing habitat data described above for Fall Run Chinook Salmon (FR) is for use in the DSM in the following format.
```{r}
battle_creek_instream
```

