---
title: "Battle Creek"
author: "[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com)"
date: "April 24, 2019"
output: html_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
```
## Future Data Improvements
None

## Instream Spawning and Rearing Habiat

**Data Sources:**   
[Payne 1995](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/Payne1995_BattleCreekIFIM.pdf) and <https://flowwest.github.io/cvpiaHabitat/reference/battle_creek_instream.html>


Instream spawning and rearing habitat for Fall Run Chinook Salmon and Steelhead in Battle Creek is based on data from Thomas R. Payne & Associates instream flow evaluations conducted for the California Department of Fish and Wildlife. These evaluations developed relationships between flow and suitable spawning and rearing habitat for Battle Creek from seven study reaches in Battle Creek. [Mark Gard](mailto:mark_gard@fws.gov) from the U.S. Fish and Wildlife Service consolidated the reach-scale date to the entire Battle Creek watershed, provided the consolidated data in a spreadsheet (see '~/cvpiaHabitat/data-raw/mark_gard_data/IFIMWUA.xlsx'), and instructed us to use A1:E36 of the 'Battle' tab within his 'IFIMWUA.xlsx'. The units for WUA in the raw data are in square feet per 1000 feet, and adult trout data are used for steelhead.

### Spawning WUA
The following plot shows the weighted usable spawning area rate in square feet per thousand feet of river for Fall Run Chinook Salmon (FR_spawn_wua) and Steelhead (ST_spawn_wua). These rates are multiplied by the total spawning reach length mapped by the SIT and applied for the flow in each month in the DSM.

ADD SPAWNING GRAPH HERE

### Rearing WUA
The following plot shows the weighted usable rearing area rate in square feet per thousand feet of river for Fall Run Chinook Salmon (FR_spawn_wua) and Steelhead (ST_spawn_wua) fry and juveniles. These rates are multiplied by the total rearing reach length mapped by the SIT and applied for the flow in each month in the DSM. 

ADD REARING GRAPH HERE

### Combine Spawning and Rearing Data
The instream spawning and rearing habitat data described above for Fall Run Chinook Salmon (FR) and Steelhead (ST) is combined for use in the DSM in the following format.

```{r,echo=FALSE}
battle <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', 
                           range = "A1:E36", sheet = 'Battle')

battle_creek_instream <- battle %>% 
  select(flow_cfs = Flow, FR_spawn_wua = Spawning,
         FR_fry_wua = `Fry Rearing`, FR_juv_wua = `Juv Rearing`,
         ST_adult_wua = `adult trout`) %>% 
  mutate(watershed = 'Battle Creek')

battle_creek_instream

# devtools::use_data(battle_creek_instream, overwrite = TRUE)
```

