---
title: "Bear Creek"
author: "[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com), [Erin Cain](mailto:ecain@flowwest.com)"
date: "July 17, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
```

## Future Data Improvements
None Planned 

## Instream Spawning and Rearing Habiat
```{r}
source('data-raw/watershed/mean_wua_description.R')
```
**Data Source:** 
`r print_mean_wua_details('Bear Creek')`

### Combine Spawning and Rearing Data 
The instream spawning and rearing habitat data described above for Fall Run Chinook Salmon (FR) is combined for use in the DSM in the following format.

```{r}
mean_wua <- upper_mid_sac_region_instream %>% 
  mutate(watershed = 'Bear Creek')
knitr::kable(align = 'c', head(mean_wua, 5))
```

*...with 18 more rows.*

### Spawning WUA 
The instream spawning habitat weighted usable areas for Fall Run Chinook Salmon in Bear Creek.

#### Spawning WUA Plot
The following plot shows the weighted usable spawning area in square feet per thousand feet of river for Fall Run Chinook Salmon. These area per length rates are multiplied by the total spawning reach length mapped by the SIT. 

```{r, echo=FALSE}
mean_wua %>%
  filter(!is.na(FR_spawn_wua)) %>%
  ggplot(aes(flow_cfs, FR_spawn_wua)) +
  geom_line(color = '#7570b3' ) +
  labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') +
  theme_minimal() 
```

### Rearing WUA
The fry and juvenile instream rearing habitat weighted usable areas for Fall Run Chinook Salmon in Bear Creek. 

#### Rearing WUA Plot

The following plot shows the weighted usable rearing area in square feet per thousand feet of river for Fall Run Chinook Salmon fry and juvenile. These rates are multiplied by the total rearing reach length mapped by the SIT. 

``` {r,echo=FALSE}
mean_wua %>% 
  gather(lifestage, wua, -flow_cfs, -watershed)  %>% 
  filter(!is.na(wua), lifestage != 'FR_spawn_wua', lifestage != "ST_fry_wua", lifestage != "ST_juv_wua") %>%
  mutate(lifestage = ifelse(lifestage == 'FR_fry_wua', 'fry', 'juvenile')) %>% 
  ggplot(aes(x = flow_cfs , y = wua, color = lifestage)) +
  geom_line() +
  labs(x = 'flow (cfs)', y = 'WUA (sqft/1000ft)') + theme_minimal() +
  scale_color_manual(values = c('#d95f02','#7570b3'))
```

```{r}
source('data-raw/floodplain/floodplain_utils.R')
```

## Floodplain Modeling Details

**Data Source:**

[Scaled from a Deer Creek flow to floodplain area relationship generated with a 1D HEC-RAS hydraulic model](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/DRAFT+Deer+Creek+Hydraulic+Models+Tech+Memo+6-08-07.pdf)

`r print_model_details('Bear Creek', 'fr')`

## Floodplain Data

```{r}
# TODO Ask mark about the difference in outcome from his sheet
bear_creek_floodplain <- scale_fp_flow_area('Bear Creek')

knitr::kable(align = 'c', head(bear_creek_floodplain, 5))
# devtools::use_data(bear_creek_floodplain, overwrite = TRUE)
```

*...with 17 more rows*

## Floodplain Plot
```{r}
bear_creek_floodplain %>% 
  select(FR_floodplain_acres, watershed, flow_cfs) %>%
  rename(`Fall Run Chinook and Steelhead` = FR_floodplain_acres) %>% 
  gather(Species, acres, -flow_cfs, -watershed) %>% 
  ggplot(aes(flow_cfs, acres, color = Species)) +
  geom_line() +
  theme_minimal() +
  scale_color_brewer(palette = 'Dark2') +
  labs(x = 'Flow (cfs)', y = 'Total Inundated Acres')
```
