---
title: "Pools"
author: "[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com), [Erin Cain](mailto:ecain@flowwest.com)"
date: "August 20, 2019"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, out.width = '100%')
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
```

## Rearing Pools

Estimates of percent pools were provided by [Mark Gard](mark_gard@@fws.gov) for some watersheds.The mean percent pools, excluding Feather River, is 23.1%. This value is used for watersheds without an estimate.The total area of pools is then caluculated by multiplying the estimated percent pools by the total channel area. Channel areas were estimated using Google Earth Engine.

### Watersheds With Estimated Percent Pools 
```{r}
lengths_table <- read_excel('data-raw/watershed/CVPIA_FloodplainAreas.xlsx',
                            sheet = 'MetaData',
                            col_types = c('text', 'text', 'text', 'text', rep('numeric', 17),
                                          'text', 'numeric', 'text'),
                            na = 'na')

pools_perc <- read_csv('data-raw/watershed/pools/pools.csv', col_types = 'cnc') %>%
  select(-FW_QAQC)

mean_pools_perc <- filter(pools_perc, watershed != 'Feather River') %>%
  pull(percent_pools) %>% mean(na.rm = TRUE)

knitr::kable(align = 'c', pools_perc)

wss <- cvpiaData::watershed_ordering$watershed

pools <- lengths_table %>%
  filter(watershed %in% wss) %>%
  mutate(SR_prop_length = SR_rearing_length_mi/FR_rearing_length_mi,
         ST_prop_length = ST_rearing_length_mi/FR_rearing_length_mi,
         FR_channel_area = FR_channel_area_of_length_modeled_acres,
         SR_channel_area = FR_channel_area * SR_prop_length,
         ST_channel_area = FR_channel_area * ST_prop_length) %>%
  select(watershed, FR_channel_area, SR_channel_area, ST_channel_area) %>%
  left_join(pools_perc) %>%
  mutate(percent_pools = replace(percent_pools, is.na(percent_pools), mean_pools_perc),
         SR_pools_sq_meters = SR_channel_area * percent_pools/100 * 4046.86,
         ST_pools_sq_meters = ST_channel_area * percent_pools/100 * 4046.86) %>%
  select(watershed, SR_pools_sq_meters, ST_pools_sq_meters) %>%
  left_join(cvpiaData::watershed_ordering) %>%
  arrange(order) %>%
  select(-order)

# put combinded values of upper sac + upper mid sac in upper sac
# upper sac + upper mid sac 6174 acres, from mark gard

pools_upper_sac <- 6174 * 7.2/100 * 4046.86

pools[pools$watershed == 'Upper Sacramento River', c('SR_pools_sq_meters', 'ST_pools_sq_meters')] <- c(pools_upper_sac, pools_upper_sac)

```

### Spring Run Chinook and Steelhead Pools 
```{r}
# use_data(pools, overwrite = TRUE)
knitr::kable(align = 'c', head(pools, 5), caption = 'SR_pools_sq_meters = Spring Run Chinook Pools in square meters, ST_pools_sq_meters = Steelhead Pools in square meters')
```

*...with 26 more rows.*
