---
title: "Cottonwood Creek"
author: "[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com), [Erin Cain](ecain@flowwest.com)"
date: "June 11, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
```
## Future Data Improvements
none 

## Instream Spawning and Rearing Habitat

**Data Source:**
[CDFW 1979 and FWS 2014](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/CDFW1979_CottonwoodSpawningIFIM.PDF) (pgs 11-15)

Instream spawning and rearing habitat for Fall Run Chinook Salmon in the Cottonwood Creek is based on data from a collaborative report by the U.S. Corps of Engineers and the California Department of Fish and Game. The traditional Department of Fish and Game flow method was used to measure spawning habitat and the federal instream flow method was used to measure rearing habitat at 3 reaches making up a 10.76 mile stretch of Cottonwood Creek. The spawning and rearing data is provided on pages 11-15 of the CDFW 1979 and FWS 2014 report. The instream spawning and rearing habitat data was provided to us by [Mark Gard](mailto:mark_gard@fws.gov) from the U.S. Fish and Wildlife Service in a spreadsheet (see '~/cvpiaHabitat/data-raw/mark_gard_data/IFIMWUA.xlsx'). Mark Gard instructed us to use A2:B8 for spawning, and A10:C40 for rearing of the 'Cottonwood' tab within his 'IFIMWUA.xlsx'. 


##Combine Spawning and Rearing Data 
The instream spawning and rearing habitat data described above for Fall Run Chinook Salmon (FR) is combined for use in the DSM in the following format. Units are in square feet per 1000 feet.


```{r,echo=FALSE}
mi <- 10.76
cotton_spawn <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', 
                            range = "A2:B8", sheet = 'Cottonwood') %>% 
  mutate(FR_spawn_wua = Spawning/mi/5.28) %>% 
   select(flow_cfs = Flow, FR_spawn_wua)
 
cotton_rear <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', 
                           range = "A10:C40", sheet = 'Cottonwood') %>% 
   mutate(FR_fry_wua = `Fry Rearing`/mi/5.28, FR_juv_wua = `Juv Rearing`/mi/5.28) %>% 
   select(flow_cfs = Flow, FR_fry_wua, FR_juv_wua)
 
cottonwood_creek_instream <- cotton_spawn %>% 
   full_join(cotton_rear) %>% 
  arrange(flow_cfs) %>% 
   mutate(watershed = 'Cottonwood Creek')

knitr::kable(align = 'c', head(cottonwood_creek_instream, 5))
# devtools::use_data(cottonwood_creek_instream, overwrite = TRUE)
```

###### ...with 31 more rows.

### Spawning WUA 
The instream spawning habitat weighted usable areas for Fall Run Chinook Salmon in Cottonwood Creek.

#### Spawning WUA Plot
The following plot shows the weighted usable spawning area in square feet per thousand feet of river for Fall Run Chinook Salmon. These area per length rates are multiplied by the total spawning reach length mapped by the SIT. 

``` {r,echo=FALSE}
cottonwood_creek_instream %>% 
    filter(!is.na(FR_spawn_wua)) %>%
    ggplot(aes(x = flow_cfs , y = FR_spawn_wua)) +
    geom_line(color = '#7570b3') +
    labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') + 
    theme_minimal()
```

### Rearing WUA
The fry and juvenile instream rearing habitat weighted usable areas for Fall Run Chinook Salmon in Cottonwood Creek.

#### Rearing WUA Plot

The following plot shows the weighted usable rearing area in square feet per thousand feet of river for Fall Run Chinook Salmon fry and juvenile. These rates are multiplied by the total rearing reach length mapped by the SIT. 
``` {r,echo=FALSE}
cottonwood_creek_instream %>% 
  gather(Lifestage, wua, -flow_cfs, -watershed)  %>% 
  filter(!is.na(wua), Lifestage != 'FR_spawn_wua') %>%
  mutate(Lifestage = ifelse(Lifestage == 'FR_fry_wua', 'Fry', 'Juvenile')) %>% 
  ggplot(aes(x = flow_cfs , y = wua, color = Lifestage)) +
  geom_line() +
  labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') + 
  theme_minimal() + 
  scale_color_manual(values = c('#d95f02','#7570b3'))
```

