---
title: "Cottonwood Creek"
author: "[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com)"
date: "December 13, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
```
## Future Data Improvements
[Future data improvements bullets]

## Instream Spawning and Rearing Habiat

**Data Source:**
[CDFW 1979 and FWS 2014](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/CDFW1979_CottonwoodSpawningIFIM.PDF)

Instream spawning and rearing habitat for Fall Run Chinook Salmon in the Cottonwood Creek is based on data from [data source description]. The [DWR FERC instream spawning and rearing habitat data] was provided by [Mark Gard](mailto:mark_gard@fws.gov) from the U.S. Fish and Wildlife Service in a spreadsheet (see '~/cvpiaHabitat/data-raw/mark_gard_data/IFIMWUA.xlsx'). Mark Gard instructed us to use A2:B8 for spawning, and A10:C40 for rearing of the 'Cottonwood' tab within his 'IFIMWUA.xlsx'. 

Units are in square feet and study length is 10.76 miles.

```{r,include=FALSE}
# mi <- 10.76
# cotton_spawn <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', 
#                            range = "A2:B8", sheet = 'Cottonwood') %>% 
#   mutate(FR_spawn_wua = Spawning/mi/5.28) %>% 
#   select(flow_cfs = Flow, FR_spawn_wua)
# 
# cotton_rear <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', 
#                           range = "A10:C40", sheet = 'Cottonwood') %>% 
#   mutate(FR_fry_wua = `Fry Rearing`/mi/5.28, FR_juv_wua = `Juv Rearing`/mi/5.28) %>% 
#   select(flow_cfs = Flow, FR_fry_wua, FR_juv_wua)
# 
# cottonwood_creek_instream <- cotton_spawn %>% 
#   full_join(cotton_rear) %>% 
#   arrange(flow_cfs) %>% 
#   mutate(watershed = 'Cottonwood Creek')
# 
# cottonwood_creek_instream
# devtools::use_data(cottonwood_creek_instream, overwrite = TRUE)
```

### Spawning WUA 
[description]

#### Spawning WUA Plot
The following plot shows the weighted usable spawning area in square feet per thousand feet of river for Fall Run Chinook Salmon. These area per length rates are multiplied by the total spawning reach length mapped by the SIT and applied for the flow in each month in the DSM.

``` {r,echo=FALSE}
cottonwood_creek_instream %>% 
    filter(!is.na(FR_spawn_wua)) %>%
    ggplot(aes(x = flow_cfs , y = FR_spawn_wua)) +
    geom_line() +
    labs(x = 'flow (cfs)', y = 'WUA (sqft/1000ft)')
```

### Rearing WUA
The fry and juvenile instream rearing habitat weighted usable areas for Fall Run Chinook Salmon in the Cottonwood River provided by Mark Gard (see '~/cvpiaHabitat/data-raw/mark_gard_data/IFIMWUA.xlsx') are used directly in the DSM. 

#### Rearing WUA Plot

The following plot shows the weighted usable rearing area in square feet per thousand feet of river for Fall Run Chinook Salmon fry and juvenile. These rates are multiplied by the total rearing reach length mapped by the SIT and applied for the flow in each month in the DSM.
``` {r,echo=FALSE}
cottonwood_creek_instream %>% 
  gather(lifestage, wua, -flow_cfs, -watershed)  %>% 
  filter(!is.na(wua), lifestage != 'FR_spawn_wua', lifestage != 'ST_adult_wua') %>%
  mutate(lifestage = ifelse(lifestage == 'FR_fry_wua', 'fry', 'juvenile')) %>% 
  ggplot(aes(x = flow_cfs , y = wua, color = lifestage)) +
  geom_line() +
  labs(x = 'flow (cfs)', y = 'WUA (sqft/1000ft)')
```

### Combine Spawning and Rearing Data
The instream spawning and rearing habitat data described above for Fall Run Chinook Salmon (FR) is for use in the DSM in the following format.
```{r,echo=FALSE}
cottonwood_creek_instream
```
