---
title: "Tuolumne River"
author: "[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com)"
date: "April 10, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
```

## Future Data Improvements
none 

## Instream Spawning and Rearing Habiat

**Data Source:**
[Flo2D model by Stillwater 2014](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/Stillwater2010_Tuolumne_P-2299-075_71_DP_FLA_AttC_StudyRept_IFIM_AppA-H_140428.pdf)

Instream spawning and rearing habitat for Fall Run Chinook Salmon and Steelhead in the Tuolumne River is based on data from a California Department of Fish and Game site specific study. This data consists of 10 sites spread over 9.2 miles of the dominant spawning reach of the Lower Tuolumne River and was reviewed by Stillwater Sciences for use in their 2014 Flo2D model. The instream spawning and rearing habitat data was provided by [Mark Gard](mailto:mark_gard@fws.gov) from the U.S. Fish and Wildlife Service in a spreadsheet (see '~/cvpiaHabitat/data-raw/mark_gard_data/IFIMWUA.xlsx'). Mark Gard instructed us to use cells A2:H32 in the 'Tuolumne' tab within his 'IFIMWUA.xlsx'.


###Combine Spawning and Rearing Data 
The instream spawning and rearing habitat data described above for Fall Run Chinook Salmon (FR) and Steelhead (ST) is combined for use in the DSM in the following format.


```{r, echo=FALSE}
tuolumne_river_instream <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx',
                                      range = "A2:H32", sheet = 'Tuolumne') %>%
  mutate(watershed = 'Tuolumne River') %>%
  select(flow_cfs = `Simulated\nDischarge\n(cfs)`, FR_spawn_wua = `Chinook\nSpawning`,
         FR_fry_wua = `Chinook\nFry`, FR_juv_wua = `Chinook\nJuvenile`,
         ST_spawn_wua = `O. mykiss\nSpawning`, ST_fry_wua = `O. mykiss\nFry`,
         ST_juv_wua = `O. mykiss\nJuvenile`, ST_adult_wua = `O. mykiss\nAdult`, watershed)
tuolumne_river_instream #%>% 
  #print(n= 10, width = Inf)
#Do we need full tibble in documentation?
```

### Spawning WUA 
The spawning habitat weighted usable areas for Fall Run Chinook Salmon in the Tuolumne River provided by Mark Gard (see ‘~/cvpiaHabitat/data-raw/mark_gard_data/IFIMWUA.xlsx’).

#### Spawning WUA Plot
The following plot shows the weighted usable spawning area in square feet per thousand feet of river for Fall Run Chinook Salmon. These area per length rates are multiplied by the total spawning reach length mapped by the SIT.

``` {r,echo=FALSE}
tuolumne_river_instream %>% 
    ggplot(aes(x = flow_cfs , y = FR_spawn_wua)) +
    geom_line(color='#7570b3') +
    labs(x = 'flow (cfs)', y = 'WUA (sqft/1000ft)') + theme_minimal() 

```

### Rearing WUA
The fry and juvenile instream rearing habitat weighted usable areas for Fall Run Chinook Salmon and Steelhead in the Tuolumne River provided by Mark Gard (see '~/cvpiaHabitat/data-raw/mark_gard_data/IFIMWUA.xlsx') are used directly in the DSM. 

#### Chinook Salmon Rearing WUA Plot

The following plot shows the weighted usable rearing area in square feet per thousand feet of river for Fall Run Chinook Salmon fry and juvenile. These rates are multiplied by the total rearing reach length mapped by the SIT.
``` {r,echo=FALSE}
tuolumne_river_instream %>% 
  gather(lifestage, wua, -flow_cfs, -watershed)  %>% 
  filter(!is.na(wua), lifestage != 'FR_spawn_wua', lifestage != 'ST_adult_wua', lifestage != "ST_spawn_wua", lifestage != "ST_juv_wua", lifestage != "ST_fry_wua") %>%
  mutate(lifestage = ifelse(lifestage == 'FR_fry_wua', 'fry', 'juvenile')) %>% 
  ggplot(aes(x = flow_cfs , y = wua, color = lifestage)) +
  geom_line() +
  labs(x = 'flow (cfs)', y = 'WUA (sqft/1000ft)') + theme_minimal() +
  scale_color_manual(values = c('#d95f02','dodgerblue4'))
```

#### Steelhead Rearing WUA Plot

The following plot shows the weighted usable rearing area in square feet per thousand feet of river for Steelhead fry and juvenile. These rates are multiplied by the total rearing reach length mapped by the SIT.

``` {r,echo=FALSE}
tuolumne_river_instream %>% 
  gather(lifestage, wua, -flow_cfs, -watershed)  %>% 
  filter(!is.na(wua), lifestage != 'FR_spawn_wua', lifestage != 'ST_adult_wua',  lifestage != "FR_juv_wua", lifestage != "FR_fry_wua", lifestage != "ST_spawn_wua") %>%
  mutate(lifestage = ifelse(lifestage == 'ST_fry_wua', 'fry', 'juvenile')) %>% 
  ggplot(aes(x = flow_cfs , y = wua, color = lifestage)) +
  geom_line() +
  labs(x = 'flow (cfs)', y = 'WUA (sqft/1000ft)') + theme_minimal() +
  scale_color_manual(values = c('#d95f02','dodgerblue4'))
```
