---
title: "Feather River"
author: "sadie"
date: "December 12, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
library(scales)
```

## Instream Spawning and Rearing Habiat

> **Data Source:** The spawning and rearing WUA values were derived from a PHABSIM model.
> <https://flowwest.github.io/cvpiaHabitat/reference/feather_river_instream.html>

Mark Gard instructed us to use A1:D8 of the 'Feather' tab within his 'IFIMWUA.xlsx'

Flow in cfs, WUA in square feet per 1000 feet

#### Mark Gard Raw Data
```{r, echo=FALSE}
feather_raw <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', range = "A1:D8", sheet = 'Feather')
feather_raw
```

## Spawning Data
We went to the FERC 2004 report and hand digitized the spawning relationship using <https://automeris.io/WebPlotDigitizer/>

```{r, echo=FALSE}
upper <- read_csv('data-raw/instream/feather/alt_upper_feather_spawn_wua_rsi.csv') %>%
  mutate(location = 'upper')

lower <- read_csv('data-raw/instream/feather/alt_lower_feather_spawn_wua_rsi.csv') %>%
  mutate(location = 'lower')

upper_approx <- approxfun(upper$flow, upper$wua_rsi, rule = 2)
lower_approx <- approxfun(lower$flow, lower$wua_rsi, rule = 2)

# modeled lengths????
upper_length <- 8.25 * 5280
lower_length <- 15 * 5280

spawning <- tibble(
  flow_cfs = c(200, 500, 750, 1000, 1250, 1500, 1750, 2000, 2500, 3000, 3500, 4000, 4500, 5000, 5500, 6000, 6500, 7000),
  upper_wua = upper_approx(flow_cfs),
  lower_wua = lower_approx(flow_cfs)) %>%
  mutate(FR_spawn_wua = upper_wua * (9.75/18) + lower_wua * (8.25/18)) %>%
  select(flow_cfs, FR_spawn_wua)

write_csv(spawning, 'data-raw/instream/feather/feather_river_spawning.csv')

cvpiaHabitat::watershed_lengths %>%
  filter(watershed == 'Feather River')

# 18 miles for our mapped spawning
(8.25 + 15) - 18
18 - 8.25
med_spawn_flow <- 2241.379

((upper_approx(med_spawn_flow) * 8.5*5280/1000) + (lower_approx(med_spawn_flow) * 9.75*5280/1000))/43560
wua_lower <- 32796.83
wua_upper <- 7277.629
(wua_lower * (8.25/18) + wua_upper * (9.75/18))* 18*5280/1000/43560
```

### Spawning WUA

```{r, echo=FALSE}
spawning %>% 
  ggplot(aes(flow_cfs, FR_spawn_wua)) +
  geom_point() +
  labs(x = 'flow (cfs)', y = 'spawning WUA (sqft/1000ft)') +
  scale_y_continuous(labels = comma)
```

### Rearing WUA from raw data
```{r, echo=FALSE}
feather_raw %>%
  filter(!is.na(`Fry Rearing`)) %>%
  select(-Spawning) %>%
  gather(species, wua, - Flow) %>%
  ggplot(aes(Flow, wua, color = species)) +
  geom_line() +
  labs(x = 'flow (cfs)', y = 'spawning WUA (sqft/1000ft)') +
  scale_y_continuous(labels = comma)
```

### Rearing Habitat in Acres
Using the habitat package `cvpiaHabitat::set_instream_habitat`

```{r, echo=FALSE}
rear_calc <- tibble(
  Flow = seq(100, 3000, length.out = 100),
  fr_fry = cvpiaHabitat::square_meters_to_acres(
    cvpiaHabitat::set_instream_habitat('Feather River', 'fr', 'fry', Flow)),
  fr_juv = cvpiaHabitat::square_meters_to_acres(
    cvpiaHabitat::set_instream_habitat('Feather River', 'fr', 'juv', Flow))) %>%
  gather(species, acres, -Flow)

rear_calc %>%
  ggplot(aes(Flow, acres, color = species)) +
  geom_line()

```

```{r}
feather_river_instream <- feather_raw %>% 
  select(-Spawning) %>% 
  rename(flow_cfs = Flow, FR_fry_wua = `Fry Rearing`, FR_juv_wua = `Juv Rearing`) %>% 
  full_join(spawning) %>% 
  arrange(flow_cfs) %>% 
  mutate(watershed = 'Feather River') %>%
  select(flow_cfs,
         FR_spawn_wua,
         FR_fry_wua,
         FR_juv_wua,
         watershed)
  
# devtools::use_data(feather_river_instream, overwrite = TRUE)
```
