---
title: "Cosumnes River"
author: "[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com)"
date: "December 17, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
library(scales)

```
## Future Data Improvements
[Future data improvement bullets]

## Instream Spawning and Rearing Habiat

**Data Source:**   
[???--combo of mokelunme and calaveras]

There is no habitat modeling, using mokelumne and calaveras

Instream spawning and rearing habitat for Fall Run Chinook Salmon in the Cosumnes River is based on data from [data source description]. The [DWR FERC instream spawning and rearing habitat data] was provided by [Mark Gard](mailto:mark_gard@fws.gov) from the U.S. Fish and Wildlife Service in a spreadsheet (see '~/cvpiaHabitat/data-raw/mark_gard_data/IFIMWUA.xlsx').

```{r,include=FALSE}
# # cosumnes -----
# region <- cvpiaHabitat::modeling_exist %>%
#   filter(Region == 'South Delta') %>%
#   select(Watershed, Region, FR_fry, FR_juv)
# 
# flows <- seq(100, 1000, by = 100)
# 
# # spawn
# cal_st_spwn <- approx(calaveras_river_instream$flow_cfs, 
#                       calaveras_river_instream$ST_spawn_wua, 
#                       rule = 2, xout = flows)$y
# 
# mok_fr_spwn <- approx(mokelumne_river_instream$flow_cfs, 
#                       mokelumne_river_instream$FR_spawn_wua, 
#                       rule = 2, xout = flows)$y
# 
# cos_spwn <- (cal_st_spwn + mok_fr_spwn) / 2
# 
# # fry
# cal_st_fry <- approx(calaveras_river_instream$flow_cfs, 
#                       calaveras_river_instream$ST_fry_wua, 
#                       rule = 2, xout = flows)$y
# 
# mok_fr_fry <- approx(mokelumne_river_instream$flow_cfs, 
#                       mokelumne_river_instream$FR_fry_wua, 
#                       rule = 2, xout = flows)$y
# 
# cos_fry <- (cal_st_fry + mok_fr_fry) / 2
# 
# # juv
# cal_st_juv <- approx(calaveras_river_instream$flow_cfs, 
#                       calaveras_river_instream$ST_juv_wua, 
#                       rule = 2, xout = flows)$y
# 
# mok_fr_juv <- approx(mokelumne_river_instream$flow_cfs, 
#                       mokelumne_river_instream$FR_juv_wua, 
#                       rule = 2, xout = flows)$y
# 
# cos_juv <- (cal_st_juv + mok_fr_juv) / 2
# 
# 
# 
# cosumnes_river_instream <- tibble(flow_cfs = flows, FR_spawn_wua = cos_spwn, 
#                                   FR_fry_wua = cos_fry, FR_juv_wua = cos_juv,
#                                   watershed = 'Cosumnes River')
# 
# cosumnes_river_instream
# devtools::use_data(cosumnes_river_instream, overwrite = TRUE)
```

### Spawning WUA 
[description]

#### Spawning WUA Plot
The following plot shows the weighted usable spawning area in square feet per thousand feet of river for Fall Run Chinook Salmon. These area per length rates are multiplied by the total spawning reach length mapped by the SIT and applied for the flow in each month in the DSM.

``` {r,echo=FALSE}
cosumnes_river_instream %>% 
    filter(!is.na(FR_spawn_wua)) %>%
    ggplot(aes(x = flow_cfs , y = FR_spawn_wua)) +
    geom_line() +
    labs(x = 'flow (cfs)', y = 'WUA (sqft/1000ft)')
```

### Rearing WUA
The fry and juvenile instream rearing habitat weighted usable areas for Fall Run Chinook Salmon in the Consumnes River provided by Mark Gard (see '~/cvpiaHabitat/data-raw/mark_gard_data/IFIMWUA.xlsx') are used directly in the DSM. 

#### Rearing WUA Plot

The following plot shows the weighted usable rearing area in square feet per thousand feet of river for Fall Run Chinook Salmon fry and juvenile. These rates are multiplied by the total rearing reach length mapped by the SIT and applied for the flow in each month in the DSM.
``` {r,echo=FALSE}
cosumnes_river_instream %>% 
  gather(lifestage, wua, -flow_cfs, -watershed)  %>% 
  filter(!is.na(wua), lifestage != 'FR_spawn_wua', lifestage != 'ST_adult_wua') %>%
  mutate(lifestage = ifelse(lifestage == 'FR_fry_wua', 'fry', 'juvenile')) %>% 
  ggplot(aes(x = flow_cfs , y = wua, color = lifestage)) +
  geom_line() +
  labs(x = 'flow (cfs)', y = 'WUA (sqft/1000ft)')
```

### Combine Spawning and Rearing Data
The instream spawning and rearing habitat data described above for Fall Run Chinook Salmon (FR) is for use in the DSM in the following format.
```{r,echo=FALSE}
cosumnes_river_instream
```

