---
title: "American River"
author: "[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com), [Erin Cain](mailto:ecain@flowwest.com)"
date: "August 14, 2019"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, out.width = '100%')
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
library(scales)

```
## Future Data Improvements
None planned

## Instream Spawning and Rearing Habitat
**Data Source:** [ARWA 2017](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/ARWA-702.pdf){target="_blank"} (pgs. 91-107) and [USFWS 1985](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/AmericanRiver1985IFIM.PDF){target="_blank"} (pgs. 10-11) 

Instream spawning habitat for Fall Run Chinook Salmon and Steelhead in the American River is based on the spawning habitat suitability and availability analysis in the Lower American River Biological Rationale Development and Performance of the Modified Flow Standard (ARWA 2017). Instream juvenile rearing habitat for Fall Run Chinook Salmon is based on instream flow studies conducted by [Mark Gard](mailto:mark_gard@fws.gov) from the U.S. Fish and Wildlife Service. Both evaluations developed relationships between flow and suitable spawning and rearing habitat for the Lower American River. The values contained in cells A35:D46 of the 'American' spreadsheet tab are the sums of the values from the instream flow evaluation reaches called Sailor Bar, Ancil Hoffman, Watt, and H Street. These values were divided by four to get composite instream rearing weighted usable area values in square feet per 1000 feet of river.       
```{r, echo=FALSE}
american_raw <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', range = "A35:D46", sheet = 'American')

spawn <- read_csv('data-raw/watershed/american/american_river_spawning_wua.csv')
 
study_len <- 800+377+1553+522+3142+3103+1000+754+3442+1500
 
rearing <- american_raw %>% 
  select(-Spawning) %>% 
  gather(lifestage, WUA, -Flow) %>% 
  rename(flow_cfs = Flow) 
 
spawning <- spawn %>%
  mutate(FR_spawn_wua = FR_wua_ft2/(study_len/1000),
         ST_spawn_wua = ST_wua_ft2/(study_len/1000)) %>% 
  select(flow_cfs, FR_spawn_wua, ST_spawn_wua) %>% 
  gather(lifestage, WUA, -flow_cfs)
 
american_river_instream <- rearing %>% 
  bind_rows(spawning) %>% 
  spread(lifestage, WUA) %>% 
  select(flow_cfs, FR_spawn_wua, FR_fry_wua = `Fry Rearing`, 
         FR_juv_wua = `Juv Rearing`, ST_spawn_wua) %>% 
  mutate(watershed = 'American River')

# usethis::use_data(american_river_instream, overwrite = TRUE)
```

### Spawning Data 
```{r, echo = FALSE}
american_spawning <- spawn %>%
  mutate(FR_spawn_wua = FR_wua_ft2/(study_len/1000),
         ST_spawn_wua = ST_wua_ft2/(study_len/1000)) %>% 
  select(flow_cfs, FR_spawn_wua, ST_spawn_wua) %>% 
  mutate(watershed = 'American River')

knitr::kable(align = 'c', head(american_spawning, 5), caption = "Header Descriptions: flow_cfs = flow in cubic feet per second, FR_spawn_wua = Fall Run Chinook Spawning wua,  ST_spawn_wua = Steelhead Spawning wua")
```

*...with 40 more rows.*

### Rearing Data 
```{r, echo = FALSE}
american_rearing <- american_raw %>% 
  select(-Spawning) %>% 
  rename(flow_cfs = Flow, FR_fry_wua = 'Fry Rearing', FR_juv_wua = 'Juv Rearing') %>% 
  mutate(watershed = 'American River') 

knitr::kable(align = 'c', head(american_rearing, 5), caption = "Header Descriptions: flow_cfs = flow in cubic feet per second, FR_juv_wua = Fall Run Chinook Juvenile wua,  FR_fry_wua = Fall Run Chinook Fry wua")
  
```

*...with 6 more rows.*

### Spawning WUA 
The Fall Run Chinook Salmon and Steelhead spawning habitat data from [ARWA 2017](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/ARWA-702.pdf) is contained in Attachment B, Table B-4 on page 91 and Figure B-8 on page 92 (Fall Run Chinook Salmon) and Attachment B, Table B-9 and Figure B-26 on page 107 (Steelhead). 

The following plot shows the composite weighted usable spawning area in square feet per thousand feet of river for Fall Run Chinook Salmon and Steelhead developed in the AWRA evaluation by combining the results for all evaluation reaches. These area per length rates are multiplied by the total spawning reach length mapped by the SIT. 
```{r, echo=FALSE}
american_river_instream %>% 
  select(flow_cfs, `Fall Run Chinook` = FR_spawn_wua, `Steelhead` = ST_spawn_wua) %>% 
  gather(Species, WUA, -flow_cfs) %>% 
  filter(!is.na(WUA)) %>% 
  ggplot(aes(flow_cfs, WUA, color = Species)) +
  geom_line() +
  labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') + 
  theme_minimal() + 
  scale_color_manual(values = c('#d95f02','#7570b3')) +
  theme(legend.justification = c(1,0), legend.position = c(.9,.7))

```

### Rearing WUA
The following plot shows the weighted usable rearing area in square feet per thousand feet of river for Fall Run Chinook Salmon fry and juvenile. These rates are multiplied by the total rearing reach length mapped by the SIT. 
```{r, echo=FALSE}
american_river_instream %>% 
  select(flow_cfs, Fry = FR_fry_wua, Juvenile = FR_juv_wua) %>% 
  gather(Lifestage, WUA, -flow_cfs) %>% 
  filter(!is.na(WUA)) %>% 
  ggplot(aes(x = flow_cfs, y = WUA, color = Lifestage)) +
  geom_line() +
  labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') + 
  theme_minimal() + 
  scale_color_manual(values = c('#d95f02','#7570b3')) +
  theme(legend.justification = c(1,0), legend.position = c(.9,.7))

```
```{r, include=FALSE}
source('data-raw/watershed/floodplain_utils.R')
```

## Floodplain Rearing Habitat
**Data Source:**
[Central Valley Floodplain Evaluation and Delineation (CVFED) HEC-RAS hydraulic model](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/CombinedTM_IQAR_Final-FULL-REPORT_20140206.pdf){target="_blank"}

`r print_model_details('American River', 'fr')`

## Floodplain Data
```{r, echo=FALSE}
american_river <- read_excel('data-raw/watershed/CVPIA_FloodplainAreas.xlsx',
                                        sheet = 'AmericanRiver') 

threshold_flow <- american_river %>% 
  filter(modeled_floodplain_area_acres == 0) %>% 
  summarise(max_flow = max(flow_cfs)) %>% 
  pull(max_flow)  
  
american_river_floodplain <- american_river %>% 
  mutate(watershed = 'American River',
         SR_floodplain_acres = modeled_floodplain_area_acres,
         ST_floodplain_acres = modeled_floodplain_area_acres) %>%
  select(flow_cfs,
         FR_floodplain_acres = modeled_floodplain_area_acres,
         SR_floodplain_acres, ST_floodplain_acres,
         watershed) %>% 
  filter(flow_cfs >= threshold_flow)

knitr::kable(align = 'c', head(american_river_floodplain, 5), caption = "Header Descriptions: flow_cfs = flow in cubic feet per second, FR_floodplain_acres = Fall Run Chinook floodplain acres, SR_floodplain_acres = Spring Run Chinook floodplain acres, ST_floodplain_acres = Steelhead floodplain acres")
# usethis::use_data(american_river_floodplain, overwrite = TRUE)
```

*...with 29 more rows*

## Floodplain Plot
```{r, echo=FALSE}
american_river_floodplain %>% 
  select(FR_floodplain_acres, flow_cfs, watershed) %>%
  rename(`Fall Run Chinook,
Spring Run Chinook,
and Steelhead` = FR_floodplain_acres) %>% 
  gather(Species, acres, -flow_cfs, -watershed) %>% 
  ggplot(aes(flow_cfs, acres, color = Species)) +
  geom_line() +
  theme_minimal() +
  scale_color_brewer(palette = 'Dark2') +
  labs(x = 'Flow (cfs)', y = 'Total Inundated Acres') +
  theme(legend.justification = c(1,0), legend.position = c(1,.1))
```

