---
title: "American River"
author: "[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com)" 
date:  "July 2020" 
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, out.width = '100%')
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
library(scales)

```
## Future Data Improvements
None planned. 

## Instream Spawning and Rearing Habitat
**Data Source:** [ARWA 2017](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/ARWA-702.pdf){target="_blank"} (pgs. 91-107) and [USFWS 1985](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/AmericanRiver1985IFIM.PDF){target="_blank"} (pgs. 10-11) 

The instream spawning habitat for Fall Run Chinook Salmon and Steelhead in the American River is based on a spawning habitat suitability and availability analysis performed by the Sacramento Water Forum. The Water Forum conducted the analysis as part of the Lower American River Biological Rationale Development and Performance of the Modified Flow Standard (ARWA 2017). The instream juvenile rearing habitat for Fall Run Chinook Salmon is from instream flow studies conducted by [Mark Gard](mailto:mark_gard@fws.gov) from the U.S. Fish and Wildlife Service. These analyses uncovered relationships between flow and both suitable spawning habitat and suitable rearing habitat for the Lower American River. Cells A35:D46 in the 'American' tab of the IFIMWUA [spreadsheet](https://cvpiahabitat-r-package.s3-us-west-2.amazonaws.com/IFIMWUA.xlsx) contain sums of the instream flow evaluation values from reaches Sailor Bar, Ancil Hoffman, Watt, and H Street. These values were equally weighted to obtain composite instream rearing weighted usable area (WUA) values in square feet per 1000 feet of river.     
```{r, echo=FALSE}
american_raw <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', range = "A35:D46", sheet = 'American')

spawn <- read_csv('data-raw/watershed/american/data/american_river_spawning_wua.csv')
 
study_len <- 800+377+1553+522+3142+3103+1000+754+3442+1500
 
rearing <- american_raw %>% 
  select(-Spawning) %>% 
  gather(lifestage, WUA, -Flow) %>% 
  rename(flow_cfs = Flow) 
 
spawning <- spawn %>%
  mutate(FR_spawn_wua = FR_wua_ft2/(study_len/1000),
         ST_spawn_wua = ST_wua_ft2/(study_len/1000)) %>% 
  select(flow_cfs, FR_spawn_wua, ST_spawn_wua) %>% 
  gather(lifestage, WUA, -flow_cfs)
 
american_river_instream <- rearing %>% 
  bind_rows(spawning) %>% 
  spread(lifestage, WUA) %>% 
  select(flow_cfs, FR_spawn_wua, FR_fry_wua = `Fry Rearing`, 
         FR_juv_wua = `Juv Rearing`, ST_spawn_wua) %>% 
  mutate(watershed = 'American River')

# usethis::use_data(american_river_instream, overwrite = TRUE)
```

### Spawning Data 
```{r, echo = FALSE}
american_spawning <- spawn %>%
  mutate(FR_spawn_wua = FR_wua_ft2/(study_len/1000),
         ST_spawn_wua = ST_wua_ft2/(study_len/1000)) %>% 
  select(flow_cfs, FR_spawn_wua, ST_spawn_wua) %>% 
  mutate(watershed = 'American River')

knitr::kable(align = 'c', head(american_spawning, 5), 
             caption = "Header Descriptions: flow_cfs = flow in cubic feet per second,
             FR_spawn_wua = Fall Run Chinook Spawning WUA,  
             ST_spawn_wua = Steelhead Spawning WUA, 
             watershed = section of stream modeled for CVPIA SDM")
```

*...with 40 more rows.*

### Rearing Data 
```{r, echo = FALSE}
american_rearing <- american_raw %>% 
  select(-Spawning) %>% 
  rename(flow_cfs = Flow, FR_fry_wua = 'Fry Rearing', FR_juv_wua = 'Juv Rearing') %>% 
  mutate(watershed = 'American River') 

knitr::kable(align = 'c', head(american_rearing, 5), 
             caption = "Header Descriptions:
              flow_cfs = flow in cubic feet per second,
              FR_fry_wua = Fall Run Chinook Fry WUA,
              FR_juv_wua = Fall Run Chinook Juvenile WUA,
              watershed = section of stream modeled for CVPIA SDM")
  
```

*...with 6 more rows.*

### Spawning WUA 
The Fall Run Chinook Salmon and Steelhead spawning habitat data from [ARWA 2017](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/ARWA-702.pdf) is located in Attachment B. Fall Run Chinook Salmon data are in Table B-4 on page 91 and Figure B-8 on page 92. Steelhead data are in Table B-9 and Figure B-26 on page 107. 

The following plot shows the composite WUA in square feet per thousand feet for Fall Run Chinook Salmon and Steelhead. The WUA values combine data from all evaluation reaches. These area per length rates are multiplied by the total spawning reach length mapped by the SIT to estimate areas of habitat. 
```{r, echo=FALSE}
american_river_instream %>% 
  select(flow_cfs, `Fall Run Chinook` = FR_spawn_wua, `Steelhead` = ST_spawn_wua) %>% 
  gather(Species, WUA, -flow_cfs) %>% 
  filter(!is.na(WUA)) %>% 
  ggplot(aes(flow_cfs, WUA, color = Species)) +
  geom_line() +
  labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') + 
  theme_minimal() + 
  scale_color_manual(values = c('#d95f02','#7570b3')) +
  theme(legend.justification = c(1,0), legend.position = c(.9,.7))

```

### Rearing WUA
The following plot shows the rearing WUA in square feet per thousand feet for fry and juvenile Fall Run Chinook Salmon. These rates are multiplied by the total rearing reach length mapped by the SIT to estimate habitat area. 
```{r, echo=FALSE}
american_river_instream %>% 
  select(flow_cfs, Fry = FR_fry_wua, Juvenile = FR_juv_wua) %>% 
  gather(Lifestage, WUA, -flow_cfs) %>% 
  filter(!is.na(WUA)) %>% 
  ggplot(aes(x = flow_cfs, y = WUA, color = Lifestage)) +
  geom_line() +
  labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') + 
  theme_minimal() + 
  scale_color_manual(values = c('#d95f02','#7570b3')) +
  theme(legend.justification = c(1,0), legend.position = c(.9,.7))

```
```{r, include=FALSE}
source('data-raw/watershed/floodplain_utils.R')
```

## Floodplain Rearing Habitat
**Data Source:**
[Central Valley Floodplain Evaluation and Delineation (CVFED) HEC-RAS hydraulic model](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/CombinedTM_IQAR_Final-FULL-REPORT_20140206.pdf){target="_blank"}

`r print_model_details('American River', 'fr')`

## Floodplain Data    
The areas represent total inundated area in acres.    
```{r, echo=FALSE}
american_river <- read_excel('data-raw/watershed/CVPIA_FloodplainAreas.xlsx',
                                        sheet = 'AmericanRiver') 

threshold_flow <- american_river %>% 
  filter(modeled_floodplain_area_acres == 0) %>% 
  summarise(max_flow = max(flow_cfs)) %>% 
  pull(max_flow)  
  
american_river_floodplain <- american_river %>% 
  mutate(watershed = 'American River',
         ST_floodplain_acres = modeled_floodplain_area_acres) %>%
  select(flow_cfs,
         FR_floodplain_acres = modeled_floodplain_area_acres,
         ST_floodplain_acres,
         watershed) %>% 
  filter(flow_cfs >= threshold_flow)

knitr::kable(align = 'c', head(american_river_floodplain, 5), 
             caption = "Header Descriptions: flow_cfs = flow in cubic feet per second,
             FR_floodplain_acres = Fall Run Chinook floodplain acres, 
             ST_floodplain_acres = Steelhead floodplain acres,
             watershed = section of stream modeled for CVPIA SDM")
# usethis::use_data(american_river_floodplain, overwrite = TRUE)
```

*...with 29 more rows*

## Floodplain Plot
```{r, echo=FALSE}
american_river_floodplain %>% 
  select(FR_floodplain_acres, flow_cfs, watershed) %>%
  rename(`Fall Run Chinook and Steelhead` = FR_floodplain_acres) %>% 
  gather(Species, acres, -flow_cfs, -watershed) %>% 
  ggplot(aes(flow_cfs, acres, color = Species)) +
  geom_line() +
  theme_minimal() +
  scale_color_brewer(palette = 'Dark2') +
  labs(x = 'Flow (cfs)', y = 'Total Inundated Acres') +
  theme(legend.justification = c(1,0), legend.position = c(1,.1))
```

