---
title: "Bear River"
author: "[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com)" 
date: "July 2020"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, out.width = '100%')
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
```
## Future Data Improvements
None planned. 

## Instream Spawning and Rearing Habitat

**Data Source:** [SSWD 2019](https://cvpiahabitat-r-package.s3-us-west-2.amazonaws.com/cvpia-sit-model-inputs/SSWD_Bear_River_2019.pdf){target="_blank"} (pg E3.3.3-78 to 82)

Instream spawning and rearing habitat for Fall Run Chinook Salmon and Steelhead in the Bear River are based on data from an instream flow study conducted by the South Sutter Water District. SSWD performed this study using River2D, a 2-dimensional hydraulic and habitat model. They used this model at two sites to approximate spawning and rearing availability in the Bear River. The U.S. Fish and Wildlife Service collected the same data at a third site, resulting in data for 3 separate reaches. The instream spawning and rearing habitat data were provided to us by [Mark Gard](mailto:mark_gard@fws.gov) from the U.S. Fish and Wildlife Service. The data were compiled into a spreadsheet. The three reaches were scaled by length and combined. The data were then converted into sqft/1000 ft.  

### Combine Spawning and Rearing Data
The instream spawning and juvenile rearing habitat data described above for Fall Run Chinook Salmon (FR) and Steelhead (ST) are combined for use in the Decision Support Model (DSM) in the following format. Units are in square feet per 1000 feet.
```{r,echo=FALSE}
# Bear river data 
bear_raw <- read_csv('data-raw/watershed/bear_river/data/updated_bear_river_instream.csv',
                     skip = 1)

bear_reach_1 <- bear_raw %>% 
  filter(reach == 1) 
bear_reach_2 <- bear_raw %>% 
  filter(reach == 2) 
bear_reach_3 <- bear_raw %>% 
  filter(reach == 3) 

total <- 3.4

bear_river_instream <- bear_reach_1 %>% bind_rows(bear_reach_2, bear_reach_3) %>% 
  gather(lifestage, sq_ft, -flow_cfs, -reach_km) %>%
  group_by(lifestage, flow_cfs) %>%
  summarise(wua = sum(sq_ft/reach_km * .3048 * reach_km/ total, na.rm = TRUE)) %>%
  spread(lifestage, wua) %>% 
  select(flow_cfs, FR_spawn_wua = FR_spawn_sqft, FR_juv_wua = FR_juv_sqft, 
         FR_fry_wua = FR_fry_sqft, ST_spawn_wua = ST_spawn_sqft, 
         ST_juv_wua = ST_juv_sqft, ST_fry_wua = ST_fry_sqft) %>%
  mutate(watershed = "Bear River")

knitr::kable(align = 'c', head(bear_river_instream, 5), 
caption = "Header Descriptions: flow_cfs = flow in cubic feet per second, 
FR_spawn_wua = Fall Run Chinook Spawning WUA, 
FR_juv_wua = Fall Run Chinook Juvenile WUA, 
FR_fry_wua = Fall Run Chinook Fry WUA, 
ST_spawn_wua = Steelhead Spawning WUA, 
ST_juv_wua = Steelhead Juvenile WUA, 
ST_fry_wua = Steelhead Fry WUA,
watershed = section of stream modeled for CVPIA SDM")
#usethis::use_data(bear_river_instream, overwrite = TRUE)
```

*...with 15 more rows.* 

### Spawning WUA 

The following plot shows the weighted usable spawning area (WUA) in square feet per thousand feet for Fall Run Chinook Salmon and Steelhead. These area per length rates are multiplied by the total spawning reach length mapped by the Science Integration Team (SIT) to estimate available habitat. 

``` {r,echo=FALSE}
bear_river_instream %>% 
  select(flow_cfs, `Fall Run Chinook` = FR_spawn_wua, `Steelhead` = ST_spawn_wua) %>% 
  gather(Species, WUA, -flow_cfs) %>% 
  filter(!is.na(WUA)) %>% 
  ggplot(aes(flow_cfs, WUA, color = Species)) +
  geom_line() +
  labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') + 
  theme_minimal() + 
  scale_color_manual(values = c('#d95f02','#7570b3')) +
  theme(legend.justification = c(1,0), legend.position = c(.9,.05))
```

### Rearing WUA
The following plot shows the rearing WUA in square feet per thousand feet for Fall Run Chinook Salmon and Steelhead juvenile and fry. These rates are multiplied by the total rearing reach length mapped by the SIT to estimate available habitat. 
``` {r,echo=FALSE}
bear_river_instream %>% 
  select(FR_juv_wua, FR_fry_wua, ST_juv_wua, ST_fry_wua, flow_cfs, watershed) %>%
  gather(species_lifestage, acres, -flow_cfs, -watershed) %>% 
  mutate(
    species = case_when( 
      str_detect(species_lifestage, 'ST') ~ 'Steelhead', 
      str_detect(species_lifestage, 'FR') ~ 'Fall Run'),
      lifestage = ifelse(str_detect(species_lifestage, 'juv'), 'Juvenile', 'Fry')) %>% 
  ggplot(aes(flow_cfs, acres, color = species)) +
  geom_line() +
  theme_minimal() +
  facet_wrap(~lifestage, ncol = 1, scales = 'free_y') +
  scale_color_brewer(palette = 'Dark2') +
  labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') + 
  theme(legend.justification = c(1,0), legend.position = c(.95,.8))
```

```{r, include=FALSE}
source('data-raw/watershed/floodplain_utils.R')
```

## Floodplain Rearing Habitat
**Data Source:** [Central Valley Floodplain Evaluation and Delineation (CVFED) HEC-RAS hydraulic model](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/CombinedTM_IQAR_Final-FULL-REPORT_20140206.pdf){target="_blank"}

`r print_model_details('Bear River', 'fr')`

## Floodplain Data
The areas represent total inundated area in acres.
```{r, echo= FALSE}
bear_fp <- read_excel('data-raw/watershed/CVPIA_FloodplainAreas.xlsx', sheet = 'BearRiver') %>%
  mutate(watershed = 'Bear River')

threshold <- bear_fp %>% 
  filter(modeled_floodplain_area_acres == 0) %>% 
  summarise(max = max(flow_cfs)) %>% 
  pull(max)

bear_df <- bear_fp %>% 
  filter(flow_cfs >= threshold)

bear_river_floodplain_fr <- 
  scale_fp_flow_area_partial_model(ws = 'Bear River', df = bear_df) 

bear_river_floodplain <- bear_river_floodplain_fr %>% 
  mutate(ST_floodplain_acres = FR_floodplain_acres) %>% 
  select(flow_cfs, FR_floodplain_acres, ST_floodplain_acres, watershed)

knitr::kable(align = 'c', head(bear_river_floodplain, 5), 
             caption = "Header Descriptions: flow_cfs = flow in cubic feet per second,
             FR_floodplain_acres = Fall Run Chinook floodplain acres,
             ST_floodplain_acres = Steelhead floodplain acres,
             watershed = section of stream modeled for CVPIA SDM") 
# usethis::use_data(bear_river_floodplain, overwrite = TRUE)
```


*...with 30 more rows*

## Floodplain Plot
```{r, echo = FALSE}
bear_river_floodplain %>% 
  select(flow_cfs, FR_floodplain_acres, watershed) %>% 
  rename(`Fall Run Chinook and Steelhead` = FR_floodplain_acres) %>% 
  gather(Species, acres, -flow_cfs, -watershed) %>% 
  ggplot(aes(flow_cfs, acres, color = Species)) +
  geom_line() +
  theme_minimal() +
  scale_color_brewer(palette = 'Dark2') +
  labs(x = 'Flow (cfs)', y = 'Total Inundated Acres') +
  theme(legend.justification = c(1,0), legend.position = c(.9,.05)) 
```
