---
title: "Mokelumne River"
author: "sadie"
date: "December 11, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
library(scales)
```



robyn steelhead rearing coming mid-January 

## Instream Spawning and Rearing Habiat
**Data Source:** EBMUD 2016 and CDFW 1998
<https://flowwest.github.io/cvpiaHabitat/reference/mokelumne_river_instream.html>

### EBMUD Raw Data

'LMR_habitat__s_v2mw_to_MT.xlsx'
spawning: 'Spawning WUA' tab A4:C8 
rearing: 'Instream WUA' tab A2:C8 and A12:C14 and A18:c27 weighted average, check F6 and H6 to compare outcome at 300 cfs
floodplain: 'Floodplain (new flows)' tab C21:D26 Camanche to tidal low estimate, 400 cfs = 0 total acres

#### Spawning WUA
```{r, echo=FALSE}
# spawning reach length
moke_spawn_len <- cvpiaHabitat::watershed_lengths %>% 
  filter(watershed == 'Mokelumne River', lifestage == 'spawning') %>% 
  pull(feet)

# 'LMR_habitat__s_v2mw_to_MT.xlsx'
# spawning: 'Spawning WUA' tab A4:C8 
spawning <- read_excel('data-raw/instream/mokelumne/LMR_habitat__s_v2mw_to_MT.xlsx', 
                       sheet = 'Spawning WUA', range = 'A4:C8', 
                       col_names = c('flow_cfs', 'FR_spawn_sqft', 'ST_spawn_sqft')) %>% 
  mutate(FR_spawn_wua = FR_spawn_sqft/(moke_spawn_len / 1000), #sqft_per_1000ft
         ST_spawn_wua = ST_spawn_sqft/(moke_spawn_len / 1000)) %>% 
  select(flow_cfs, FR_spawn_wua, ST_spawn_wua)

spawning
```
### Spawning WUA

```{r, echo=FALSE}
spawning %>% 
  gather(species, wua, -flow_cfs) %>% 
  ggplot(aes(flow_cfs, wua, color = species)) +
  geom_line() +
  labs(x = 'flow (cfs)', y = 'spawning WUA (sqft/1000ft)') +
  scale_y_continuous(labels = comma)

# devtools::use_data(mokelumne_river_instream, overwrite = TRUE)
```

### Rearing WUA
```{r}
# spawning reach length
moke_rear_len <- cvpiaHabitat::watershed_lengths %>% 
  filter(watershed == 'Mokelumne River', lifestage == 'rearing') %>% 
  pull(feet)

# 'LMR_habitat__s_v2mw_to_MT.xlsx'
# rearing: 'Instream WUA' tab A2:C8 and A11:C14 and A17:C27 weighted average, check F6 and H6 to compare outcome at 300 cfs
total_modeled_len <- 1.1 + 9.4 + 24.7

rearing1 <- read_excel('data-raw/instream/mokelumne/LMR_habitat__s_v2mw_to_MT.xlsx', 
                       sheet = 'Instream WUA', range = 'A2:C8') 
rearing2 <- read_excel('data-raw/instream/mokelumne/LMR_habitat__s_v2mw_to_MT.xlsx', 
                       sheet = 'Instream WUA', range = 'A11:C14')
rearing3 <- read_excel('data-raw/instream/mokelumne/LMR_habitat__s_v2mw_to_MT.xlsx', 
                       sheet = 'Instream WUA', range = 'A17:C27')

Flow <- unique(sort(c(rearing1$Flow, rearing2$Flow, rearing3$Flow)))

# reach 1 ------------
r1 <- tibble(
  Flow = Flow,
  `Fall-run fry (ft2)` = approx(rearing1$Flow, rearing1$`Fall-run fry (ft2)`, 
                                xout = flow_out, rule = 2)$y,
  `Fall-run juvenile (ft2)` = approx(rearing1$Flow, rearing1$`Fall-run juvenile (ft2)`, 
                                     xout = flow_out, rule = 2)$y,
  rm = 1.1
)

rearing1 %>% 
  ggplot(aes(Flow, `Fall-run fry (ft2)`)) +
  geom_point()

r1 %>% 
  ggplot(aes(Flow, `Fall-run fry (ft2)`)) +
  geom_point()

# reach 2 ----------
r2 <- tibble(
  Flow = Flow,
  `Fall-run fry (ft2)` = approx(rearing2$Flow, rearing2$`Fall-run fry (ft2)`, 
                                xout = Flow, rule = 2)$y,
  `Fall-run juvenile (ft2)` = approx(rearing2$Flow, rearing2$`Fall-run juvenile (ft2)`, 
                                     xout = Flow, rule = 2)$y,
  rm = 9.4
)

rearing2 %>% 
  ggplot(aes(Flow, `Fall-run fry (ft2)`)) +
  geom_point()

r2 %>% 
  ggplot(aes(Flow, `Fall-run fry (ft2)`)) +
  geom_point()

# reach 3 ----------
  # mutate(FR_spawn_wua = FR_spawn_sqft/(moke_spawn_len / 1000), #sqft_per_1000ft
  #        ST_spawn_wua = ST_spawn_sqft/(moke_spawn_len / 1000)) %>% 
  # select(flow_cfs, FR_spawn_wua, ST_spawn_wua)
r3 <- tibble(
  Flow = Flow,
  `Fall-run fry (ft2)` = approx(rearing3$Flow, rearing3$`Fall-run fry (ft2)`, 
                                xout = Flow, rule = 2)$y,
  `Fall-run juvenile (ft2)` = approx(rearing3$Flow, rearing3$`Fall-run juvenile (ft2)`, 
                                     xout = Flow, rule = 2)$y,
  rm = 24.7
)

rearing3 %>% 
  ggplot(aes(Flow, `Fall-run fry (ft2)`)) +
  geom_point()

r3 %>% 
  ggplot(aes(Flow, `Fall-run fry (ft2)`)) +
  geom_point()

# combine reaches
rearing <- r1 %>% 
  bind_rows(r2, r3) %>% 
  gather(lifestage, sq_ft, -Flow, -rm) %>%
  group_by(lifestage, Flow) %>% 
  summarise(wua = sum(sq_ft/rm/5.28 * rm/total_modeled_len)) %>% 
  spread(lifestage, wua) %>% 
  rename(flow_cfs = Flow, FR_fry_wua = `Fall-run fry (ft2)`, FR_juv_wua = `Fall-run juvenile (ft2)`)

# 300 cfs fry, double check calculation
(203823/1.1/5.28 * 1.1/35.2 + 1718536/9.4/5.28 * 9.4/35.2 + 1697397/24.7/5.28 * 24.7/35.2) -
(rearing %>% 
  filter(flow_cfs == 300) %>% 
  pull(FR_fry_wua))
```
