---
title: "Mokelumne River"
author: "[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com)"
date: "July 2020"
output: html_document
---
```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, out.width = '100%')
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
library(scales)
```

## Future Data Improvements

* EBMUD is developing new steelhead rearing habitat flow:area relationships (expected 2020)
* EBMUD is developing new Fall Run floodplain rearing habitat flow:area relationships with Mark Gard (expected 2020)

## Instream Spawning and Rearing Habitat
**Data Sources:** [CDFW 1991](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/Lower+Mokelumne+RIver+Fisheries+Management+Plan%2C+CDFG+1991.pdf){target="_blank"} (pg. 105 - 107) 

Instream spawning and rearing habitat for Fall Run Chinook Salmon and Steelhead in the Mokelumne River is based on data from USFWS's 1998 instream flow evaluation and EBMUD's 2016 instream habitat evaluation. The East Bay Municipal Utility Distract (EBMUD) instream spawning and rearing habitat data were provided in a spreadsheet. Together, these evaluations determined relationships between flow and both suitable spawning and rearing habitat for reaches spanning approximately 35 miles of the Mokelumne River. 

### Combine Instream Spawning and Rearing Habitat Data
The instream spawning and rearing habitat data described above are combined for use in the Decision Support Model (DSM) in the following format. Units are in square feet per 1000 feet. 
```{r, message=FALSE}
# spawning reach length
moke_spawn_len <- cvpiaHabitat::watershed_lengths %>% 
  filter(watershed == 'Mokelumne River', lifestage == 'spawning') %>% 
  pull(feet)

# 'LMR_habitat__s_v2mw_to_MT.xlsx'
# spawning: 'Spawning WUA' tab A4:C8 
spawning_moke <- read_excel('data-raw/watershed/mokelumne/data/LMR_habitat__s_v2mw_to_MT.xlsx', 
                       sheet = 'Spawning WUA', range = 'A4:C8', 
                       col_names = c('flow_cfs', 'FR_spawn_sqft', 'ST_spawn_sqft')) %>% 
  mutate(FR_spawn_wua = FR_spawn_sqft/(moke_spawn_len / 1000), #sqft_per_1000ft
         ST_spawn_wua = ST_spawn_sqft/(moke_spawn_len / 1000)) %>% 
  select(flow_cfs, FR_spawn_wua, ST_spawn_wua)
```

```{r,message=FALSE}
# rearing reach length
moke_rear_len <- cvpiaHabitat::watershed_lengths %>% 
  filter(watershed == 'Mokelumne River', lifestage == 'rearing') %>% 
  pull(feet)

# 'LMR_habitat__s_v2mw_to_MT.xlsx'
# rearing: 'Instream WUA' tab A2:C8 and A11:C14 and A17:C27 weighted average
total_modeled_len <- 1.1 + 9.4 + 24.7

rearing1 <- read_excel('data-raw/watershed/mokelumne/data/LMR_habitat__s_v2mw_to_MT.xlsx', 
                       sheet = 'Instream WUA', range = 'A2:C8') 
rearing2 <- read_excel('data-raw/watershed/mokelumne/data/LMR_habitat__s_v2mw_to_MT.xlsx', 
                       sheet = 'Instream WUA', range = 'A11:C14')
rearing3 <- read_excel('data-raw/watershed/mokelumne/data/LMR_habitat__s_v2mw_to_MT.xlsx', 
                       sheet = 'Instream WUA', range = 'A17:C27')

Flow <- unique(sort(c(rearing1$Flow, rearing2$Flow, rearing3$Flow)))

# reach 1 ------------
r1 <- tibble(
  Flow = Flow,
  `Fall-run fry (ft2)` = approx(rearing1$Flow, rearing1$`Fall-run fry (ft2)`, 
                                xout = Flow, rule = 2)$y,
  `Fall-run juvenile (ft2)` = approx(rearing1$Flow, rearing1$`Fall-run juvenile (ft2)`, 
                                     xout = Flow, rule = 2)$y,
  rm = 1.1
)

# reach 2 ----------
r2 <- tibble(
  Flow = Flow,
  `Fall-run fry (ft2)` = approx(rearing2$Flow, rearing2$`Fall-run fry (ft2)`, 
                                xout = Flow, rule = 2)$y,
  `Fall-run juvenile (ft2)` = approx(rearing2$Flow, rearing2$`Fall-run juvenile (ft2)`, 
                                     xout = Flow, rule = 2)$y,
  rm = 9.4
)

# reach 3 ----------
r3 <- tibble(
  Flow = Flow,
  `Fall-run fry (ft2)` = approx(rearing3$Flow, rearing3$`Fall-run fry (ft2)`, 
                                xout = Flow, rule = 2)$y,
  `Fall-run juvenile (ft2)` = approx(rearing3$Flow, rearing3$`Fall-run juvenile (ft2)`, 
                                     xout = Flow, rule = 2)$y,
  rm = 24.7
)

# combine reaches
rearing <- r1 %>% 
  bind_rows(r2, r3) %>% 
  gather(lifestage, sq_ft, -Flow, -rm) %>%
  group_by(lifestage, Flow) %>% 
  summarise(wua = sum(sq_ft/rm/5.28 * rm/total_modeled_len)) %>% 
  spread(lifestage, wua) %>% 
  rename(flow_cfs = Flow, FR_fry_wua = `Fall-run fry (ft2)`, FR_juv_wua = `Fall-run juvenile (ft2)`)
```
```{r, echo = FALSE}
#combine spawning and rearing 
mokelumne_river_instream <- spawning_moke %>% 
  full_join(rearing) %>% 
  arrange(flow_cfs) %>% 
  mutate(watershed = 'Mokelumne River') %>% 
  filter(flow_cfs > 0) %>% 
  select(flow_cfs, FR_spawn_wua, FR_fry_wua, FR_juv_wua, ST_spawn_wua, watershed)

#usethis::use_data(mokelumne_river_instream, overwrite = TRUE)
```

```{r, echo = FALSE,message=FALSE}
knitr::kable(align = 'c' , head(mokelumne_river_instream, 5), 
             caption = "Header Descriptions: flow_cfs = flow in cubic feet per second,
             FR_spawn_wua = Fall Run Chinook Spawning wua,
             FR_fry_wua = Fall Run Chinook Fry wua,
             FR_juv_wua = Fall Run Chinook Juvenile wua,
             ST_spawn_wua = Steelhead Spawning wua,
             watershed = section of stream modeled for CVPIA SDM")
```

*...with 47 more rows.*

### Spawning WUA   
The following plot shows the weighted usable spawning area (WUA) rate in square feet per thousand feet for Fall Run Chinook Salmon (FR). 
```{r, echo=FALSE}
spawning_moke %>% 
  select(flow_cfs, `Fall Run Chinook` = FR_spawn_wua, `Steelhead` = ST_spawn_wua) %>% 
  gather(Species, WUA, -flow_cfs) %>% 
  filter(!is.na(WUA)) %>% 
  ggplot(aes(flow_cfs, WUA, color = Species)) +
  geom_line() +
  labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') + 
  theme_minimal() + 
  scale_color_manual(values = c('#d95f02','#7570b3')) +
  theme(legend.justification = c(1,0), legend.position = c(.9,.3))
```

### Rearing WUA
The flow:area relationships from USFWS and EBMUD were first expanded using linear interpolation to ensure the relationships for each evaluation reach included the same flows. A single rearing habitat area representing the entire river was then estimated for each flow. Each area was weighted in proportion to the length of the evaluation reach from which the area was originally estimated. The resulting areas were then converted to an area per unit length by dividing the area by the length of the rearing reach mapped by the Scince Integration Team (SIT).

The following plot shows the rearing WUA in square feet per thousand feet for Fall Run Chinook Salmon fry and juvenile.
``` {r,echo=FALSE}
mokelumne_river_instream %>% 
  gather(Lifestage, wua, -flow_cfs, -watershed)  %>% 
  filter(!is.na(wua), Lifestage != 'FR_spawn_wua', Lifestage != 'ST_adult_wua', Lifestage != "ST_spawn_wua", Lifestage != "ST_juv_wua", Lifestage != "ST_fry_wua") %>%
  mutate(Lifestage = ifelse(Lifestage == 'FR_fry_wua', 'Fry', 'Juvenile')) %>% 
  ggplot(aes(x = flow_cfs , y = wua, color = Lifestage)) +
  geom_line() +
  labs(x = 'Flow (cfs)', y = 'WUA (sqft/1000ft)') + theme_minimal() +
  scale_color_manual(values = c('#d95f02','#7570b3')) +
  theme(legend.justification = c(1,0), legend.position = c(.9,.3))
```

## Floodplain Rearing Habitat

**Data Source:** [CVPIA Annual Progress Report Fiscal Year 2019](https://cvpiahabitat-r-package.s3-us-west-2.amazonaws.com/CVPIA+Annual+Progress+Report+Fiscal+Year+2019.pdf){target="_blank"}

The flow : floodplain area relationship for the Mokelumne River is based on high resolution (4 square foot cell size) two-dimensional hydraulic modeling completed by EBMUD in 2017 (Technical Memorandum in preparation by Robyn Bilski at EBMUD). The Mokelumne model was broken into seven reaches (Figure 14) based on the availability of stage-discharge relationships. The extent of this modeling is from River Mile 62.8 to River Mile 64. All area inundated flows less than or equal to 400 cfs were considered as instream.  The inundated area per length value from this modeled reach was extrapolated to the entire rearing reach in the Mokelumne River, from Camanche Dam downstream to the point of tidal influence. This flow : floodplain area relationship was used for all salmon and steelhead runs in the Mokelumne River. The following table contains the flow : floodplain area relationships used in the Decision Support Model (DSM). 
```{r, echo=FALSE}
#'LMR_habitat__s_v2mw_to_MT.xlsx', 'Floodplain (new flows)' tab C21:D26 Camanche to tidal #low estimate
#400 cfs = 0 acres 

mokelumne_river_floodplain <- read_excel('data-raw/watershed/mokelumne/data/LMR_habitat__s_v2mw_to_MT.xlsx', 
                      sheet = 'Floodplain (new flows)', range = 'C21:D26',
                      col_names = c('flow_cfs', 'FR_floodplain_acres')) %>% 
 bind_rows(tibble(flow_cfs = 400, FR_floodplain_acres = 0)) %>% 
 arrange(flow_cfs) %>% 
 mutate(SR_floodplain_acres = FR_floodplain_acres,
        ST_floodplain_acres = FR_floodplain_acres,
        watershed = "Mokelumne River")

knitr::kable(align ='c', head(mokelumne_river_floodplain, 5), 
             caption = "Header Descriptions: flow_cfs = flow in cubic feet per second,
             FR_floodplain_acres = Fall Run Chinook floodplain acres, SR_floodplain_acres
             = Spring Run Chinook floodplain acres, ST_floodplain_acres = 
             Steelhead floodplain acres, 
             watershed = section of stream modeled for CVPIA SDM")
# usethis::use_data(mokelumne_river_floodplain, overwrite = TRUE)
```

*...with 2 more rows.*

## Floodplain Plot
```{r, echo = FALSE}
mokelumne_river_floodplain %>% 
  select(FR_floodplain_acres, flow_cfs, watershed) %>%
  rename(`Fall Run Chinook,
Spring Run Chinook, 
and Steelhead` = FR_floodplain_acres) %>% 
  gather(Species, acres, -flow_cfs, -watershed) %>% 
  ggplot(aes(flow_cfs, acres, color = Species)) +
  geom_line() +
  theme_minimal() +
  scale_color_brewer(palette = 'Dark2') +
  labs(x = 'Flow (cfs)', y = 'Total Inundated Acres') +
  theme(legend.justification = c(1,0), legend.position = c(.9,.1))
```

