---
title: "Paynes Creek"
author: "[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com)" 
date: "July 2020"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, out.width = '100%')
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
```

## Future Data Improvements
None planned.  

## Instream Spawning and Rearing Habitat
**Data Source:** [Calculated with a relationship between flow and mean weighted usable area from watersheds in the region with modeled habitat](http://cvpia-habitat-docs-markdown.s3-website-us-west-2.amazonaws.com/watershed/Regional_Approximation.html){target="_blank"}
```{r, include=FALSE}
source('data-raw/watershed/regional_approximation/description_regional_approx.R')
```

`r print_regional_approx('paynes Creek')`

```{r, include=FALSE}
source('data-raw/watershed/floodplain_utils.R')
```

## Floodplain Rearing Habitat

**Data Source:** [Scaled from a Deer Creek flow to floodplain area relationship generated with a 1D HEC-RAS hydraulic model](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/DRAFT+Deer+Creek+Hydraulic+Models+Tech+Memo+6-08-07.pdf){target="_blank"}

### Fall Run
`r print_model_details('Paynes Creek', 'fr')`

### Steelhead
`r print_model_details('Paynes Creek', 'st')`

## Floodplain Data
The areas represent total inundated area in acres.    
```{r}
# TODO Ask mark about the difference in outcome from his sheet
paynes_creek_floodplain <- scale_fp_flow_area('Paynes Creek')

knitr::kable(align = 'c', head(paynes_creek_floodplain, 5), 
             caption = "Header Descriptions: flow_cfs = flow in cubic feet per second,
             FR_floodplain_acres = Fall Run Chinook floodplain acres, 
             ST_floodplain_acres = Steelhead floodplain acres,
             watershed = section of stream modeled for CVPIA SDM") 
# devtools::use_data(paynes_creek_floodplain, overwrite = TRUE)
```

*...with 17 more rows*

## Floodplain Plot
```{r}
paynes_creek_floodplain %>% 
  rename(`Fall Run Chinook` = FR_floodplain_acres,
         `Steelhead` = ST_floodplain_acres) %>% 
  gather(Species, acres, -flow_cfs, -watershed) %>% 
  ggplot(aes(flow_cfs, acres, color = Species)) +
  geom_line() +
  theme_minimal() +
  scale_color_brewer(palette = 'Dark2') +
  labs(x = 'Flow (cfs)', y = 'Total Inundated Acres') +
  theme(legend.justification = c(1,0), legend.position = c(1,.1))
```
