---
title: "Butte Creek"
author: "sadie"
date: "December 13, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
```

## Instream Spawning and Rearing Habiat

> **Data Source:**   TODO
> [FWS and 2008 FERC relicensing of DeSabla](https://s3-us-west-2.amazonaws.com/cvpiahabitat-r-package/cvpia-sit-model-inputs/DeSabla2008ButteIFIM.pdf)
> <https://flowwest.github.io/cvpiaHabitat/reference/butte_creek_instream.html>

Mark Gard instructed us to use cells A3:B33 for spring run spawning Above Centerville Powerhouse Segment (6.5 mi) and cells A38:B68 for the reach Below Centerville Powerhouse Segment.
Units are in square feet.   

Rearing habitat is inn the same sheet in cells A71:D101. Units are square feet per 1000 ft.

```{r}

above_centerville <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', 
                                range = "A3:B33", sheet = 'Butte') 
below_centerville <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', 
                                range = "A38:B68", sheet = 'Butte') 

spawn <- bind_rows(above_centerville, below_centerville) %>% 
  group_by(Flow) %>% 
  summarise(SR_spawn_wua = sum(Spawning)/(6.5 + 9)/5.28) %>% 
  ungroup()

rear <- read_excel('data-raw/mark_gard_data/IFIMWUA.xlsx', 
                   range = "A71:D101", sheet = 'Butte') 

butte_creek_instream <- spawn %>% 
  full_join(rear) %>% 
  mutate(FR_spawn_wua = SR_spawn_wua, watershed = 'Butte Creek') %>% 
  select(flow_cfs = Flow, FR_spawn_wua, FR_fry_wua = `Fry Rearing`, 
         FR_juv_wua = `Juv Rearing`, adult_trout_wua = `adult trout`, 
         watershed)

butte_creek_instream
# devtools::use_data(butte_creek_instream, overwrite = TRUE)
```

