---
title: "Bypasses"
author: "[Sadie Gill](mailto:sgill@flowwest.com), [Mark Tompkins](mailto:mtompkins@flowwest.com)"
date: "July 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(cvpiaHabitat)
library(scales)
```

## Modeling Details
**Data Source:**
Central Valley Floodplain Evaluation and Delineation (CVFED) HEC-RAS hydraulic model refined for use in the [NOAA-NMFS Winter Run Chinook Salmon life cycle model](https://cvpiahabitat-r-package.s3-us-west-2.amazonaws.com/2017Hendrix.pdf). The ESRI [shapefile](https://cvpiahabitat-r-package.s3-us-west-2.amazonaws.com/delta_bypass_extents.zip) and Google Earth [kmz file](https://cvpiahabitat-r-package.s3-us-west-2.amazonaws.com/Sutter+Yolo+Delta+Habitat+Extents.kmz) show the delineation of the bypass and Delta regions used to generate habitat inputs.

High quality habitat defined by:

* Channel depth > 0.2 m and < 1.5 m
* Velocity <= 0.15 m/s

## Sutter Bypass

The Sutter Bypass is divided into four sections:

* Section 1 - Butte Sink from Rotavele to Moulton Weir
* Section 2 - Butte Sink from Moulton Weir to Colusa Weir
* Section 3 - Sutter Bypass from Colusa Weir to Tisdale Weir
* Section 4 - Sutter Bypass from Tisdale Weir to the Sacramento River

Area is in square meters.

```{r}
# 1 sq ft = 0.092903 sq meters
sutter1 <- read_excel('data-raw/correigh_greene_data/River Rearing_Habitat vs flow.xls', sheet = 2) %>% 
  mutate(flow_cfs = Flow_cfs,
         total_sq_meters = Sutter_Bypass_1_Pref11_TotlArea_Sq_ft * 0.092903,
         bypass = 'Sutter Bypass 1') %>%
  select(flow_cfs, total_sq_meters, bypass)

sutter2 <- read_excel('data-raw/correigh_greene_data/River Rearing_Habitat vs flow.xls', sheet = 3) %>%
  mutate(flow_cfs = Flow_cfs,
         total_sq_meters = Sutter_Bypass_2_Pref11_TotlArea_Sq_ft * 0.092903,
         bypass = 'Sutter Bypass 2') %>%
  select(flow_cfs, total_sq_meters, bypass)

sutter3 <- read_excel('data-raw/correigh_greene_data/River Rearing_Habitat vs flow.xls', sheet = 4) %>%
  mutate(flow_cfs = Flow_cfs,
         total_sq_meters = Sutter_Bypass_3_Pref11_TotlArea_Sq_ft * 0.092903,
         bypass = 'Sutter Bypass 3') %>%
  select(flow_cfs, total_sq_meters, bypass)

sutter4 <- read_excel('data-raw/correigh_greene_data/River Rearing_Habitat vs flow.xls', sheet = 5) %>%
  mutate(flow_cfs = Flow_cfs,
         total_sq_meters = Sutter_Bypass_4_Pref11_TotlArea_Sq_ft * 0.092903,
         bypass = 'Sutter Bypass 4') %>%
  select(flow_cfs, total_sq_meters, bypass)

sutter_bypass_habitat <- bind_rows(sutter1, sutter2, sutter3, sutter4) %>% 
  spread(bypass, total_sq_meters) 

knitr::kable(align = 'c', head(sutter_bypass_habitat, 5))

# devtools::use_data(sutter_bypass_habitat)
```

*... with 36 more rows*

## Sutter Bypass Plot

```{r, out.width = '90%'}
sutter_bypass_habitat %>% 
  gather(Section, square_meters, -flow_cfs) %>% 
  ggplot(aes(flow_cfs, square_meters / 4046.86, color = Section)) +
  geom_line() +
  theme_minimal() +
  scale_color_brewer(palette = 'Dark2') +
  scale_y_continuous(labels = comma) +
  labs(x = 'Flow (cfs)', y = 'Suitable Habitat Acres') +
  theme(legend.position = "bottom")
```

## Yolo Bypass

The Yolo Bypass is divided into two sections:

* Section 1 - Fremont Weir to Sacramento Weir
* Section 2 - Below Sacramento Weir

Area is in square meters.

```{r}
yolo1 <- read_excel('data-raw/correigh_greene_data/River Rearing_Habitat vs flow.xls', sheet = 6) %>%
  mutate(flow_cfs = Flow_cfs,
         total_sq_meters = Yolo_Bypass_1_Pref11_TotlArea_Sq_ft * 0.092903,
         bypass = 'Yolo Bypass 1') %>%
  select(flow_cfs, total_sq_meters, bypass)

yolo2 <- read_excel('data-raw/correigh_greene_data/River Rearing_Habitat vs flow.xls', sheet = 7) %>%
  mutate(flow_cfs = Flow_cfs,
         total_sq_meters = Yolo_Bypass_2_Pref11_TotlArea_Sq_ft * 0.092903,
         bypass = 'Yolo Bypass 2') %>%
  select(flow_cfs, total_sq_meters, bypass)

yolo_bypass_habitat <- bind_rows(yolo1, yolo2) %>% 
  spread(bypass, total_sq_meters)

knitr::kable(align = 'c', head(yolo_bypass_habitat, 5))

# devtools::use_data(yolo_bypass_habitat)
```

*... with 36 more rows*

## Yolo Bypass Plot

```{r, out.width = '90%'}
yolo_bypass_habitat %>% 
  gather(Section, square_meters, -flow_cfs) %>% 
  ggplot(aes(flow_cfs, square_meters / 4046.86, color = Section)) +
  geom_line() +
  theme_minimal() +
  scale_color_brewer(palette = 'Dark2') +
  scale_y_continuous(labels = comma) +
  labs(x = 'Flow (cfs)', y = 'Suitable Habitat Acres') +
  theme(legend.position = "bottom")
```
